### 2015-10-26 analysis/template.Rmd
add template.Rmd and chunk options

---
title: "Untitled"
author: "First Last"
date: YYYY-MM-DD
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

This is a template for writing reports with R Markdown, copied from @jdblischak.

## Section title

You can create a new file from the template using the command `cp`.

```bash
cp analysis/template.Rmd analysis/newfile.Rmd
```

## Session information

```{r info}
sessionInfo()
```