### 2015-10-26 analysis/summarize_dsc_robust.Rmd
add summarize_dsc_robust.Rmd

---
title: "Summarize results on robustness of EM algorithm"
author: "Matthew Stephens"
date: 2015-10-26
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

Here we summarize the results of the simulation study to look at convergence of ash.
```{r}
  library(dscr)
  library(ashr)
  library(ggplot2)
  library(reshape2)
  library("dplyr")
  load("../output/dsc-robust-files/dsc_robust.RData") 
  res = dsc_robust$res
```

```{r}
  quantile((res %>% filter(method=="ash.n"))$score.diff1,c(0.8,0.9,0.95,0.99,1))
  quantile((res %>% filter(method=="ash.hu"))$score.diff1,c(0.8,0.9,0.95,0.99,1))
  quantile((res %>% filter(method=="ash.u"))$score.diff1,c(0.8,0.9,0.95,0.99,1))

  plot(ecdf((res %>% filter(method %in% c("ash.u","ash.hu")))$score.diff1),ylim=c(0.8,1))
  plot(ecdf((res %>% filter(method %in% c("ash.n")))$score.diff1),ylim=c(0.8,1))
```

```{r}
  mean((res %>% filter(method %in% c("ash.u","ash.hu")))$score.diff1<0.02)
  mean((res %>% filter(method %in% c("ash.u","ash.hu")))$score.diff1>1)
  mean((res %>% filter(method %in% c("ash.n")))$score.diff1<0.02)  
```

Now check out the example that showed the biggest convergence problem
```{r}
#check out which one showed biggest convergence problem
prob = res[which.max(res$score.diff1),]
setwd("../output")
prob_eg = loadExample(dsc_robust,seed=prob$seed,scenario=as.character(prob$scenario),method=as.character(prob$method))
setwd("../analysis")
```

Run ash on the example to check we can recapitulate the convergence problem.
```{r}
set.seed(prob$seed+1) #dscr sets the seed for the method to 1+seed 
ash.res=list()
ash.res[[1]] = ash(prob_eg$input$betahat,prob_eg$input$sebetahat,mixcompdist="uniform",cxx=TRUE)
for(i in 2:11){ash.res[[i]] = ash(prob_eg$input$betahat,prob_eg$input$sebetahat,mixcompdist="uniform",randomstart=TRUE,cxx=TRUE)}
```

```{r}
loglik = mapply(get_loglik,ash.res)
loglik
```

```{r}
best=which.max(loglik)
plot(ash.res[[best]]$lfsr,ash.res[[1]]$lfsr,xlab="lfsr (best run)",ylab="lfsr (default run)",main="Comparison of lfsr (best vs default run)")
abline(a=0,b=1,col=2)
plot(ash.res[[best]]$PosteriorMean,ash.res[[1]]$PosteriorMean,xlab="Posterior Mean (best run)",ylab="Posterior Mean (default run)",main="Comparison of Posterior Means (best vs default run)")
abline(a=0,b=1,col=2)

```

Compare Cdfs:
```{r}
  x = seq(-5,5,length=100)
  cdf_res = data.frame(x=x,
                     best=as.numeric(cdf.ash(ash.res[[best]],x)$y),
                     default=as.numeric(cdf.ash(ash.res[[1]],x)$y))
  cdf_res=melt(cdf_res,id.vars="x",value.name="y",variable.name="run")
  qplot(x=x,y=y,data=cdf_res,geom="line",colour=run,main="comparison of estimated cdf (best run and default run)")
```

## Session information

```{r info}
sessionInfo()
```