### 2015-10-26 analysis/check_mixfdr_lownoise.Rmd

---
title: "Check mixfdr in low noise simulation"
author: "Matthew Stephens"
date: 2015-10-26
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

Here we check that mixfdr actually does OK at fitting the bimodal distribution when 
noise is small compared with bimodality.

First simulate bimodal data with well-separated peaks. The histogram
shows the multi-modal nature of the underlying g is clear.
```{r}
  source("../code/dsc-shrink/datamakers/datamaker.R")
  sim.bimodal=rnormmix_datamaker(
              args=list(g=normalmix(c(0.5,0.5),c(-5,5),c(1,1)),
              min_pi0=0.4,
              max_pi0=0.4,
              nsamp=1000,
              betahatsd=1)
            )
  hist(sim.bimodal$input$betahat,nclass=20,xlim=c(-10,10),prob=TRUE,main="histogram of simulated betahat values")
```

Now run mixfdr on it. Note that the fit captures the multi-modal nature of g, 
although mixfdr continues to overestimate pi0.
```{r}
  source("../code/dsc-shrink/methods/mixfdr.wrapper.R")
  sim.bimodal.mixfdroutput = mixfdr.wrapper(sim.bimodal$input,args=list(theonull=TRUE))
  g = mixfdr2fitted.g(sim.bimodal.mixfdroutput)$fitted.g
  x=seq(-10,10,length=100)
  plot(ecdf(sim.bimodal$meta$beta),main="Cdf of true beta: empirical (black) and fitted (red)")
  lines(x,mixcdf(g,x),col=2)
```

## Session information

```{r info}
sessionInfo()
```