### 2015-10-26 analysis/plot_egdens.Rmd

---
title: "Plot densities for simulation studies"
author: "Matthew Stephens"
date: 2015-10-26
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

```{r}
library("ashr")
library("ggplot2")
library("dplyr")
library("dscr")
load("../output/dsc-shrink-files/res.RData")
PLOTSCENARIOS = c("spiky","near-normal","flat-top","skew","big-normal","bimodal")
PLOTNAMES = PLOTSCENARIOS
#PLOTFILE = "../paper/figures/scenarios_density.pdf"

df=data.frame()

for(i in PLOTSCENARIOS){
  s=dsc_shrink$scenarios[[i]]
  g=s$args$g
  x = seq(-6,6,length=100)
  y = as.numeric(dens(g,x))
  df = rbind(df,data.frame(x=x,y=y,scenario=i))
}

df$scenario = factor(df$scenario,levels=PLOTSCENARIOS)
levels(df$scenario)= PLOTNAMES
```

```{r scenario_density, dev='pdf',fig.height=3, fig.width=9, crop=TRUE}
#pdf(PLOTFILE,height=3,width=9)
ggplot(df, aes(x=x,y=y)) + geom_line(size=1.2,linetype=1) + facet_grid(.~scenario) + ylab("density")
#dev.off()
```

## Session information

```{r info}
sessionInfo()
```