### 2015-10-26 analysis/plot_pi0est.Rmd

---
title: "Plot estimates of pi0"
author: "Matthew Stephens"
date: 2015-10-26
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

```{r plot_pi0est, dev='pdf', crop=TRUE}
load("../output/dsc-shrink-files/res.RData")
source("../R/set_plot_colors.R")
library(dplyr)

PLOTMETHODS= c("mixfdr.tnull","ash.n","ash.u","qvalue","locfdr")
PLOTSCENARIOS = c("spiky","near-normal","flat-top","skew","big-normal","bimodal")
PLOTNAMES = PLOTSCENARIOS
ALPHALEVEL = 0.8 # controls transparency

df = res$pi0_score %>% filter(scenario %in% PLOTSCENARIOS) %>% filter(method %in% PLOTMETHODS)
df$scenario = factor(df$scenario,levels=PLOTSCENARIOS)
levels(df$scenario)= PLOTNAMES

pi0_plot=ggplot(df,
         aes(pi0,pi0_est,colour=method,alpha=ALPHALEVEL)) +geom_point(shape=1) +
  facet_grid(. ~ scenario) + 
  guides(alpha=FALSE) +
  geom_abline(colour = "black") +
  xlab("True pi0") +
  ylab("Estimated pi0") 
print(pi0_plot +scale_y_continuous(limits=c(0,1)) +
        scale_x_continuous(limits=c(0,1)) +
        #scale_colour_manual(values=cbbPalette,breaks=breaks,labels=labels) +
        coord_equal(ratio=1) + colScale + theme(legend.position = "top",axis.text.x = element_text(size = 8,angle=45)))

```

## Session information

```{r info}
sessionInfo()
```