### 2015-11-22 analysis/checkIP.Rmd
add checkIP analysis

---
title: "Check interior point method"
author: "Matthew Stephens"
date: 2015-11-18
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

We want to compare the likelihoods achieved by the newly implemented interior point method with the EM algorithm.

## Log-likelihood comparison

The 'dsc-opt' simply runs both methods (IP and EM) on the same data sets, 
and compares the difference in the log-likelihoods. We can see that
in >90% of cases the difference is negligible. Then there is a tail of
cases where the IP loglik is higher - presumably where the EM doesn't really converge so well.

```{r}
  load("../output/dsc-opt-files/dsc_opt.RData")
  plot(ecdf(dsc_opt$res$diff1),main="ECDF of log-likelihood difference (IP-EM)")
```

## lfsr comparison

However, what we really care about is inference quantities, such as the lfsr, rather than the log-likelihood. Here we compare the lfsr for the simulations where
we ran both the IP method and EM method (nocxx) in dsc-shrink.

```{r}
  load("../output/dsc-shrink-files/res.RData")
  library("ggplot2")
  library("dplyr")
  library("reshape2")
  # select out the lfsr estimates for each method

  df1 = res$lfsr %>% dplyr::filter(method %in% c("ash.hu.nocxx")) %>% select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
    melt(id.vars=c("method","scenario","seed",".id"),value.name="lfsr.nocxx")
  df2 = res$lfsr %>% dplyr::filter(method %in% c("ash.hu")) %>% select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
    melt(id.vars=c("method","scenario","seed",".id"),value.name="lfsr.IP")

  df = inner_join(df1,df2,by=c("scenario","seed","variable"))
  ggplot(df,aes(lfsr.nocxx,lfsr.IP)) +geom_point(shape=1) +
  facet_wrap(~ scenario, nrow=2) + 
  geom_abline(colour = "black") +
  xlab("EM algorithm (no cxx)") +
  ylab("IP method")
```

## Session information

```{r info}
sessionInfo()
```