### 2016-05-31 analysis/ash_npmle.Rmd
add example of npmle

---
title: "Using ash for NPMLE"
author: "Matthew Stephens"
date: 2016-05-31
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

## The NPMLE

Here we show how we can use ash to (approximately) compute the unconstrained
NPMLE - that is, estimate the underlying distribution $g$ by maximizing the likelihood without the unimodal constraint.
See Koenker and Mizera, JASA 2014, for background.

Within ash one can approximate the npmle as a mixture of uniforms
on a dense grid of non-overlapping values - this results in a piecewise constant density with changes in the density only at the grid points.

The following example comes from the REBayes vignette by Koenker and Gu.
The underlying $g$ is a mixture of a point mass (weight 0.8) at 0 and a point mass (weight 0.2) at 2.
```{r}
  set.seed(102)
  y <- c(rep(0,800), rnorm(200, 2)) + rnorm(1000)
  z <- REBayes::GLmix(y)
```

Now we fit the NPMLE using 'ashr' and compare with the REBayes solution.
```{r}
  grid = seq(from=min(z$x),to=max(z$x),length=1000)
  k=length(grid)
  y.ash.npmle = ashr::ash(y,1,g=ashr::unimix(pi=rep(1/(k-1),(k-1)),a=grid[-k],b=grid[-1]),method="shrink")

  plot(z$x,cumsum(z$y)/sum(z$y),col=2,main="Estimated cdf",type="l")
  lines(ashr::cdf.ash(y.ash.npmle,x=z$x),type="l")
  z$logLik
  y.ash.npmle$loglik
```

## Session information

```{r info}
sessionInfo()
```