### 2016-06-06 analysis/make_GOODPOOR_figs.Rmd
update

---
title: "Make figures for simple simulation with good and poor precision observations"
author: "Matthew Stephens"
date: 2015-10-26
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

First perform a simple simulated example to illustrate high and low signal.
True values of beta simulated from $(0.5 N(0,2^2) + 0.5 \delta_0)$.
```{r}
ntest = 10000
set.seed(112)
require(ashr)
require(qvalue)
require(locfdr)
require(mixfdr)
source("../R/plot_FDReg_hist.R")

null_alt = rbinom(ntest,1,0.5)
beta = rnorm(ntest,0,sd=1)
beta = ifelse(null_alt==1,beta,0)
GOOD= 1:(ntest/2)
sebetahat = rep(1,ntest)
sebetahat[-GOOD] = 10

betahat = beta + rnorm(ntest, 0, sebetahat)
zscore = betahat/sebetahat
pval = pchisq(zscore^2,df=1,lower.tail=F)
```

Show how poor precision observations dilute good ones:
```{r, GOODPOOReg_hist, dev='pdf',fig.width=6.5, fig.height=3}
par(mai=c(0.3,0.3,0.2,0.2),mgp = c(3, 0.5, 0))
layout(matrix(1:3,ncol=3,byrow=TRUE))
plot_FDReg_hist(pval[GOOD],1,type=1,title="Good-precision observations",ylab="",nc=20,cex.axis=1,cex.main=1.2,ylim=c(0,2.5))
plot_FDReg_hist(pval[-GOOD],1,type=1,title="Poor-precision observations",ylab="",nc=20,yaxt='n',cex.axis=1,cex.main=1.2,ylim=c(0,2.5))
axis(side=2, labels=FALSE,tck=-0.01)
plot_FDReg_hist(pval,1,type=1,title="Combined",yaxt='n',ylab="",nc=20,cex.axis=1,cex.main=1.2,ylim=c(0,2.5))
axis(side=2, labels=FALSE,tck=-0.01)
```

Apply different methods
```{r}
res.qvalue = qvalue(pval)
res.locfdr = locfdr(zscore,nulltype=0,plot=0)
res.ash = ash(betahat,sebetahat,method="fdr",outputlevel=4)

res.qvalue.good = qvalue(pval[GOOD])
res.locfdr.good = locfdr(zscore[GOOD],nulltype=0,plot=0)
res.ash.good = ash(betahat[GOOD],sebetahat[GOOD],method="fdr")
```

```{r GOODPOOReg_scatter, dev='pdf', fig.width=6.5, fig.height=3}
res = rbind(data.frame(x= res.qvalue.good$qvalues, y = res.qvalue$qvalues[GOOD],type="qvalue"),
            data.frame(x=res.locfdr.good$fdr,y=res.locfdr$fdr[GOOD],type='locfdr'), 
            data.frame(x=res.ash.good$lfsr,y=res.ash$lfsr[GOOD],type="ashr"))

library("ggplot2")
pp= ggplot(data=res,aes(x,y)) +geom_point(shape=1) +
  facet_grid(. ~ type) +
  geom_abline(colour = "red") +
  xlab("Analysing good-precision data only") +
  ylab("Analysing combined data")

print(pp +scale_y_continuous(limits=c(0,1)) +
        scale_x_continuous(limits=c(0,1))  +
        coord_equal(ratio=1))

```

```{r lfsr_vs_pval_GOODPOOR, dev='pdf', fig.width=6.5, fig.height=3}
make_df_for_pval=function(ash,method="default"){
  df = data.frame(p=pnorm(-abs(ash$data$betahat/ash$data$sebetahat)),lfsr=ash$lfsr,s=ash$data$sebetahat,method=method)
}

plot_pval_vs_lfsr=function(df,plot.it=TRUE){
  if(length(unique(df$s))>2){df$s = log(df$s)} else{ df$s=as.factor(df$s)}

  p=ggplot(df,aes(x=p,y=lfsr,color=s)) + geom_point() +facet_grid(. ~ method) +xlim(c(0, 0.025)) + xlab("p value") + ylab("lfsr")

  if(length(unique(df$s))>2){
    p = p+scale_colour_gradient2(midpoint = 1, low="blue", mid="white", high="red", space="Lab")
  } 
  if(plot.it){print(p)}
  return(p)
}

res.ash.ET = ash(betahat,sebetahat,method="fdr",model="ET",mixcompdist="normal",outputlevel=4)
p=plot_pval_vs_lfsr(rbind(
  make_df_for_pval(res.ash,method="Default prior (alpha=0)"),
  make_df_for_pval(res.ash.ET,method="p-value prior (alpha=1)")))
print(p+ theme(axis.text.x = element_text(size = 8,angle=45)) + 
        scale_colour_manual(values=c("blue","red")))

```

```{r  lfsr_vs_pval_GOODPOOR_single, dev='pdf', fig.width=4.5, fig.height=3}
#now plot one at a time
p=plot_pval_vs_lfsr(make_df_for_pval(res.ash,method="ash"))
print(p+ theme(axis.text.x = element_text(size = 8,angle=45)))

p=plot_pval_vs_lfsr(make_df_for_pval(res.ash.ET,method="ash (p-value prior)"))
print(p+ theme(axis.text.x = element_text(size = 8,angle=45)))
```

Compare the log-likelihoods:
```{r}
res.ash.ET$logLR
res.ash$logLR
```

Plot number of true positives against false positives
```{r tp-fp, dev='pdf', fig.width=4.5, fig.height=3}
  df0 = cbind(make_df_for_pval(res.ash,method="Default prior (alpha=0)"),beta=beta)
  df1 = cbind(make_df_for_pval(res.ash.ET,method="p-value prior (alpha=1)"),beta=beta)
  #false positives and true positives
  fp = function(df){cumsum((df$beta==0)[order(df$lfsr,decreasing = FALSE)])}
  tp = function(df){cumsum((df$beta!=0)[order(df$lfsr,decreasing = FALSE)])}
  df.fptp = rbind(data.frame(method="ash lfsr", tp = tp(df0), fp = fp(df0)),
             data.frame(method="p values", tp = tp(df1), fp = fp(df1)))

  ggplot(df.fptp,aes(x=fp,y=tp,col=method)) + geom_line()  + xlab("False positives") + ylab("True positives") + xlim(c(0,100)) + ylim(c(0,400)) + scale_colour_manual(values=c("black","green")) 
```

## Session information

```{r info}
sessionInfo()
```