### 2016-06-06 paper/main.tex

% Template for PLoS
% Version 1.0 January 2009
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template

\documentclass[11pt]{article}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
%\usepackage{natbib}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{xspace}
% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

%\usepackage{endfloat}
\usepackage{hyperref}

\usepackage{sectsty}
\subsectionfont{\normalsize}
\subsubsectionfont{\mdseries\itshape}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}
\usepackage{caption}
\usepackage{subcaption}

\usepackage{color}

% Use doublespacing - comment out for single spacing
\usepackage{setspace} 
\doublespacing

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm 
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

% Use the PLoS provided bibtex style
\bibliographystyle{/Users/stephens/Dropbox/Documents/stylefiles/plos2009}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

\pagestyle{myheadings}
%% ** EDIT HERE **
\usepackage{enumerate}
\usepackage{multirow} 
\usepackage{url}
\usepackage{xr} %for cross-referencing
%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW
\newtheorem{algorithm}{Algorithm}
\newtheorem{proposition}{Proposition}
\newtheorem{restateproposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{result}{Result}
\newtheorem{note}{Note}
\newtheorem{definition}{Definition}

\def\lfdr{\textit{lfdr}}
\def\lfsr{\textit{lfsr}}
\def\x{\mbf{x}}
\def\y{\mbf{y}}
\def\e{\mbf{e}}
\def\g{\mbf{g}}
\def\p{p}
\def\Yb{\hat{Y}_1^{\text{BAYES}}}
\def\Xperp{X_{1 \perp 0}}
\def\Pperp{P_{1 \perp 0}}
\def\Pperpb{\Pperp^{\text{B}}}
\def \hatd { \widehat{ d } }
\def \dhatin {\hatd^{\mbox{\tiny IN}}}
\def \dhatsw {\hatd^{\mbox{\tiny SW}}}
\def \mvn{\text{MVN}}
\def\mn{\text{MN}}
\def \iw{\text{W}^{-1}}
\def \fp {fastPHASE\ }
\def\bhat{\hat{\beta}}
\def\shat{\hat{s}}
\def\var{\rm Var}
\def\mvr{\text{BMVR}}
\def\bfuni{\text{BF}_\text{uni}}
\def\one{\mbf{1}}
\def\zero{\mbf{0}}
\def\hall{H_{\text{all}}}
\def\pall{\p_{\text{all}}}
\def\bfall{\text{BF}_\text{all}}
\def\bfav{\text{BF}_\text{av}}
\def\gall{\gamma_\text{all}}
\def\limbfall{\bfall^\rightarrow}
\def\limbfav{\bfav^\rightarrow}
\def\limbfgamma{\BF_\gamma^\rightarrow}
\def\sxx{S_{xx}}
\def\syy{S_{yy}}
\def\vyy{V_{yy}}
\def\vyx{V_{yx}}
\def\vxx{{V_{xx}}}
\def\rss{\text{\rm RSS}}
\def\dss{\text{\rm $\Delta$SS}}
\def\U{U}
\def\D{D}
\def\I{I}
\def\luu{L_{UU}}
\def\indep{\perp}
\def\qvalue{{\tt qvalue}\xspace}
\def\locfdr{{\tt locfdr}\xspace}
\def\mixfdr{{\tt mixfdr}\xspace}
\def\ash{{\tt ash}\xspace}
\def\ashr{{\tt ashr}\xspace}

\newcommand{\mbf}[1]{\mbox{\boldmath$#1$}} % true mathbf works in math
                                % mode too
\newcommand\height{{\it height}}
\newcommand\weight{{\it weight}}
\def\BF{\text{BF}}

%% END MACROS SECTION
%\externaldocument{plos_one_rev1_suppinfo}

\begin{document}

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{False Discovery Rates: A New Deal}
}
% Insert Author names, affiliations and corresponding author email.
\\
Matthew Stephens$^{1*}$, 
\\
\bf{1} Department of Statistics and Department of Human Genetics, University of Chicago, Chicago, IL, USA
\\
$\ast$ E-mail: Corresponding mstephens@uchicago.edu
\end{flushleft}

% Please keep the abstract between 250 and 300 words
\section*{Abstract}

We introduce a new Empirical Bayes approach for large-scale hypothesis testing, including  
estimating False Discovery Rates (FDRs), and effect sizes. This approach has two key differences
from existing approaches to FDR analysis. First, it
assumes that the distribution of the actual (unobserved) effects is unimodal, with a mode at 0.
This ``unimodal assumption" (UA), although natural in many contexts, is not usually incorporated into 
standard FDR analysis,
and we demonstrate how incorporating it brings many benefits.  Specifically, the UA
facilitates efficient and robust computation -- estimating the unimodal distribution involves solving a simple convex optimization problem --
and enables more accurate inferences provided that it holds.
Second, the method takes as its input two numbers for each test (an effect size estimate, and corresponding standard error),
rather than the one number usually used ($p$ value, or $z$ score). When available, using two numbers instead of one 
helps account for variation in measurement precision across tests. It also facilitates estimation of  
effects, and unlike standard FDR methods
our approach provides interval estimates (credible regions) for each effect in addition to measures of significance.
To provide a bridge between interval estimates and significance measures we introduce the term ``local false sign rate"
to refer to the probability of getting the sign of an effect wrong, and argue that it is a superior
measure of significance than the local FDR because it is both more generally applicable, and
can be more robustly estimated. Our methods are implemented in an R package {\tt ashr}
available from \url{http://github.com/stephens999/ashr}.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
%\section*{Author Summary}

\section*{Introduction}

Since its introduction in 1995 by Benjamini and Hochberg 
\cite{benjamini1995controlling}, the ``False Discovery Rate" (FDR) has quickly established itself
as a key concept in modern statistics, and the primary tool by which most practitioners handle large-scale multiple testing
in which the goal is to identify the non-zero ``effects" among a large number of imprecisely-measured effects.

Here we consider an Empirical Bayes (EB) approach to FDR. This idea is, of course, far from new:
indeed, the notion that EB approaches could be helpful in handling multiple comparisons predates introduction of
the FDR  (e.g.~\cite{greenland1991empirical}). More recently, EB approaches to the FDR
have been extensively studied by several authors, 
especially B.~Efron and co-authors 
\cite{efron2001empirical, efron2002empirical,efron2003robbins,efron2008microarrays,efron2010large}; 
see also \cite{kendziorski2003parametric,newton2004detecting, 
datta2005empirical,muralidharan2010empirical} for example.

So what is the ``New Deal" here? We introduce two simple ideas that are new (at least compared with
existing widely-used FDR pipelines) and can substantially affect
inference. The first idea is to {\it assume that the distribution of effects is unimodal}.  We provide a simple, fast, and stable computer 
implementation for perfoming EB inference under this assumption, and illustrate how it can improve inferences 
when the unimodal assumption is correct.  
The second idea is to use two numbers -- effect sizes, and their standard errors -- rather than just one -- $p$ values, or $z$ scores -- 
to summarize each measurement. Here we use this idea to allow variation in measurement precision to be better accounted for,
 avoiding a problem with standard pipelines that poor-precision measurements can inflate estimated FDR.
(\cite{ploner2006multidimensional} also suggest using more than one number in FDR analyses, taking a
rather different approach to the one we use here.)

In addition to these two new ideas, we highlight a third idea that is old, but which remains under-used in practice:
the idea that it may be preferable to focus on estimation rather than on testing.
In principle, Bayesian approaches can naturally unify testing and estimation into a single framework -- testing is
simply estimation with some positive prior probability that the effect is exactly zero.
However, despite ongoing interest in this area from both frequentist \cite{benjamini2005false} and Bayesian \cite{zhao2012empirical,gelman2012we} 
perspectives, in practice large-scale studies that assess many effects almost invariably focus on testing significance and
controlling the FDR, and not on estimation. To help provide a bridge between FDR and estimation we introduce the term
``local false sign rate" (\lfsr), which is analogous to the ``local false discovery rate" (\lfdr) \cite{efron2008microarrays}, but which measures confidence  
in the {\it sign} of each effect rather than confidence in each effect being non-zero. We show that in some settings, particularly those with many discoveries, 
the $\lfsr$ and $\lfdr$ can be quite different, and emphasize benefits of the $\lfsr$, particularly its increased robustness to modeling assumptions.

%terminology
%``signed discovery" for a discovery (an effect declared to be non-zero) for which the sign of the effect (positive or negative) is declared,
%and correspondingly define the ``false signed discovery rate", or simply ``false sign rate" (FSR), in a natural way.

%The methods we provide here can address both problems, and we provide results that highlight important benefits of focussing on estimation, and particularly on assessing the correctness of the {\it sign} of the estimated effects, rather than on testing for non-zero effects.

Although we focus here on FDR applications, the idea of performing EB inference using a flexible unimodal prior distribution
is useful more generally. For example, the methods described here can be applied directly to perform shrinkage estimation for
wavelet denoising \cite{donoho:1995}, an idea explored in a companion paper \cite{xing2016smoothing}. And analogous ideas can be used to perform
EB inference for variances \cite{lu2016variance}.
Importantly, and perhaps surprisingly, our work demonstrates how EB inference under a general unimodal assumption is, if anything, computationally simpler  
than commonly-used more restrictive assumptions -- such as a spike and slab or Laplace prior distribution \cite{johnstone2004needles} --  as well as being more flexible.

We refer to our EB method as {\bf a}daptive {\bf sh}rinkage, or \ash, to emphasise its key points:
using a unimodal prior naturally results in shrinkage estimation,
and the shrinkage is adaptive to both the amount of signal in the data and the measurement precision of each observation. We provide implementations in an R package, \ashr, available at 
\url{http://github.com/stephens999/ashr}.
Code and instructions for reproducing analyses and figures in this paper are at
\url{https://github.com/stephenslab/ash}.

%Measurement precision will often vary among genes --  for example, log-fold change estimate for highly expressed genes may have greater precision than those for lower-expressed genes -- and one goal  of our methods is to allow for thiis variation.

%Let $\mu^0_j$ and $\mu^1_j$ denote the mean expression 
%of gene $j$ ($j=1,\dots,J$) in the two conditions, and define $\beta_j:= \mu^0_j - \mu^1_j$ to be the difference. Typically expression
%measurements are made on only a small number of
%samples in each condition - sometimes as few as one
%sample in each condition. Thus the error in estimates
%of $\mu^0_j$ and $\mu^1_j$ is appreciable, and the error
%in estimates of $\beta_j$ still greater.
%

%Many tools have been developed for estimating or controlling the FDR and related quantities \cite{}, and some are widely used.
%Here we present new tools that bring two important ideas to bear on the problem.
%First, these tools are designed to take account of variations in the {\it measurement precision} of different effects, reducing the influence of low-precision measurements.
%For example, in the example of detecting differentially expressed genes, measurements of log fold change for high-expressed genes 
%may have better precision than those for low-expressed genes; by incorporating this information our methods can avoid the tendancy for 
%imprecise measurements at low-expressed genes to dilute the signal at higher-expressed genes.
%Second, our tools explicitly assume that the non-zero effects have a unimodal distribution centered at zero. This assumption 
%is, we argue, often realistic, and comes with considerable benefits, both statistical and computational. 
%Our tools are implemented in an R package \ash (short for ``Adaptive Shrinkage") and we 
%demonstrate how the incorporation of these two ideas can lead to a substantial reduction in estimated FDRs (or, equivalently,
%a substantial increase in the number of discoveries at a given FDR).

%Although we motivate our tools by the need to address the FDR problem, our \ash package, as its name suggests,  provides a general tool for shrinkage estimation. 
%As an important special case, these methods address the "multiple comparisons" setting, where interest usually focuses on which $\beta_j$ can be confidently inferred to be non-zero. Such problems are usually tackled by computing a $p$ value for each $j$, often by applying a $t$ test to $\bhat_j/s_j$,
%and then applying a generic procedure, such as that of \cite{benjamini1995controlling} or \cite{storey2003statistical}, designed to control or
%estimate the false discovery rate (FDR) or the positive FDR \cite{storey.03}. In essence we aim to provide analagous
%generic methods that work directly with two numbers for each 
%measurement $(\bhat_j,s_j$), rather than a single number (e.g.~ the $p$ value, or $t$ statistic). Working with these two numbers has two important benefits: first, it permits estimation and not only testing; second, the 
%uncertainty in each measurement $\bhat_j$ can be more fully accounted for, reducing the impact of ``high-noise" measurements (large $s_j$) that can reduce the effectiveness of a standard FDR analysis.

\def\df{df}
\def\FDR{\text{FDR}}
\def\fdr{\text{\lfdr}}
\def\FSR{\text{FSR}}
\def\fsr{\text{lfsr}}

 \section*{Methods}

 \subsection*{Model Outine}

% Here we consider one setting in which FDR-based methods are widely used:
%an experimenter has measured a large numbers of ``effects", each with an associated
%measurement precision (standard error), 
%and the goal is to identify which of the effects we can declare to be non-zero, with some prescribed level of confidence.
%Although the usual focus of FDR-based methods is on distinguishing zero and non-zero effects,
%here we also focus on providing estimates -- both point estimates and interval estimates -- for these effects.
%This setting arises commonly in many areas of statistics, and particularly in genomics applications.
%For example, in genomics
%a common goal is to identify ``differentially-expressed genes" -- that is, genes
%whose mean expression (activity) level differs in two conditions. Here the ``effect" of interest is the difference in (log) 
%mean expression level between the two conditions, often referred to as the ``log-fold change". 
Here we describe the simplest version of the method, and briefly discuss embellishments we have also implemented.

Let $\beta=(\beta_1,\dots,\beta_J)$ denote $J$ ``effects" of interest. For example, in a genomics application 
$\beta_j$ might be the difference in the mean (log) expression of gene $j$ in two conditions.
In contexts where FDR methods are applied, interest often focuses on identifying ``significant" non-zero effects;
that is, testing the null hypotheses $H_j:\beta_j=0$.  Here we tackle both this problem, and the  
more general problem of estimating, and assessing uncertainty in, $\beta_j$.

Assume that the available data are estimates $\bhat=(\bhat_1,\dots,\bhat_J)$ of the effects,
and corresponding (estimated) standard errors $\shat=(\shat_1,\dots,\shat_J)$.  Our goal is to compute a posterior distribution for $\beta$ given the observed data $\bhat,\shat$,
which by Bayes theorem can be written as 
\begin{equation}
p(\beta | \bhat, \shat) \propto p(\beta | \shat) p(\bhat | \beta, \shat).
\end{equation}

For $p(\beta | \shat)$ we assume that the $\beta_j$ are independent from a unimodal distribution $g$.
This unimodal assumption (UA) is a key assumption that distinguishes our approach from previous EB approaches to FDR analysis.
A simple way to implement the UA is to assume that $g$ is
a mixture of a point mass at 0 and a mixture of {\it zero-mean} normal distributions:
\begin{align} \label{eqn:beta}
p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\   
g(\cdot; \pi) & = \pi_0 \delta_0(\cdot) + \sum_{k=1}^K \pi_k N(\cdot; 0, \sigma_k^2), \label{eqn:mixnorm}
\end{align}
where $\delta_0(\cdot)$ denotes a point mass on 0, and $N(\cdot; \mu, \sigma^2)$ denotes the density of the normal distribution with mean $\mu$ and variance $\sigma^2$.
Here the mixture proportions $\pi=(\pi_0,\dots,\pi_K)$ are hyper-parameters, which are non-negative and sum to one, and are to be estimated,
while the mixture component standard deviations $\sigma_1,\dots,\sigma_K$ 
represent a large and dense grid of {\it fixed} positive numbers spanning a range from very small to very big (so $K$ is fixed and large). 
We encourage the reader to think of this grid as becoming infinitely large and dense, as a non-parametric limit,
although of course in practice we use a finite grid -- see Implementation Details.

For the likelihood $p(\bhat | \beta, \shat)$ we assume a normal approximation:
\begin{equation} \label{eqn:normlik}
p(\bhat | \beta, \shat) = \prod_j N(\bhat_j; \beta_j, \shat_j^2).
\end{equation}
%Here, in addition to some conditional independence assumptions, we are 
%effectively  assuming that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.

This simple model features both the key ideas we want to emphasize in this paper: the UA is encapsulated in (\ref{eqn:mixnorm}) while
the different measurement precision of different observations is encapsulated in the likelihood (\ref{eqn:normlik}) -- specifically, observations
with larger standard error will have a flatter likelihood, and therefore have less impact on inference.
However, the model also has several additional assumptions that can be relaxed. Specifically,
\begin{enumerate}
\item  The use of a mixture of zero-mean normals (\ref{eqn:mixnorm}) implies that $g$ is symmetric about 0.
More flexibility can be obtained using mixtures of uniforms [see equation (\ref{eqn:g-general})]; indeed this 
can allow $g$ to approximate {\it any} unimodal distribution. 
\item The model (\ref{eqn:beta}) assumes that the effects are identically distributed, independent of their standard errors $\shat$.
This can be relaxed [see (\ref{eqn:beta-alpha})].
\item The normal likelihood (\ref{eqn:normlik}) can be generalized to a $t$ likelihood [see (\ref{eqn:tlik})].
\end{enumerate}
These embellishments, detailed in Detailed Methods, are all implemented in our software.
Of course there remain limitations that are harder to relax, most notably the independence and conditional independence assumptions
(which are also made by most existing approaches). Correlations among tests certainly arise in practice, either due to genuine correlations 
or due to unmeasured confounders, and their potential impact on estimated FDRs is important to consider
whatever analysis methods are used \cite{efron2007correlation,leek:2007}.

%We assume that the observations $\bhat$ are conditionally independent given $\beta,\shat$, and that the distribution of
%$\bhat_j$ depends only on the corresponding $\beta_j,\shat_j$, so
%\begin{equation}
%p(\bhat | \beta, \shat) = \prod_j p(\bhat_j | \beta_j, \shat_j).
%\end{equation}
%Initially we assume that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation,
%\begin{equation} \label{eqn:normlik}
%p(\bhat_j | \beta_j, \shat_j) =  N(\bhat_j; \beta_j, \shat^2_j),
%\end{equation}
%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and variance $\sigma^2$.
%
%We assume that the (unobserved) effects $\beta_j$ come from some distribution $g$, which is to be estimated, and which we write 
%as a mixture of null and alternative components:
%\begin{equation} \label{eqn:betaj0}
%g(\cdot) = \pi_0 \delta_0(\cdot) + (1-\pi_0) g_1(\cdot)
%\end{equation}
%where $\pi_0$ denotes the proportion of effects that are null, $\delta_0(\cdot)$ denotes a point mass at 0, and $g_1(\cdot)$ denotes the distribution of the non-zero $\beta_j$.
%We also assume that (given $g$) the $\beta_j$ are independent, and independent of $\shat$ (a point we return to later), so
%\begin{equation} \label{eqn:beta-iid}
%p(\beta | \shat) = \prod_j g(\beta_j).
%\end{equation}
%
%Our key further assumption, which distinguishes our approach from previous EB approaches to FDR analysis, is that $g_1$ is unimodal,
%with a mode at 0. One simple way to implement this ``Unimodal Assumption" (UA)  is
%to use a mixture of zero-mean normal distributions:
%\begin{equation} \label{eqn:mixnorm}
%g_1(\cdot) = \sum_{k=1}^K w_k N(\cdot; 0, \sigma_k^2)
%\end{equation}
%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and standard deviation $\sigma$.
%Here the $w_k$ are non-negative mixture weights that sum to $1$ and are to be estimated, and the component standard deviations $\sigma_1,\dots,\sigma_K$ 
%represent a large and dense grid of fixed positive numbers spanning a range from very small to very big (so $K$ is fixed and large).
% The symmetry of the normal distribution implies that (\ref{eqn:mixnorm}) is not only unimodal, but also symmetric about zero. We introduce
% more flexible alternatives based on using mixtures of uniform distributions, and further details (such as choice of gridpoints $\{\sigma_k:k=1,\dots,K\}$) below. 
% 
% Substituting (\ref{eqn:mixnorm}) for $g_1$ in (\ref{eqn:betaj0}), and defining $\pi_k := (1-\pi_0)w_k$ for $k=1,\dots,K$ (and also $\sigma_k:=0$, and $N(\cdot;0,0):=\delta_0(\cdot)$)
% yields a notationally-convenient form for $g$, with hyperparameters $\pi:=(\pi_0,\pi_1,\dots,\pi_K)$ whose elements are non-negative and sum to 1:
% \begin{equation} \label{eqn:g}
% g(\cdot; \pi) = \sum_{k=0}^K \pi_k N(\cdot; 0, \sigma_k^2).
% \end{equation}
% 
% Putting all this together, and making the dependence on hyperparameters $\pi$ explicit, we have:
% \begin{align} \label{eqn:betaj}
% p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\  \label{eqn:normlik2}
% p(\bhat | \beta, \shat, \pi) &= \prod_j N(\bhat_j; \beta_j, \shat_j^2), 
% \end{align}
% with $g$ given by the mixture of normals $(\ref{eqn:g})$.

\subsection*{Fitting the model}

Together, (\ref{eqn:beta})-(\ref{eqn:normlik}) imply that $\bhat_j$ are independent observations from a mixture of normal distributions:
\begin{equation} \label{eqn:marginallik}
p(\bhat | \shat, \pi) =   \prod_j  \left[\sum_{k=0}^K \pi_k N(\bhat_j; 0, \sigma_k^2 + \shat_j^2) \right],
\end{equation}
where we define $\sigma_0:=0$.

The usual EB approach to fitting this model would involve two simple steps:
\begin{enumerate}
\item Estimate the hyper-parameters $\pi$ by maximizing the likelihood $L(\pi)$, given by (\ref{eqn:marginallik}), yielding $\hat{\pi} := \arg \max L(\pi)$.
\item Compute quantities of interest from the conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$. For example, the evidence
against the null hypothesis $\beta_j=0$ can be summarized by $p(\beta_j \neq 0 | \bhat, \shat,\hat{\pi})$.
\end{enumerate}
Both steps 1 and 2 are very straightforward. Step 1 is a convex optimization problem, and can be solved quickly and reliably using interior point methods \cite{boyd2004convex,koenker2013convex}. (Alternatively a simple EM algorithm can also work well, particularly for modest $J$, say $J<2000$;
see \url{http:://stephenslab.github.io/ash/analysis/IPvsEM.html}.) 
And the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ in Step 2 are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
(The simplicity of step 1 is due to our use of a fixed grid for $\sigma_k$ in (\ref{eqn:mixnorm}), instead of estimating $\sigma_k$,
which may seem more natural but is not straightforward when $\shat_j$ varies among $j$. This simple device may be useful in other applications.)

Here we slightly modify this usual procedure: instead of obtaining $\hat\pi$ by maximizing the likelihood, we maximize a penalized
likelihood [see (\ref{eqn:penalty})], where the penalty encourages $\hat\pi_0$ to be as big as possible whilst remaining consistent with the observed data. 
We introduce this penalty because in FDR applications it is considered desirable to avoid underestimating $\pi_0$ so as to avoid underestimating the FDR.

Our {\tt R} package implementation typically takes $~20s$ on a modern laptop for $J=100,000$, and scales linearly with $J$.

\subsection*{The local False Discovery Rate and local False Sign Rate}

As noted above, the posterior distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ have a simple analytic form.
In practice it is common, and desirable, to summarize these distributions to convey the ``significance" of each observation $j$.
One natural measure of the significance of observation $j$ is its ``local FDR" \cite{efron2008microarrays}, which is
the probability, given the observed data, that effect $j$ would be a false discovery, if we were to declare it a discovery.
In other words it is the posterior probability that $\beta_j$ is actually zero:
\begin{equation} \label{eqn:fdr}
\fdr_j := \Pr(\beta_j =0  |  \bhat, \shat, \hat{\pi}).
\end{equation}

The $\lfdr$, like most other measures of significance, is rooted in the hypothesis testing paradigm which focuses on 
whether or not an effect is exactly zero. This paradigm is popular, despite the fact that many statistical practitioners have argued that it is often inappropriate because
the null hypothesis $H_j: \beta_j=0$ is often implausible. For example, Tukey \cite{tukey1991philosophy} argued that
``All we know about the world teaches us that the effects of $A$ and $B$ are always different -- in some decimal place -- for any $A$ and $B$. Thus asking `Are the effects different?' is foolish." Instead, Tukey (\cite{tukey1962future}, p32) suggested  that one should address
\begin{quote}
...the more meaningful question: ``is the evidence strong enough to support a belief that the observed difference has the correct sign?"
\end{quote}
Along the same lines, Gelman and co-authors \cite{gelman2000type, gelman2012we} suggest  
focussing on ``type S errors", meaning errors in sign, rather than the more traditional type I errors.

Motivated by these suggestions, we define the ``local False Sign Rate" for effect $j$, $\lfsr_j$, to be the probability that we
would make an error in the sign of effect $j$ if we were forced to declare it either positive or negative. Specifically,
\begin{equation} \label{eqn:lfsr}
\lfsr_j := \min[ \Pr(\beta_j \geq 0| \hat\pi, \bhat, s), \Pr(\beta_j \leq 0| \hat\pi, \bhat, s) ].
\end{equation}
To illustrate, suppose that 
\begin{gather*}
\Pr(\beta_j < 0| \bhat, s, \hat\pi)=0.95, \\
\Pr(\beta_j =0| \bhat, s, \hat\pi) = 0.03, \\
\Pr(\beta_j >0| \bhat, s, \hat\pi) = 0.02.
\end{gather*} 
Then 
from (\ref{eqn:lfsr}) $\lfsr_j=\min(0.05,0.98)=0.05$ (and, from (\ref{eqn:fdr}), $\lfdr_j=0.03$). This $\lfsr$ corresponds to the fact that, given these results, 
we would guess that $\beta_j$ is negative, with probability $0.05$ of being wrong.

As our notation suggests, $\lfsr_j$ is analogous to $\lfdr_j$: whereas small values of $\lfdr_j$ indicate that we can be {\it confident that $\beta_j$ is non-zero}, 
small values of $\lfsr_j$ indicate that we can be {\it confident in the sign of $\beta_j$}. 
Of course, being confident in the sign of an effect logically implies that we are confident it is non-zero, and
this is reflected in the fact that $\lfsr_j \geq \lfdr_j$ 
(this follows from the definition because both the events $\beta_j \geq 0$
and $\beta_j \leq 0$ in (\ref{eqn:lfsr}) include the event $\beta_j=0$).
In this sense, as a measure of ``significance", $\lfsr$ is more conservative than $\lfdr$. More importantly, as we illustrate in Results,
$\lfsr$ is more robust to modeling assumptions.

From these ``local" measures of significance, we can also compute average error rates over subsets of observations
$\Gamma \subset \{1,\dots,J\}$. For example, 
\begin{equation} \label{eqn:FSRhat}
\widehat\FSR(\Gamma):= (1/|\Gamma|) \sum_{j \in \Gamma} \lfsr_j 
\end{equation}
is an estimate of the total proportion of errors made if we were to estimate the sign of all effects in $\Gamma$
(an error measure analogous to the usual (tail) FDR).
And, we can define the ``$s$-value" 
\begin{equation} \label{eqn:qvalue}
s_j := \widehat \FSR(\{k: \lfsr_k \leq \lfsr_j\}) 
\end{equation}
as a measure of significance analogous to Storey's $q$-value \cite{storey.03}.
(Replacing $\lfsr_j$ with $\lfdr_j$ in these definitions gives an estimate for the usual $\FDR(\Gamma)$, and an estimated $q$ value for each observation respectively.)

%Because our methods involve explicitly modelling effects $\beta_j$, rather than 
%$p$ values or $z$ scores, they can provide estimates of the local false sign rates.

%Indeed, even in settings where some
%$\beta_j$ are exactly zero, separately identifying which are positive and which negative may be useful. For example, when identifying differentially expressed genes,
%analysts will often separate the genes that are ``upregulated" from those that ``downregulated" in a condition.

\subsection*{Related work}

\subsubsection*{Previous approaches focussed on FDR}

Among previous methods that explicitly consider the FDR and related quantities,
our work here seems most naturally compared with the EB methods of \cite{efron2008microarrays} and \cite{muralidharan2010empirical}
 (implemented in the R packages \locfdr and \mixfdr respectively) and with the widely-used methods from  \cite{storey.03} (implemented in the R package \qvalue), which although not formally an EB approach, shares some elements in common.
  %Here we highlight the key differences between our approach and these three existing approaches.

There are two key differences between our approach and all of these three existing methods. First, 
whereas these existing methods summarize the information on $\beta_j$ by a single number -- either a $z$ score (\locfdr and \mixfdr), or a $p$ value (\qvalue) -- 
we instead work with two numbers ($\bhat_j,\shat_j$). Here we are building on \cite{wakefield:2009}, who develops Bayesian tests 
for individual null hypotheses in a similar way; see also \cite{efron1993bayes}.
Using two numbers instead of one clearly has the potential to be more informative, and this is illustrated in Results later (Figure \ref{fig:goodpoor}).

%(And in the special case where $\shat_j$ are equal for all $j$, our approach essentially comes down to the same thing as these existing methods.)
% Or more generally Section xx illustrates the potential practical effects

%allows for different modelling assumptions to be compared through likelihood comparisons.
%\def\sZA{ZA}
Second, our unimodal assumption (UA) that the effects are unimodal about zero, is an assumption 
not made by existing methods, and one that we argue to be both plausible and beneficial in many contexts.
Although the UA will not hold in all settings, it will often be reasonable, especially in FDR-related contexts
that focus on rejecting the null hypotheses $\beta_j=0$. This is because if ``$\beta_j=0$" is a plausible null hypothesis
then ``$\beta_j$ very near 0" should also be plausible. Further, it seems reasonable to expect that larger effects become
decreasingly plausible, and so the distribution of the effects will be unimodal about 0. 
To paraphrase Tukey, ``All we know about the world teaches us that large effects are rare, whereas small effects abound."
We emphasize that the UA relates to the distribution of {\it all} effects, and not only the {\it detectable} effects (i.e.~those that are significantly different from zero). It is very likely that the distribution of {\it detectable} non-zero effects will be multimodal, with one mode for detectable positive effects and another for detectable negative effects, and the UA does not contradict this.

% Indeed, neither $\locfdr$ nor $\qvalue$ explicitly model the effect distribution at all:  they effectively avoid this by simply subtracting a null distribution 
%(of $z$ scores or $p$ values) from the observed empirical distribution, with the remains necessarily 
%representing the alternative distribution. When doing this they assume that all the $z$ scores near zero (or, equivalently,
%all $p$ values near 1) are null, which results in a ``hole" in the alternative distribution

%Indeed, $\locfdr$ \cite{efron2008microarrays} assumes that all $z$ scores near 0 are null 
%(which we call the Zero Assumption, \sZA), which implies that 
%under the alternative hypothesis the distribution of $z$ scores has {\it no support at 0}.  
%(Although \cite{efron2008microarrays} motivates and defines the \sZA as ``most $z$ scores near 0 are null",
%results from $\locfdr$ actually obey the stronger version of the \sZA  that we use here: {\it all} $z$ scores near 0 are null.)
%Similarly, $\qvalue$ assumes -- when estimating $\pi_0$ -- that all $p$ values near 1 are null.
%which, if the $p$ values were derived from $z$ scores, is the ZA because $p$ values near 1 correspond to $z$ scores near 0.
%And although \mixfdr does not formally make the \sZA,
%we have found that in practice, with default settings, the results often approximately satisfy this assumption (due, we believe, to the default choice of penalty term $\beta$ described in \cite{muralidharan2010empirical}).
%Tukey "peri-null"

%Interestingly, although the unimodal assumption seems natural, much previous related work on estimating False Discovery Rates  
%has assumed, either explicitly or implicitly, that the distribution of effects is multimodal - see Figure \ref{fig:ZA}.

In further support of the UA, note that large-scale regression methods almost always make the UA for the regression coefficients,
which are analogous to the ``effects" we estimate here. 
Common choices of uni-modal distribution for regression coefficients 
include the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors \cite{carvalho2010horseshoe}.
These are all less flexible than our approach, which provides for general uni-modal distributions, and 
it may be fruitful to apply our methods to the regression context; indeed see \cite{moser:2015} for work in this vein. 
Additionally, the UA can be motivated by its effect on point estimates, which is to ``shrink" the estimates towards the mode - 
such shrinkage is desirable from several standpoints for improving estimation accuracy. Indeed most model-based approaches to shrinkage
make parametric assumptions that obey the UA (e.g. \cite{johnstone2004needles}).

Besides its plausibility, the UA has two important practical benefits: it facilitates more accurate estimates
of FDR-related quantities, and it yields simple algorithms that are both computationally
and statistically stable (see Results).

\subsubsection*{Other work}

There is also a very considerable literature that does not directly focus on the FDR problem, but which involves similar ideas and methods. 
 Among these, a paper about deconvolution \cite{cordy1997deconvolution} is most similar, methodologically, to our work here:
indeed, this paper includes all the elements of our approach outlined above, except for the point mass on 0 and corresponding penalty term.
This said, the focus is very different: \cite{cordy1997deconvolution} focuses entirely on estimating $g$, whereas our primary focus is on estimating $\beta_j$.
Also, they provide no software implementation.

More generally, the related literature is too large to review comprehensively, but relevant key-words include ``empirical bayes", ``shrinkage", ``deconvolution", ``semi-parametric", ``shape-constrained", and ``heteroskedastic". Some pointers to recent papers in which other relevant
citations can be found include \cite{xie2012sure, sarkar:2014, koenker2014convex}. Much of the literature focusses on the homoskedastic case (i.e. $\shat_j$ all equal) whereas we allow for heteroskedasticity. And much of the recent shrinkage-oriented literature focuses only on point estimation of $\beta_j$, whereas for FDR-related applications measures of uncertainty are essential. Several recent papers consider more flexible non-parametric assumptions on 
$g$ than the UA assumption we make here. In particular, \cite{jiang2009general,koenker2014convex} consider the unconstrained non-parametric maximum likelihood
estimate (NPMLE) for $g$. These methods may provide alternatives to our approach in settings where the UA assumption is considered too restrictive. However, the  NPMLE for $g$ is a discrete distribution, which will induce a discrete posterior distribution on $\beta_j$, and so although the NPMLE
may perform well for point estimation, it may not adequately reflect uncertainty in $\beta_j$, and some regularization on $g$ may be necessary.
Indeed, one way of thinking about the UA is as a way to regularize $g$.

%The conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ 
%encapsulate uncertainty in $\beta_j$, combining information across {\it all} observations 
%$j=1,\dots,J$. Specifically, the combining of the information occurs through estimation of
%$\pi$, which involves all of the data. Importantly, when estimating $\pi$ our approach accounts for
%the precision of each measurement through the likelihood (\ref{eqn:normlik}). For example, 
%observations with a large $\shat_j$ have an essentially flat likelihood, and so have little effect on estimation of $\pi$, and
%corresondingly little effect on the estimated FDR of other observations. 
%We highlight this because it contrasts with approaches that model the $p$ values or $z$ scores directly: observations with large $\shat_j$ will tend to produce 
%$p$ values with approximately $U[0,1]$ distribution, and $z$ scores with an approximately $N(0,1)$ distribution, under both the null and the alternative hypothesis (assuming
%that the effect $\beta_j$ is small compared with $\shat_j$). Such noisy observations therefore do affect inference; specifically they add to the
%estimated weight of the null component $\pi_0$, and inflate estimates of the FDR. 
%Because of this our approach can take better account of the informativeness of each measurement, as illustrated in Results below.

\def\dt{\text{dt}}

%
%assume that the likelihood $L(\beta) := p(D | \beta)$ can be approximated by 
%$\hat{L}(\beta) := \prod_j \hat{L}_j(\beta_j)$ where
% $\hat{L}_j(\beta_j) := p(\bhat_j | \beta_j, \shat_j)$ is given by 
%  \begin{equation} \label{eqn:lik}
% \hat{L}_j(\beta_j) \propto \dt_\nu(\bhat_j; \beta_j, \shat_j)
% \end{equation}
% where $\dt_\nu(\cdot; \mu,\sigma)$ denotes the density of the generalized $t$ distribution on $\nu$ degrees of freedom with mean $\mu$ and scale parameter $\sigma$ (i.e. the density of $\mu+\sigma T_\nu$). In the special case $\nu \rightarrow \infty$ this corresponds, of course, to a normal approximation to the likelihood:
%  \begin{equation} \label{eqn:normlik}
% \hat{L}_j(\beta_j) \propto N(\bhat_j; \beta_j, \shat_j).
% \end{equation}
%which, in light of (\ref{eqn:betahat}), is equivalent to
 % \begin{equation} \label{eqn:lik2}
% L(\beta_j) \propto p(\bhat_j | \beta_j, \shat_j).
 %\end{equation}

%The $t$ likelihood (\ref{eqn:lik}) can derived from an assumption of conditional independence of the data on each effect given the true effects, plus two additional assumptions: a) $\bhat_j, \shat_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e. they are ``approximately sufficient" for $D_j$); and b) $\shat_j$ alone contains little information about $\beta_j$. The former (a) suggests the approximation $\hat{L}_j(\beta_j) \propto p(\bhat_j, \shat_j | \beta) = p(\bhat | \beta, \shat_j) p(\shat_j | \beta)$, while the latter (b) suggests assuming that $p(\shat_j|\beta_j)$ does not depend on $\beta_j$. Putting these two together yields $\hat{L}_j(\beta_j) \propto p(\bhat_j | \beta, \shat_j)$. If $\bhat_j, \shat_j$ are not sufficient for $D_j$ then there is some loss of efficiency in making this approximation, but all Bayesian inferences remain valid provided they are interpreted as conditional on $\bhat_j,\shat_j$, rather than conditional on $D_j$. See \cite{johnson:2008} and \cite{wakefield:2009} for more general discussion and examples of Bayesian inference based on summary statistics.

%This likelihood approximation can be motivated by assuming that i) $\bhat$
% $(\bhat_j -\beta_j)/\shat_j | \shat_j$ has a $T_\nu$ distribution, and then writing
% $p(D_j | \beta_j) = p(\bhat_j, \shat_j | \beta) = $ as sufficient for $\beta$. 
%

%Alternatively (\ref{eqn:lik}) could be considered definitional for $\bhat_j,\shat_j$: that is, choose $\bhat_j,\shat_j,\nu$ such that (\ref{eqn:lik}) holds.

%While has parallels with approaches like \cite{efron2008microarrays, sun2007oracle}
 %that model the $Z$ scores $Z_j = \bhat_j/\shat_j$ as normally distributed under the null, 
%there is an important difference: here,

%The choice of normal likelihood seems natural, and indeed it can be motivated in multiple ways. For example, we can write $p(\bhat_j, \shat_j | \beta_j) = p(\bhat_j | \shat_j, \beta_j)p(\shat_j | \beta_j)$. Now, if we are willing to assume that $\shat_j$ alone contains no information about $\beta_j$, or equivalently that $p(\shat_j | \beta_j)$ does not depend on $\beta_j$, then $L(\beta_j) \propto p(\bhat_j | \shat_j, \beta_j)$, and if $\bhat_j | \shat_j, \beta_j \sim N(\beta_j, \shat_j^2)$, as is often asymptotically the case, then we obtain the likelihood (**) above.

\section*{Results}

We compare results of \ash with existing FDR-based methods implemented
in the R packages \qvalue (v2.1.1), \locfdr (v1.1-8), and \mixfdr (v1.0, from \url{https://cran.r-project.org/src/contrib/Archive/mixfdr/}).
In all our simulations we assume that the test statistics follow the expected theoretical distribution under the null, and we indicate this
to \locfdr using {\tt nulltype=0} and to \mixfdr using {\tt theonull=TRUE}. Otherwise all packages were used with default options.

\subsection*{Effects of the Unimodal Assumption}

Here we consider the effects of making the UA. 
To isolate these effects we consider the simplest case, where every observation has the same
standard error, $s_j=1$ and all methods are provided that information. That is,
$\bhat_j | \beta_j \sim N(\beta_j,1)$ and $\shat_j=s_j=1$. In this case the $z$ scores $z_j:=\bhat_j/\shat_j=\bhat_j$, so modelling the $z$ scores is the same as modelling the $\bhat_j$, and so the primary difference among methods lies in how they estimate $g$.

To briefly summarize the results in this section:
\begin{enumerate}
\item The UA can produce quite different results from existing methods.
\item The UA can yield conservative estimates of the proportion of true nulls, $\pi_0$, and hence conservative estimates of $\lfdr$ and $\FDR$.
\item The UA results in a stable procedure, both numerically and statistically, and is somewhat robust to deviations from unimodality.
\end{enumerate}

\subsubsection*{The UA can produce quite different results from existing methods.}

\begin{figure}
\center \includegraphics[height=6in]{../analysis/figure/makefig_FDReg.Rmd/decomp_ZA-1.pdf}
\caption{Illustration that the unimodal assumption (UA) in \ash can produce very different results from existing methods.
The figure shows, for a single simulated dataset, the way different methods decompose $p$ values (left) and $z$ scores (right) into a null component (dark blue) and an alternative component (cyan). In the $z$ score space the alternative distribution is placed on the bottom to highlight the differences in its shape among methods.
The three existing methods (\qvalue, \locfdr, \mixfdr) all produce a ``hole" in the alternative $z$ score distribution around 0.
In contrast the method introduced here (\ash) makes the Unimodal Assumption -- that the effect sizes, and thus the $z$ scores, have a unimodal distribution about 0 -- which yields a very different decomposition. (In this case the \ash decomposition is closer to the truth: the data were simulated under a model where all of the effects are non-zero, so the ``true" decomposition would make everything cyan.)} \label{fig:ZA}
\end{figure}

We illustrate the effects of the UA, and the difference with existing methods, using a simple simulated dataset
 with the true effects $\beta_j \sim N(0,1)$ (so with $s_j=1$, $\bhat_j  \sim N(0,2)$). Note that none of the effects are truly null,
but nonetheless there are many $p$ values near 1 and $z$ scores near 0 (Figure \ref{fig:ZA}).  We used each of the methods
\qvalue, \locfdr, \mixfdr and \ash to decompose the $z$ scores ($z_j = \bhat_j$), or their corresponding $p$ values, into null and alternative components.
Here we are using the fact that these methods all provide an estimate of the \lfdr{} for each observation, which
implies such a decomposition;  specifically the average \lfdr{} within each histogram bin estimates the fraction of observations in that bin that come from the null vs the alternative component. The results (Figure \ref{fig:ZA}) illustrate a clear difference between the existing methods and our method:
all the existing methods have a ``hole"  in the alternative $z$ score distribution near 0, whereas \ash, due to the UA, has a mode near 0.
%Although \mixfdr does not formally make the \sZA, its decomposition
%exhibits a similar hole. 
%In contrast, due to the UA, the alternative $z$ score distribution for \ash is required to have a mode at 0, 
%effectively ``filling in" the hole.  
(Of course the null distribution also has a peak at 0, and the local fdr 
 under the UA is still smallest for $z$ scores that are far from zero -- i.e. large $z$ scores remain the ``most significant".)

This qualitative difference among methods in the inferred alternative distribution of the $z$ scores is quite general, and 
also occurs in simulations where most effects are null (e.g.~\url{http://stephenslab.github.io/ash/analysis/referee_uaza.html}). To understand why the alternative distribution of $z$ scores from $\locfdr$ and $\qvalue$ has a hole at zero, note that neither
of these methods explicitly models the alternative distribution: instead they simply subtract a null distribution (of $z$ scores or $p$ values) from the observed empirical distribution, letting the alternative
distribution be defined implicitly, by what remains.
 In deciding how much null distribution to subtract -- that is, in estimating the null proportion, $\pi_0$ -- 
 both methods assume that all $z$ scores near zero (or, equivalently,
all $p$ values near 1) are null. The consequence of this is that their (implicitly-defined) distribution for the alternative $z$ scores has 
a ``hole" at 0 -- quite different from our assumption of a mode at zero. (Why \mixfdr exhibits similar behaviour is less clear, since it does explicitly
model the alternative distribution; however we believe it may be due to the default choice of penalty term $\beta$ described in \cite{muralidharan2010empirical}.)

Figure \ref{fig:ZA} may also be helpful in understanding the interacting role of the UA and the penalty term (\ref{eqn:penalty}) that attempts to make $\pi_0$ as ``large as possible" while remaining consistent with the UA. Specifically, consider the panel of Figure \ref{fig:ZA} that shows \ash's decomposition of $z$ scores, and imagine increasing $\pi_0$ further. This would increase the null component (dark blue) at the expense of
the alternative component (light blue). Because the null component is $N(0,1)$, and so is biggest at 0, this would eventually create a ``dip" in the light-blue histogram at 0. The role of the penalty term is to push the dark blue component as far as possible, right up to (or, to be conservative, just past) the point where this dip appears. In contrast the 
existing methods effectively push the dark blue component until the light-blue component {\it disappears} at 0. See \url{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd} for an interactive demonstration.

 \subsubsection*{The UA can produce conservative estimates of $\pi_0$}

The illustration in Figure \ref{fig:ZA} suggests that the UA will produce smaller estimates of $\pi_0$ than existing methods.
Consequently \ash will estimate smaller $\lfdr$s and FDRs, and so identify more significant discoveries at a given threshold. 
This is desirable, provided that these estimates remain conservative: that is,
provided that $\hat\pi_0$ does not underestimate the true $\pi_0$ and $\hat\lfdr$ does not underestimate the true $\lfdr$.
The penalty term (\ref{eqn:penalty}) aims to ensure this conservative behavior. To check its effectiveness
we performed simulations under various alternative scenarios (i.e. various distributions for the non-zero effects, which we denote $g_1$), and values
for $\pi_0$. The alternative distributions are shown in Figure \ref{fig:altdens}, with details in Table \ref{table:scenarios}.
They range from a ``spiky" distribution -- where many non-zero $\beta$ are
too close to  zero to be reliably detected, making reliable estimation of $\pi_0$ essentially impossible -- to a much
flatter distribution, which is a normal distribution with large variance (``big-normal") -- where most non-zero $\beta$ are easily detected
making reliable estimation of $\pi_0$ easier. We also include one asymmetric distribution (``skew"), and one clearly bimodal distribution (``bimodal"),
which, although we view as generally unrealistic, we include to assess robustness of \ash~to deviations from the UA.

% How do these differences influence estimates of the FDR? To address this, note that  -- in this simple case where the 
% $\shat_j$ are all identical --  the primary driver of differences in FDR estimates among methods is {\it differences in their estimates
% of $\pi_0$}. To see why, observe that, given an estimate $\hat\pi_0$ for $\pi_0$, equation (\ref{eqn:FDRest}) provides a 
% strikingly natural non-parametric estimate for the FDR (provided the denominator is not very small). Thus, 
% although the methods considered here differ from one another in many ways, having estimated $\pi_0$
% they should, if sensible, produce estimates of FDR that agree closely with (\ref{eqn:FDRest}). And, indeed, limited
% empirical checks suggest that this is indeed the case (\href{https://www.github.com/stephens999/ash/examples/qvalue.nonp.rmd}).
%

%Thus, differences in FDR estimates among methods are driven primarily by differences in their estimates of $\pi_0$,
% and so to assess differences in FDR estimates among methods we focus on assessing differences in estimates of $\pi_0$.
%Figure \ref{fig:ZA} suggests that, compared with the ZA, the UA will tend to reduce the estimate of $\pi_0$. 
%At the same time, provided the UA holds, we hope still to provide ``conservative" estimates of $\pi_0$ (that is,
%overestimates of $\pi_0$, which will yield overestimates of the FDR).

\begin{figure} 
\begin{center}
\begin{subfigure}{\textwidth}
	\includegraphics[width=\textwidth]{../analysis/figure/plot_egdens.Rmd/scenario_density-1.pdf} 
	\caption{Densities of non-zero effects, $g_1$, used in simulations.} \label{fig:altdens}
\end{subfigure}
\begin{subfigure}{\textwidth}
\includegraphics[width=\textwidth]{../analysis/figure/plot_pi0est.Rmd/plot_pi0est-1.pdf} 
\caption{Comparison of true and estimated values of $\pi_0$. When the UA holds all methods yield conservative (over-)estimates for $\pi_0$, with \ash being least conservative, and hence most accurate. When the UA does not hold (``bimodal" scenario) the \ash estimates are slightly anti-conservative.} \label{fig:pi0sims}
\end{subfigure}
\begin{subfigure}{\textwidth}
\includegraphics[width=\textwidth]{../analysis/figure/plot_lfsr.Rmd/plot_lfdr-1.png} 
\caption{Comparison of true and estimated $\lfdr$ from \ash (ash.n). Black line is $y=x$ and red line is $y=2x$. Estimates of $\lfdr$ are conservative when UA holds, due to conservative estimates of $\pi_0$.} \label{fig:lfdr}
\end{subfigure}
\begin{subfigure}{\textwidth}
\includegraphics[width=\textwidth]{../analysis/figure/plot_lfsr.Rmd/plot_lfsr-1.png} 
\caption{As in c), but for $\lfsr$ instead of $\lfdr$. Estimates of $\lfsr$ are consistently less conservative than $\lfdr$ when UA holds, and also less anti-conservative in bimodal scenario.}  \label{fig:lfsr}
\end{subfigure}
\end{center}
\caption{Results of simulation studies (constant precision $s_j=1$).}
\end{figure}

For each simulation scenario we simulated 100 independent data sets, each with $J=1000$ observations. For each data set we simulated
data as follows:
\begin{enumerate}
\item Simulate $\pi_0  \sim U[0,1]$.
\item For $j=1,\dots,J$, simulate $\beta_j \sim \pi_0 \delta_0 + (1-\pi_0) g_1(\cdot)$.
\item For $j=1,\dots,J$, simulate $\bhat_j | \beta_j \sim N(\beta_j,1)$.
\end{enumerate}
%Thus these simulations assume the same
%precision $1$ for each measurement, and for all methods this precision is assumed known (i.e. $\shat_j=1$).

Figure \ref{fig:pi0sims} compares estimates of $\pi_0$ from \qvalue, \locfdr, \mixfdr and \ash ($y$ axis) with the true values ($x$ axis). 
For \ash we show results for $g_1$ modelled as a mixture of normal components (``ash.n") 
and as a mixture of symmetric uniform components (``ash.u"). (Results using the asymmetric uniforms, which we refer to
as ``half-uniforms", and denote ``ash.hu" in subsequent sections, are here generally similar to ash.u and omitted to avoid over-cluttering figures.)
The results show that \ash provides the smallest more accurate, estimates for $\pi_0$, while remaining conservative
in all scenarios where the UA holds. When the UA does not hold (``bimodal" scenario) the \ash estimates can be slightly anti-conservative.
We view this as a minor concern in practice, since we view such a strong bimodal scenario as unlikely in most applications where FDR methods
are used. (In addition, the effects on $\lfsr$ estimates turn out to be relatively modest; see below).

%\subsubsection{The UA produces less conservative estimates of $\pi_0$}
%
%The data in Figure \ref{fig:ZA} were simulated assuming that the true effects are standard normal, $\beta_j \sim N(0,1)$,
%and so all effects are non-zero and no nulls hold. This scenario is rather different
%from the ``mostly null" scenarios that originally motivated FDR methods, where most tests are null,
%and relatively few are non-null. It could be argued that the existing FDR methods are not well adapted to the 
%``mostly non-null"
%scenario, and we would agree! On the other hand, the distribution of $p$ values
%in Figure \ref{fig:ZA} looks like data that, before this work, we would have been happy to apply
%existing FDR methods to. Further, we believe that both ``mostly non-null" and ``mostly null" scenarios likely occur in practical applications.
%Thus it seems desirable to have methods that perform well in both cases, and this is
%an important feature of the methods we introduce here.
%
%TODO: put FDRs on graph? Simulate with 50\% null?

\subsubsection*{The $\lfsr$ is more robust than \lfdr}

The results above show that \ash can improve on existing methods in producing smaller, more accurate, 
estimates of $\pi_0$, which will lead to more accurate estimates of FDR.
Nonetheless, in many scenarios \ash continues
to substantially over-estimate $\pi_0$ (see the ``spiky" scenario for example). 
This is because these scenarios include 
an appreciable fraction of ``small non-null effects" that are essentially indistinguishable from 0, making accurate
estimation of $\pi_0$ impossible. Put another way, and as is well known, $\pi_0$ is not identifiable:
 the data can effectively provide an upper bound on plausible values of $\pi_0$,
but not a lower bound (because the data cannot rule out that everything is non-null, but with minuscule effects).
To obtain conservative behavior we must estimate $\pi_0$ by this upper bound, which can
be substantially larger than the true value.

Since FDR-related quantities depend quite sensitively on $\pi_0$, the consequence of this 
overestimation of $\pi_0$ is corresponding overestimation of FDR (and $\lfdr$, and $q$ values).
To illustrate, Figure \ref{fig:lfdr} compares the estimated $\lfdr$ from ash.n with the true value (computed using
Bayes rule from the true $g_1$ and $\pi_0$). As predicted, $\lfdr$ is overestimated, especially in scenarios which involve many
non-zero effects that are very near 0 (e.g.~the spiky scenario with $\pi_0$ small) where
$\pi_0$ can be grossly overestimated.  
(Of course other methods will be similarly affected by this: those that more grossly overestimate $\pi_0$, will more grossly overestimate $\lfdr$ and FDR/$q$-values.)

The key point we want to make here is that estimation of $\pi_0$, and the accompanying identifiability issues,
 become substantially less troublesome if we use the local false sign rate $\lfsr$ (\ref{eqn:lfsr}), rather than $\lfdr$, to measure significance. 
 This is because $\lfsr$ is less sensitive to the estimate of $\pi_0$.
To illustrate, Figure \ref{fig:lfsr} compares the estimated $\lfsr$ from ash.n with the true value: although the estimated $\lfsr$ continue to
be conservative, overestimating the truth, the overestimation is substantially less pronounced than for the $\lfdr$, especially for
the ``spiky" scenario. Further, in the bi-modal scenario, the anti-conservative behavior is less pronounced in $\lfsr$ than $\lfdr$.

Note that, compared with previous debates regarding testing, 
this section advances an additional reason for focussing on the sign of the effect, rather than just testing whether it is 0. 
In previous debates authors have argued against testing whether an effect is 0
because it is {\it implausible that effects are exactly 0}. Here we add that {\it even if one believes
that some effects may be exactly zero}, it is still better to focus on the sign, because generally {\it the data are more informative about that question}
and so inferences are more robust to, say, the inevitable mis-estimation of $\pi_0$.
To provide some intuition, consider an observation with a $z$ score of 0. The $\lfdr$ of this observation can range from 0 (if $\pi_0=0$)
to 1 (if $\pi_0=1$). But, assuming a symmetric $g$, the $\lfsr>0.5$ whatever the value of $\pi_0$, because the observation $z=0$ says
nothing about the sign of the effect.
Thus, there are two reasons to use the $\lfsr$ instead of the $\lfdr$: it answers a question that is more generally meaningful (e.g. it applies
whether or not zero effects truly exist),  and estimation of $\lfsr$ is more robust.

Given that we argue for using $\lfsr$ rather than $\lfdr$, one might ask whether we even need a point mass on zero in our analysis.
Indeed, one advantage of the $\lfsr$ is that it makes sense even if no effect is exactly zero. And, 
if we are prepared to assume that no effects are exactly zero, then removing the point mass 
yields smaller and more accurate estimates of $\lfsr$ when that assumption is true (Figure \ref{fig:lfsr-nn}). 
However, there is ``no free lunch":  if in fact some effects are exactly zero
then the analysis with no point mass will tend to be anti-conservative, underestimating $\lfsr$ (Figure \ref{fig:lfsr-s}). 
We conclude that {\it if} ensuring a ``conservative" analysis is important then one should allow for a point mass at 0.

\subsubsection*{The UA helps provide reliable estimates of $g$}

An important advantage of our EB approach based on modelling the effects $\beta_j$, rather than $p$ values or $z$ scores, is that it
can estimate the effects $\beta_j$.
Specifically, it provides a posterior distribution for each $\beta_j$, which can be used
to construct interval estimates, etc.
%Importantly, because the posterior distribution for $\beta_j$ depends on the ``prior" effect-size distribution $g$, which is estimated from {\it all} the observations,
%this posterior reflects information from all the observations and not only from the observation $\bhat_j$. 
%In our case, because we assume $g$ to be unimodal about zero, this prior always ``shrinks" the posterior distributions
%towards zero, with the strength of the shrinkage being estimated from the data.
Further, because the posterior distribution is, by definition,
conditional on the observed data, interval estimates based on posterior distributions are also valid Bayesian inferences for any subset of the effects that have
been selected based on the observed data. This kind of ``post-selection" validity is much harder to achieve in the frequentist paradigm.
In particular the posterior distribution solves the (Bayesian analogue of the) ``False Coverage Rate" problem posed by
\cite{benjamini2005false} which \cite{efron2008microarrays} summarizes as follows: ``having applied FDR methods to select a set of nonnull cases,
how can confidence intervals be assigned to the true
effect size for each selected case?". \cite{efron2008microarrays} notes the potential for EB approaches to tackle this problem,
and \cite{zhao2012empirical} considers in detail the case where the non-null effects are normally distributed.
%but also notes the difficulties of implementing a stable general algorithm.

The ability of the EB approach to provide valid ``post-selection" interval estimates is extremely attractive in principle.
But its usefulness in practice
depends on reliably estimating the distribution $g$. Estimating $g$ is a ``deconvolution problem",
which are notoriously difficult in general. Indeed, Efron emphasizes
the difficulties of implementing a stable general algorithm, noting in his rejoinder
``the effort foundered on practical difficulties involving the perils of deconvolution... Maybe I am trying
to be overly nonparametric ... but it is hard to imagine a
generally satisfactory parametric formulation..." (\cite{efron2008microarrays} rejoinder, p46).
We argue here that the UA can greatly simplify the deconvolution problem, producing both computationally and statistically stable estimates of $g$.

To illustrate, we compare the
estimated $g$ from \ash (under the UA) with the non-parametric maximum likelihood estimate (NPMLE) for 
$g$ (i.e.~estimated entirely non-parametrically without the unimodal constraint). 
The NPMLE is straightforward to compute, thanks to 
\cite{koenker2014convex}. Figure \ref{fig:egcdf} shows results for one dataset under each of six different scenarios.
The estimated cdf from \ash is generally closer to the truth, even in the bi-modal scenario.
(\ash tends to systematically overestimate the mass of $g$ near zero; this can be avoided 
by removing the penalty term (\ref{eqn:penalty}); Figure \ref{fig:pen-nopen}.)
Furthermore, the estimate from \ash is also substantially more ``regular" than the  NPMLE, 
which has several almost-vertical segments indicative of a concentration of density in the estimated $g$ at those locations. 
Indeed the NPMLE is a discrete distribution \cite{koenker2014convex}, so this kind of concentration will always occur. 
The UA prevents this concentration, effectively regularizing the estimated $g$. 
While the UA is not the only way to achieve this, we view it as attractive and widely applicable (although not universally applicable).

%Interestingly, even in the ``bimodal" scenario $\ash$ 
%is more accurate than \mixfdr: although \mixfdr is capable, in principle, of fitting the multiple modes of $g$, it does not do this well here. 
%Possibly the noise level here is sufficiently large to make reliable estimation of the multiple modes difficult. 
%Indeed, in multi-modal simulations where the multiple modes are sufficiently well-spaced
%to be clearly visible in the observed $\bhat$, \mixfdr fits these modes 
%(\url{http://stephenslab.github.io/ash/analysis/check_mixfdr_lownoise.html}). 
%Of course, we would not advocate the UA in settings where multi-modality is 
%clearly visible in the observed $\bhat$.
%
%We note one caveat on the accuracy of estimated $g$: due to the penalty term (\ref{eqn:penalty}) \ash tends to systematically
%overestimate the mass of $g$ near zero. This can be seen in Figure \ref{fig:egcdf-reduce}: the estimated cdf is generally below the true cdf just to the left of zero,
%and above the true cdf just to the right of zero. Averaging the cdf over many replicates confirms this systematic effect (Figure \ref{fig:egcdf-reduce}b),
%and applying our methods without the penalty term removes this systematic effect, although at the cost of sometimes 
%under-estimating $\pi_0$ (Figure \ref{fig:egcdf-reduce}c).

\begin{figure}[h!] 
\begin{center}
	\includegraphics[height=4.5in]{../analysis/figure/plot_cdf_eg.Rmd/egcdf-reduce-with-npmle-1.pdf}
	\caption{Comparison of estimated cdfs from ashr and NPMLE. Different ash methods perform similarly, so only {\tt ash.hu} is shown for clarity. Each panel shows results for a single example data set, one for each scenario in Figure \ref{fig:altdens}. The results illustrate how the unimodal assumption made by ash regularizes the estimated cdfs compared with the NPMLE.} \label{fig:egcdf}
\end{center}
\end{figure}

%making the UA provides an attractive and simple way to achieve estimating $g$  large
%subset of cases of interest (those where the UA can reasonably be assumed to hold).

%In practice, the utility of the posterior distributions noted above depends primarily on the accuracy
%of the estimate of the distribution $g(\cdot)$.

 %quote from efron rejoinder , p46:
%MorrisÕ Section 3 is especially pertinent. His formula
%for p(?|z) is related to my discussion of the
%BenjaminiÐYekutieli False Coverage Rate method in
%Section 7, particularly (7.2)Ð(7.4). Originally I had
%hoped to develop an empirical Bayes method for estimating
%such models, but the effort foundered on practical
%difficulties involving the perils of deconvolution.
%Section 6 on the Òone-group modelÓ is the ugly duckling
%of the current paper, but it bears on some important
%points raised in the discussion. Figure 7 concerns
%a fuzzy version of simultaneous hypothesis testing,
%where, as in MorrisÕ hospital example, the usual singlepoint
%null hypotheses seem unequal to the task. The
%development from (6.6) onwards, particularly (6.12),
%bears on the possibility of nonnormal null distributions,
%and is about as far as I can go in answering Professors
%Rice and SpiegelhalterÕs penultimate question.
%With g(?) a normal distribution, model (6.1) returns
%us to the realm of Stein estimation, the scene of my
%happy collaborations with Carl Morris. I continue to
%be surprised at how much harder bumpy, nonnormal
%models like (7.1) are to deal with. JamesÐStein estimation
%works fine with, say, N = 10 component problems,
%while the RobbinsÕ form of empirical Bayes appropriate
%to (7.1) seems to require hundreds or thousands.
%The information calculations in Efron (2008)
%reinforce this gloomy assessment. Maybe I am trying
%to be overly nonparametric in constructing the empirical
%Bayes Fdr estimates, but it is hard to imagine a
%generally satisfactory parametric formulation for (6.1).
%Or perhaps it is just that hypothesis testing is more demanding
%than estimation.
%My Note: (6.1) is the general model (mu \sim g()).
%7.1 is the case where the effects are a mixture of zero and N(-3,1).

\subsubsection*{Calibration of posterior intervals}

To quantify the effects of errors in estimates of $g$ we examine the calibration of the resulting posterior distributions (averaged over 100 simulations in each Scenario). 
Specifically we examine the empirical coverage of nominal lower 95\% credible bounds for a) all observations; b) significant negative discoveries; c) significant positive discoveries.  We examine only lower bounds because the results for upper bounds follow by 
symmetry (except for the one asymmetric scenario). We separately examine positive and negative discoveries because the lower bound plays a different role
in each case: for negative discoveries the lower bound is typically large and negative and limits how big (in absolute value) 
the effect could be; for positive discoveries the lower bound 
is positive, and limits how small (in absolute value) the effect could be. Intuitively, the lower bound for negative discoveries depends on the accuracy of $g$ in its tail,
whereas for positive discoveries it is more dependent on the accuracy of $g$ in the center.

The results are shown in Table \ref{tab:coverage}.  Most of the empirical coverage rates are in the range 0.92-0.96 for nominal coverage of 0.95, which we view
as adequate for practical applications. The strongest deviations from nominal rates are noted and discussed in the table captions.
One general issue is that the methods based on mixtures of uniform distributions often slightly curtail the tail of $g$, causing the probability of very large outlying effects to be understated;  see also \url{http://stephenslab.github.io/ash/analysis/efron.fcr.html}.

% As noted above, the penalty term (\ref{eqn:penalty}) causes systematic over-estimation of the mass of $g$ near 0,  which one would expect to overshrink interval estimates towards zero. 
%The effect of overshrinking on the left tail area is different for ``negative discoveries" and ``positive discoveries": for negative discoveries the lower credible bound is negative and overshrinking (moving it too far towards zero) makes the left tail too big; conversely, for positive discoveries, the lower confidence bound is positive, and  
%overshrinking (moving it too far towards zero) makes the tail too small. In the context of reliably identifying ``large" effects we might worry more about under-shrinkage, 
%and be prepared to put up with some overshrinkage to avoid it.

%Coverage rates are given in:
%\href{file:///Users/stephens/Dropbox/Documents/git/ash/dsc-shrink/investigate_cdf.html}

%Note - this could probably be solved by making the small components symetric - the idea is that the asymetry only comes in with the tail.

\begin{table}[!ht]
\begin{subtable}{\textwidth}
	\centering\input{../analysis/table/scoverage_all.tex}
	\caption{All observations. Coverage rates are generally satisfactory, except for the extreme ``spiky" scenario. This is due to the penalty term (\ref{eqn:penalty}) which tends to cause over-shrinking towards zero. Removing this penalty term produces coverage rates closer to the nominal levels for uniform and normal methods (Table \ref{tab:nopen}). Removing the penalty in the half-uniform case is not recommended (see Appendix).}
\end{subtable}

\begin{subtable}{\textwidth}
\centering\input{../analysis/table/scoverage_neg.tex}
\caption{``Significant" negative discoveries. Coverage rates are generally satisfactory, except for the uniform-based methods in the spiky and near-normal scenarios,
and the normal-based method in the flat-top scenario. These results likely reflect inaccurate estimates of the tails of $g$ due to a disconnect between the tail of $g$ and the component distributions in these cases. For example, the uniform methods sometimes substantially underestimate the length of the tail of $g$ in these long-tailed scenarios, 
causing over-shrinkage of the tail toward 0.}
\end{subtable}

\begin{subtable}{\textwidth}
\centering\input{../analysis/table/scoverage_pos.tex}
\caption{``Significant" positive discoveries. Coverage rates are generally satisfactory, except for the symmetric methods under the asymmetric (``skew") scenario. }
\end{subtable}

\caption{Table of empirical coverage for nominal 95\% lower credible bounds} \label{tab:coverage}
\end{table}

%
%\subsubsection*{Reverse asymptotics}
%
%Note that if observation precisions are all similar and most observations are null then all the methods likely to give similar results?
%
%Note: we'd like to make sure it behaves ok in the case where most observations are null.
%Perhaps simulate with 999 null and 1 alternative?
%Also maybe look at performance for small sample sizes - e.g. n=1?
%
%Note that, even though the exact value of the point mass $\pi_0$ cannot, in general, be reliably estimated, the actual underlying distribution $g$ 
%can be quite accurately estimated from the data, provided we assess accuracy by a metric that is not sensitive to $\pi_0$: for example
%by measuring the difference between the true and estimated cumulative distribution function (cdf). See supplementary Figure \ref{} for examples.
%

\subsection*{Differing measurement precision across units}

 We turn now to the second important component of our work: allowing for varying
 measurement precision across units. The key to this is the use of a likelihood,
 (\ref{eqn:normlik}) or (\ref{eqn:tlik}), that explicitly incorporates the measurement precision (standard error) of each $\bhat_j$.

  To illustrate, we conduct a simulation where half the measurements are quite precise (standard error $s_j = 1$), and the other half are very poor
 ($s_j=10$).  In both cases,
we assume that half the effects are null
and the other half are normally distributed with standard deviation 1:
\begin{equation}
p(\beta) = 0.5 \delta_0(\beta) + 0.5 N(\beta; 0,1).
\end{equation}
In this setting, the poor-precision measurements $(s_j=10)$ tell us very little, and any sane analysis should effectively ignore them. 
 However, this is not the case in standard FDR-type analyses (Figure \ref{fig:goodpoor}). This is because the poor measurements 
 produce $p$ values that are approximately uniform  (Figure \ref{fig:goodpoor}a), 
 which, when combined with the good-precision measurements, dilute the overall signal 
 (e.g.~they reduce the density of $p$ values near 0).
 This is reflected in the results of FDR methods like $\qvalue$ and $\locfdr$:
the estimated error rates ($q$-values, or $\lfdr$ values) for the good-precision observations increase when the low-precision observations are included in the analysis
(Figure \ref{fig:goodpoor}b). In contrast, the results from $\ash$ 
for the good-precision observations are unaffected by including the low-precision observations in the analysis (Figure \ref{fig:goodpoor}b).

\begin{figure}[hp!]
\begin{subfigure}{\textwidth}
\centering\includegraphics[width=4in]{../analysis/figure/make_GOODPOOR_figs.Rmd/GOODPOOReg_hist-1.pdf} 
    \label{fig:goodpoor-hist}
    \caption{Density histograms of $p$ values for good-precision, poor-precision, and combined observations. The combined data show less signal than the good-precision data, due to the contamination effect of the poor-precision measurements.}
 \end{subfigure}
\begin{subfigure}{\textwidth}
\centering\includegraphics[width=4in]{../analysis/figure/make_GOODPOOR_figs.Rmd/GOODPOOReg_scatter-1.pdf}
    \caption{Results of different methods applied to good-precision observations only ($x$ axis) and combined data ($y$ axis). Each point shows the ``significance" ($q$ values from \qvalue; $\lfdr$ for \locfdr; $\lfsr$ for \ash) of a good-precision observation under the two different analyses. For existing methods including the poor-precision observations reduces significance of good-precision observations, whereas for \ash the poor-precision observations have little effect (because they have a very flat likelihood).} \label{fig:goodpoor-scatter}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\centering\includegraphics[width=3in]{../analysis/figure/make_GOODPOOR_figs.Rmd/lfsr_vs_pval_GOODPOOR_single-2.pdf}
    \caption{The relationship between $\lfsr$ and $p$-value is different for good-precision ($s=1$) and low-precision ($s=10$) measurements: \ash assigns the low-precision measurements a higher $\lfsr$, effectively downweighting them.    %The default prior produces better performance here, as expected because the simulations match its assumptions. In practice one \ash can assess which prior best matches the data by comparing likelihoods (see text).
    } \label{fig:lfsr_pval}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\centering\includegraphics[width=3in]{../analysis/figure/make_GOODPOOR_figs.Rmd/tp-fp-1.pdf}
    \caption{Trade-off between true positives ($y$) vs false positives ($x$) as the significance threshold ($\lfsr$ or $p$ value) is varied. By downweighting the low-precision observations \ash re-orders the significance of observations, producing more true positives at a given number of false positives.
    %The default prior produces better performance here, as expected because the simulations match its assumptions. In practice one \ash can assess which prior best matches the data by comparing likelihoods (see text).
    } \label{fig:tpfp}
\end{subfigure}
\caption{Simulations showing how, with existing methods, but not \ash, 
poor-precision observations can contaminate signal from good-precision observations.} \label{fig:goodpoor}
\end{figure}

Another consequence of incorporating measurement precision into the likelihood
is that \ash can re-order the significance of observations compared
 with the original $p$ values or $z$ scores. Effectively \ash downweights the poor-precision 
 observations  by assigning them a higher $\lfsr$ than 
 good precision measurements that have the same $p$ value
 (Figure \ref{fig:lfsr_pval}). 
 The intuition is that, due to their poor precision, these measurements contain very little information about the sign of the effects (or indeed any other aspect of the effects),
 and so the $\lfsr$ for these poor-precision measurements is always high; see \cite{guan.stephens.08} for related discussion. Here this re-ordering results in improved performance: \ash identifies more true positive effects at a given level of false positives (Figure \ref{fig:tpfp}).

 \subsubsection*{Dependence of $\beta_j$ on $s_j$}

 Although downweighting low-precision observations may seem intuitive, we must now
 confess that the issues are more subtle than our treatment above suggests. 
 Specifically, it turns out that the downweighting behaviour depends on an assumption 
that we have made up to now, that $\beta_j$ is independent
of $s_j$ (Equation (\ref{eqn:beta})). In practice this assumption may not hold. 
For example, in gene expression studies, 
genes with higher biological variance may tend to exhibit larger effects $\beta_j$ (because they are
less constrained). These genes will also tend to have larger $s_j$, 
inducing a dependence between $\beta_j$ and $s_j$.

Motivated by this, we generalize the prior (\ref{eqn:beta}) to
 \begin{equation} \label{eqn:beta-alpha}
 \frac{\beta_j}{\shat_j^\alpha} \, \big |  \shat_j \sim g(\cdot; \pi)
 \end{equation}
where $\alpha$ is to be estimated or specified. 
Setting $\alpha=0$ yields (\ref{eqn:beta}). Setting $\alpha>0$ implies that 
the effects with larger standard error tend to be larger (in absolute value).
Fitting this model for any $\alpha$ is straightforward using the same methods as for $\alpha=0$ (see Methods).
%(Note that results above that were generated with all $\shat_j=1$ would not be affected by choice of $\alpha$.)

The case $\alpha=1$ in (\ref{eqn:beta-alpha}) is of special interest because it corresponds most closely
to existing methods. Specifically, it can be shown that with $\alpha=1$ the $\lfsr_j$ values from {\tt ash.n}
are monotonic in the $p$ values: effects with smaller $p$ values have smaller $\lfsr_j$. 
This result generalizes a result in \cite{wakefield:2009}, 
who referred to a prior that produces the same
ranking as the $p$ values as a ``$p$ value prior".
Of course, if the $\lfsr_j$ are monotonic in the $p$ values then the downweighting and
reordering of significance illustrated in Figure \ref{fig:goodpoor} will not occur.
The intuition is that under $\alpha=1$ the poor precision observations have larger effects sizes,
and consequently the same power 
as the high-precision observations -- under these conditions the poor precision observations are not ``contaminating" the high precision observations, and so downweighting them is unnecessary.
Thus running \ash with $\alpha=1$ will produce the same significance ranking as existing methods.
Nonetheless, it is not equivalent to them, and indeed still has the benefits outlined previously: due to the UA \ash can produce less conservative estimates of $\pi_0$ and $\lfdr_j$; and because \ash models the $\beta_j$ it can produce interval estimates.

As an aside, we note that with $\alpha=1$ the \ash estimates of both $g$ and $\lfsr_j$ depend on the pairs $(\bhat_j,\shat_j)$ only through  the $z$ scores $z_j = \bhat_j/\shat_j$ (though interval estimates for $\beta_j$ still depend on 
$(\bhat_j,\shat_j)$). This means that \ash can be run (with $\alpha=1$) in settings where the $z_j$ are available but 
the $(\bhat_j,\shat_j)$ are not. It also opens the intriguing possibility 
of running \ash on $p$ values obtained from any test (e.g.~a permutation test), by first converting each $p$ value to a corresponding $z$ score. However, the meaning of and motivation for the UA may be unclear in such settings, and caution seems warranted before proceeding along these lines.

In practice, appropriate choice of $\alpha$ will depend on the actual relationship between $\beta_j$ and $s_j$, which will be dataset-specific. Further, although we have focussed on the special cases $\alpha=0$ and $\alpha=1$, there seems no strong reason to expect that either will necessarily be optimal in practice.
Following the logic of the EB approach we suggest selecting $\alpha$ by maximum likelihood, which is 
implemented in our software using a simple 1-d grid search (implementation due to C. Dai).

%Framing the problem in this way -- that is, as comparing different modelling assumptions for $\beta_j$, rather than as comparing ``modelling $\beta_j$" vs  ``modelling $z_j$" (or ``modelling $p_j$") -- has the important advantage that 
%likelihood-based methods can be used to select $\alpha$. For example,

% He named this prior the ``$p$-value prior" because it can be thought of as the 
%implicit prior assumption that is made when using $p$ values to rank observations.
%Wakefield's $p$-value prior assumes that the prior for $\beta_j$ depends on its standard error $s_j$,
%and specifically that 
%\begin{equation}
%\beta_j | s_j \sim N(0,K s^2_j) \text{for some constant $K$.}
%\end{equation}
%Here we observe that Wakefield's result extends to our mixture of (zero-mean) normal priors. 
%Specifically, if, instead of assuming that $\beta_j$ is independent
%of $s_j$ as we have up to now, we assume that $z_j=\beta_j/s_j$ is independent of $s_j$, and drawn from a mixture of zero-mean normal distributions,
%then the $\lfsr$ provides the {\it same ranking} of observations as the $z$ scores and $p$ values,
%as is illustrated in Figure \ref{fig:lfsr_pval}, right panel.

%\begin{result}
%If the prior distribution on $\beta/s | s$ is a mixture of zero-centered normal distributions, (\ref{eqn:mixnorm}), and the likelihood for $\beta$ is 
%given by $\bhat | s, \beta \sim N(0,s^2)$ then $p(\beta>0 | \bhat, s)$ is a monotonic increasing function of the $z$ score, $\bhat/s$. 
%\end{result}

%To summarize: the behaviour of \ash depends on the value of $\alpha$ assumed in (\ref{eqn:beta-alpha}). 
%Up to now we have assumed $\alpha=0$, which implies that effects are identically distributed, independent of their standard error.
%In contrast, existing methods effectively assume $\alpha=1$, which implies that effects tend to scale proportional to their standard error.
%Whether it is better to set $\alpha=0$ or $\alpha=1$ in practice will depend on actual relationship between $\beta_j$ and $s_j$,
%which will be dataset-specific. Indeed, it seems quite likely that, in general, the optimal $\alpha$ may be some other (non-integer) value.
%(Alternatively, since fractional values of $\alpha$ may be tricky to interpret, users
%may prefer to simply choose whichever of $\alpha=0,1$ yields the highest likelihood.)

%\begin{result}
%Treat the standard errors $s_j$ as fixed and given. Under the normal likelihood (\ref{eqn:normlik}), and using a mixture of normals for the prior distribution 
%(\ref{eqn:beta-alpha}) with $\alpha=1$, 
%the (Bayesian) $\lfsr_j$ and $\lfdr_j$ are all monotonic with $|t_j|$ where $t_j = \bhat_j/s_j$, and with the corresponding $p$ value $(p_j)$. Also, the shrinkage factor ($\text{Posterior Mean($\beta_j$)}/\bhat_j$) 
%is montonic with $|t_j|$. 
%\end{result}

%It is straightforward to prove, by performing Bayesian analysis for the standardized b$-J$
%Actually the key requirement is that the prior be such that the $t$ statistics $beta_j/s_j | s_j$ are identically distributed under the alternative.

%Most previous authors of EB approaches to false discovery rates \cite{efron2008microarrays,muralidharan2010empirical}
%have worked with models for $z$ scores $\beta_j/s_j$, rather than models for the effects $\beta_j$.
%In doing this they assume (implicitly or explicitly) that the $z$ scores are identically distributed under the alternative model.
%This is analogous
%to assuming $\alpha=1$ in (ref{eqn:beta-alpha}). And, given the result above, 
%standard frequentist FDR methods, like \qvalue and Benjamini--Hochberg, 
%that rank tests by their $p$ value, could also be viewed as making the same implicit assumption.

%\begin{figure}
%\begin{subfigure}{\textwidth}
%\centering\centering\includegraphics[width=4in]{../analysis/figure/make_GOODPOOR_figs.Rmd/lfsr_vs_pval_GOODPOOR-1.pdf} \label{fig:lfsr_pval}
%    \caption{Illustration of how prior assumptions re-order significance. Left panel shows results under the ``default prior", which assumes $\beta_j$ are identically distributed, independent of $s_j$ ($\alpha=0$ in (\ref{eqn:beta-alpha})). Right panel shows results under the ``$p$-value prior", which assumes $z$ scores $\beta_j/s_j$ are identically distributed, independent of $s_j$ (\ref{eqn:beta-alpha}). This right panel is exactly the same as would be obtained by any monotonic function of the $p$ values to measure significance (e.g. $p$ values, $q$ values, $|z|$ scores)}
% \end{subfigure}
%
%\caption{} 
%\end{figure}

%\subsubsection*{Adaptive Shrinkage}
%
%Johnstone and Silverman \cite{johnstone2004needles} (see also \cite{raykar2011empirical}) note the ability of EB methods to adapt to overall signal strength
%when attempting to identify and estimate strong signals. The idea is that, if most $\beta_j$ are truly at or near zero then (given enough data) 
%the estimated prior distribution $\hat{g}$ will reflect this by being concentrated near 0, and the posterior distributions
% $p(\beta_j | \bhat_j, \shat_j,\hat{g})$ will consequently also tend to concentrate around zero (strong shrinkage). 
% In contrast, if most $\beta_j$ are large then $\hat{g}$ will be flatter, resulting in less shrinkage.
% Indeed, in the limit as $\hat{g}$ becomes flatter and flatter the posterior distribution on $\beta_j$ becomes $N(\bhat_j,\shat_j)$,
% and the Bayesian credible intervals match standard confidence intervals, which might be considered ``no shrinkage".
%
%Our methods we build on this idea in two ways. First,
%by using flexible semi-parametric approaches to estimate $g$ we aim to maximise the potential for this adaptive behaviour.
%Second, by incorporating the precision of each measurement into the likelihood, (\ref{eqn:normlik}) or (\ref{eqn:tlik}), we ensure that shrinkage
%adapts to measurement precision: more precise measurements undergo
% less shrinkage than less precise measurements.  This is because observations $\bhat_j$ with larger standard errors have flatter likelihoods 
%than observations with
%small standard error, and so their posteriors will be more affected by the prior, which, being unimodal at 0, tends to shrink estimates towards 0. 
%To emphasize these two key features we refer to our method as ``{\bf a}daptive {\bf sh}rinkage".
%Figure \ref{fig:ash} illustrates the idea by contrasting results for two simulation scenarios with different signal strengths, where observations vary in their measurment precision.
%
%\begin{figure}[!ht]
%\begin{center}
%\includegraphics[width=400px]{../examples/ash.pdf} 
%\end{center}
%\caption{Figure illustrating adaptive shrinkage. Results are shown for two different scenarios: (left) ``big-normal" where the effects have a wide distribution; (right) ``spiky", where the effects are more concentrated near 0. The standard error of each observation, $s_j$, was simulated from Inverse Gamma(5,5), which resulted in $s_j$ varying from 0.34 to 6.7.  The shrunken estimates ($y$ axis) are plotted against the observed estimates ($x$ axis), with color indicating the (square root of the) precision of each measurement (blue=lower-precision; red=higher-precision). The two key features are i) shrinkage is adaptive to signal in data, so shrinkage is stronger for the spiky scenario; ii) shrinkage is adaptive to the precision of each measurement, so less precise observations (blue) are shrunken more strongly.} \label{fig:ash}
%\end{figure}

%DISCUSSION ON ADAPTIVE SHRINKAGE; BEST TO PUT IN RESULTS?
%To understand how this approach leads to ``adaptive shrinkage", think of (\ref{eqn:betaj}) as defining the ``prior distribution" for $\beta_j$ (determined by hyper-parameters $\pi$),  and (\ref{eqn:bhat}) as defining the likelihood for $\beta_j$. Because the prior distribution is unimodal about zero, 
%the posterior distribution of $\beta$ will be ``shrunk" towards zero (compared with the likelihood), with the
%amount of shrinkage being determined by a) how concentrated the prior is near zero (determined by $\hat{\pi}$) and b) how concentrated the likelihood is around $\bhat_j$.
%For example, if most $\beta_j$ are truly at or near zero then $\hat\pi$ will concentrate on zero and very small values of $\sigma_k$, and this will produce stronger shrinkage
%than if $\beta_j$ are mostly large. And observations $\bhat_j$ with large standard error will have flatter likelihoods than observations with
%small standard error, and so their posteriors will be more affected by the prior, and undergo more shrinkage.

%previous work based on normal means:
%http://link.springer.com/article/10.1007/BF02562623

%\subsection{Null behaviour}

%Results of simulations under the null.

%\subsection{Comparison of different underlying component distributions}
%
%Because the mixture of zero-anchored uniforms is more flexible than the mixture of zero-centered uniforms, which is more flexible
%than the mixture of zero-centered normals, we expect that the likelihoods will satisfy $L_{hu} > L_u > L_n$. However, for Scenario A we expect the differences in
%likelihood to be small, because the true model here is a mixture of zero-centered normals, so any gain of $L_{hu}$ and $L_u$ over $L_n$ is necessarily due to ``overfitting".
%For Scenario B, we expect that $L_{hu}$ and $L_{u}$ to be significantly larger than $L_n$, but any 
%increase of $L_{hu}$ over $L_u$ must be due to overfitting (because the truth is symmetric). Finally, for Scenario C we
%expect $L_{hu}$ to be significantly larger than the others.
%
%Figure \ref{fig:rmse_loglik_boxplot} illustrates this.
%
%\begin{figure}[!ht] \label{fig:rmse_loglik_boxplot}
%\begin{center}
%\includegraphics[width=6.5in]{Rcode/figures/simABC_egdens.pdf}
%%\includegraphics[width=6.5in]{Rcode/figures/rmse_loglik_boxplot.pdf}
%\includegraphics[width=6.5in]{Rcode/figures/rmse_boxplot.pdf}
%\includegraphics[width=6.5in]{Rcode/figures/loglik_boxplot.pdf}
%\end{center}
%\caption{}
%\end{figure}

%Notes from Adrian Raftery - 
%Newcombe (1886)
%Dean and Raftery
%Uniform is super-efficient

\section*{Discussion}

We have presented an Empirical Bayes approach to large-scale multiple testing that emphasizes two ideas.
First, we emphasize the potential benefits of using two numbers ($\bhat$, and its standard error)
rather than just one number (a $p$ value or $z$ score)  to summarize the information on each test.
While requiring two numbers is slightly more onorous than requiring one, in many settings these
numbers are easily available and if so we argue it makes sense to use them.
Second, we note the potential benefits -- both statistical and computational -- of assuming that the effects come from a unimodal distribution, and provide
flexible implementations for performing inference under this assumption. We also introduce the ``false sign rate" as
an alternative measure of error to the FDR, and illustrate its improved robustness to errors in model fit, particularly mis-estimation
of the proportion of null tests, $\pi_0$.

Multiple testing is often referred to as a ``problem" or a ``burden". In our opinion, EB approaches turn this idea on its head,
treating multiple testing as an {\it opportunity}: specifically, an opportunity to learn about the prior distributions, and other modelling assumptions,
to improve inference and make informed decisions about significance (see also \cite{greenland1991empirical}). This view
also emphasizes that, what matters in multiple testing settings is {\it not} the number of tests, but the {\it results} of the tests.
Indeed, the FDR at a given fixed threshold does not depend on the number of tests: as the number of tests increases, both the true positives and false positives increase linearly,
and the FDR remains the same. (If this intuitive argument does not convince, see \cite{storey.03}, and note that the FDR at a given $p$ value threshold does not depend on the number of tests $m$.) Conversely, the FDR {\it does} depend on the overall distribution of effects, and particularly on $\pi_0$ for example. 
The EB approach captures this dependence in an intuitive way:
if there are lots of strong signals then we infer $\pi_0$ to be small, and the estimated FDR (or $\lfdr$, or $\lfsr$) at a given threshold may be low, even if a large number of tests were performed; and conversely
if there are no strong signals then we infer $\pi_0$ to be large and the FDR at the same threshold may be high, even if relatively few tests were performed. 
More generally, overall signal strength is reflected in the estimated $g$, which in turn influences the estimated FDR.

Two important practical issues that we have not addressed here are correlations among tests,
and the potential for deviations from the theoretical null distributions of test statistics. These two
issues are connected: specifically, unmeasured confounding factors can cause both correlations among tests
and deviations from the theoretical null \cite{efron2007correlation,leek:2007}. And although there are certainly other
factors that could cause dependence among tests, unmeasured confounders are perhaps the most worrisome in practice
because they can induce strong correlations among large numbers of tests and profoundly impact results, ultimately resulting
in too many hypotheses being rejected and a failure to control FDR.
We are acutely aware that, because our method is less conservative than existing methods, it may unwittingly
exacerbate these issues if they are not adequately dealt with.
Approaches to deal with unmeasured confounders can be largely divided into two types: those
that simply attempt to correct for the resulting inflation of test statistics \cite{devlin1999genomic,efron2004large}, and those
that attempt to infer confounders using clustering, principal components analysis, or factor models 
\cite{pritchard.stephens.rosenberg.donnelly.00,price:2006,leek:2007,gagnon2012using},
and then correct for them in computation of the test statistics (in our case, $\bhat,\shat$). When these latter approaches are viable
they provide perhaps the most satisfactory solution, and are certainly a good fit for our framework. Alternatively, our methods
could be modified to allow for test statistic inflation, an idea that may be worth pursuing in future work.

Another important practical issue is the challenge of small sample sizes. For example, in genomics applications researchers sometimes
attempt to identify differences between two conditions based on only a handful of samples in each. In such settings the
normal likelihood approximation (\ref{eqn:normlik}) will be inadequate. And, although the $t$ likelihood (\ref{eqn:tlik}) partially
addresses this issue, it is also, it turns out, not entirely satisfactory.  The root of the problem is that, with small sample sizes, raw estimated standard
errors $\shat_j$ can be horribly variable. In genomics it is routine to address this issue by applying EB methods \cite{smyth:2004} to ``moderate" (i.e.~shrink) variance estimates,
before computing $p$ values from ``moderated" test statistics. We are currently investigating how
our methods should incorporate such ``moderated" variance estimates to make it applicable to small sample settings.

Our approach involves compromises between flexibility, generality, and simplicity on the one hand, and statistical efficiency and principle on the other.
For example, in using an EB approach that uses a point estimate for $g$, rather than a fully Bayes approach that accounts for uncertainty in $g$,
we have opted for simplicity over statistical principle. And in summarizing every test by
two numbers and making a normal or $t$ approximation to the likelihood, we have aimed to produce generic methods that 
 can be applied whenever such summary data are available -- just as \qvalue can be applied to any set of $p$ values for example -- although
 possibly at the expense of statistical efficiency compared with developing multiple tailored approaches based on context-specific likelihoods. 
 Any attempt to produce generic methods will involve compromise between generality and efficiency.
 In genomics, many analyses -- not only FDR-based analyses -- involve first computing a series of $p$ values before
 subjecting them to some further downstream analysis. An important message here is 
 that working with two numbers ($\bhat_j,\shat_j$), rather than one ($p_j$ or $z_j$), 
 can yield substantial gains in functionality (e.g.~ estimating effect sizes, as well as testing; accounting for variations in
measurement precision across units) while losing only a little in generality.  We hope that our work will
encourage development of methods that exploit this idea in other contexts.

%This is connected with work by \cite{wakefield:2009} (see also \cite{johnson:2008}) who performs Bayesian inference
%from summary statistics, computing $p(\beta_j | \bhat_j, \shat_j)$ for a single $j$,
%with a specific fixed prior on $\beta_j$. Essentially we extend this to multiple $j$, using a hierarchical model to connect the observations.

%
%
%Both \locfdr and \mixfdr allow for deviations from
%the theoretical null distribution by allowing that, under the null, the $z$ scores are $N(0,\sigma^2)$ with $\sigma$ estimated from the data rather
%than fixed to 1. Within our approach this idea could be incorporated by changing the likelihood - for example, to assume $\bhat_j \sim N(0,s_ 
%
%These are tricky issues that, as far as we are aware, can affect
%all approaches to FDR analysis. An important focus of Efron's work, also implemented in \locfdr, 
% is to relax the assumption that the null $z$ scores are $N(0,1)$, to allow for empirical deviations from this
%theoretical null distribution. 
%
%For example, we and others have worked on Bayesian approaches to genetic association testing, using subjective prior specification of
%effect sizes based on external data (e.g.~\cite{stephens:2009}). However, the large numbers of tests brings the opportunity to learn about
%priors, and here we aim to exploit this

%Choice of mixture components. Use of likelihood to compare different mixture components?

%  
%Directly modeling the p values, or z scores, say via non-parametric methods, 
%can lead to unrealistic distributions being fitted. Put another way, because z scores are the result of adding noise to some
%distribution, the range of distributions they can take is limited Using entirely non-parametric methods loses this information.
%The solution is to model beta hat as a convolution of some distribution g and an error component.

%Before proceeding, we feel obliged to correct the common misconception
% that the use of FDR methods is to ``correct" for the number of tests performed. In fact the FDR does
% not depend on the number of tests. 
%  A better way to think of it is that the FDR accounts for the {\it amount of signal} in the tests that were performed: 
% 
% In other words, FDR corrects for the {\it results} of all tests performed, not the {\it number} of all tests performed.

\section*{Detailed Methods} \label{sec:implement}

\subsection*{Embellishments}

\subsubsection*{More flexible unimodal distributions}

Using a mixture of zero-centered normal distributions for $g$ in (\ref{eqn:mixnorm}) implies that $g$ is not only unimodal, but also symmetric.
Furthermore, even some symmetric unimodal distributions, such as those with a flat top, cannot be well approximated by a mixture of zero-centered normals.
 Therefore, we have implemented a more general approach based on 
\begin{equation} \label{eqn:g-general}
g(\cdot; \pi) = \sum_{k=0}^K \pi_k f_k(\cdot) ,
\end{equation}
where $f_0$ is a point mass on 0, and $f_k$ ($k=1,\dots,K$) are pre-specified component distributions with one of the following forms: 
\begin{enumerate}[(i)]
\item $f_k(\cdot) = N(\cdot; 0, \sigma^2_k)$, \qquad (``ash.n")
\item $f_k(\cdot) = U[\cdot; -a_k,a_k]$,  \qquad (``ash.u")
\item $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$,  \qquad (``ash.hu")
\end{enumerate}
where $U[\cdot; a,b]$ denotes the density of a uniform distribution on $[a,b]$.
(In (iii) we include both components in the mixture (\ref{eqn:g-general}), so a grid of values $a_1,\dots,a_K$ defines $2K+1$ mixture component densities, 
and $\pi$ is a $2K+1$ vector that sums to 1.)
The simplest version (\ref{eqn:mixnorm}) corresponds to (i).
Replacing these with uniform components (ii)-(iii) only slightly complicates calculations
under the normal likelihood (\ref{eqn:normlik}), and greatly simplifies the calculations under the $t$ likelihood
(\ref{eqn:tlik}) introduced below. The use of uniform components here closely mirrors \cite{cordy1997deconvolution}.
(In fact our implementation can handle {\it any} pre-specified uniform or normal distributions for $f_k$ provided they are all from the same family;
however, we restrict our attention here to (i)-(iii) which imply a unimodal $g$.)

Moving from (i) to (iii) the representation (\ref{eqn:g-general}) becomes increasingly flexible. Indeed,
using a large dense grid of $\sigma^2_k$ or $a_k$, (i)-(iii) can respectively approximate,
with arbitrary accuracy,
\begin{enumerate}[(i)]
\item any scale mixture of normals, which includes as special cases
the double exponential (Laplace) distribution, any $t$ distribution, and a very large number of other distributions used in  high-dimensional regression settings.
\item any symmetric unimodal distribution about 0.
\item any unimodal distribution about 0.
\end{enumerate}
The latter two claims are related to characterizations of unimodal distributions due to \cite{khintchine1938unimodal} and  \cite{shepp1962symmetric}; see \cite{feller1971introduction}, p158. 
In other words, (ii) and (iii) provide fully non-parametric estimation for $g$ under the constraints that it is (ii) both unimodal and symmetric, or (iii) unimodal only.

Although our discussion above emphasizes the use of large $K$, in practice modest values of $K$ can provide reasonable performance. 
The key point is that the value of $K$ is not critical provided it is sufficiently large, and the grid of $\sigma_k$ or $a_k$ values suitably chosen.
See Implementation for details of our software defaults.

%Although it is natural to worry about ``overfitting" if $K$ is large,  in fact the unimodal constraint is sufficient to prevent serious problems with overfitting. For example, note that the unimodal constraint is sufficiently strong that there will exist a non-parametric maximum likelihood estimate (NPMLE) for $g$ under this constraint. (Indeed, if the $\beta_j$ are observed without error, corresponding to the standard errors $\shat_j \rightarrow 0$, this NPMLE has an explicit and elegant form; \cite{grenander1956theory}.)
%Thus allowing $K$ to be large in ii) or iii), and maximizing the likelihood with respect to $\pi$, will approximate the NPMLE for $g$, under the constraints of symmetry and unimodality (ii), or unimodality alone (iii).

\subsubsection*{Dependence of effects on standard errors}

The model (\ref{eqn:beta-alpha}) for general $\alpha$ can be fitted using the algorithm for $\alpha=0$. 
To see this, define $b_j:= \beta_j/ \shat_j^\alpha$, and $\hat{b} := \bhat_j/\shat_j^\alpha$. Then $\hat{b}_j$ is an estimate of $b_j$ with
 standard error $\shat'_j:=\shat_j^{1-\alpha}$. 
 Applying the algorithm for $\alpha=0$ to effect estimates $\hat{b}_1,\dots,\hat{b}_J$ with standard errors $\shat'_1,\dots,\shat'_J$
yields a posterior distribution $p(b_j | \shat_j, \hat{b}_j, \hat{\pi}, \alpha)$, which induces a posterior distribution on $\beta_j = b_j \shat_j^\alpha$.
%In the special case $\alpha=1$, $\hat{b}_j$ is the $t$ score $t_j$ and $\shat'{b}=1$, so this involves applying our algorithm to the $t$ scores
%treating them all as having the same standard error. In this case the local fdr from our method becomes monotonic in each direction as
%the $t$ score increases away from zero, just as with \qvalue, and (usually) \mixfdr and \locfdr.

%Different values of $\alpha$ in (\ref{eqn:beta-alpha}) essentially correspond to different models for
%how effect sizes $\beta_j$ scale with the standard error $\shat_j$. Since this scaling is unknown it may be desirable to estimate
%it from the data, and this can be done here by jointly maximizing the likelihood $p(\hat{b} | \shat,\pi,\alpha)$ over $\pi$ and $\alpha$
%rather than just $\pi$.  Although values of $\alpha$ other than 0 and 1 may not be easy interpret,
%and the actual relationship between $\beta_j$ and $\shat_j$ is unlikely to exactly follow (\ref{eqn:beta-alpha}) for any $\alpha$,
%this approach seems preferable in practice than simply assuming $\alpha=0$ or $\alpha=1$. We
%do not investigate this in detail here, but it has been implemented in our software by C.Dai, using a simple grid search over $\alpha$.

\subsubsection*{Replace normal likelihood with $t$ likelihood}

We generalize the normal likelihood (\ref{eqn:normlik}) by replacing it with a $t$ likelihood:
  \begin{equation} \label{eqn:tlik}
 \bhat_j \, | \, \beta_j, \shat_j \sim T_\nu(\beta_j, \shat_j)
 \end{equation}
where $T_\nu(\beta_j,\shat_j)$ denotes the distribution of $\beta_j+\shat_j T_\nu$ where $T_\nu$ has a standard $t$ distribution on $\nu$ degrees of freedom,
and  $\nu$ denotes the degrees of freedom used to estimate $\shat_j$ (assumed known, and for simplicity assumed to be the same for each $j$).
The normal approximation (\ref{eqn:normlik}) corresponds to the limit $\nu \rightarrow \infty$.
This generalization does not complicate inference when the mixture components $f_k$ in (\ref{eqn:g-general}) are uniforms; see Implementation below.
When the $f_k$ are normal the computations with a $t$ likelihood are considerably more difficult and we have not implemented this combination.

Equation (\ref{eqn:tlik}) is, of course, motivated by the standard asymptotic result
\begin{equation} \label{eqn:tdist}
(\bhat_j-\beta_j)/\shat_j \sim T_\nu.
\end{equation}
However (\ref{eqn:tdist}) does not imply (\ref{eqn:tlik}), because in (\ref{eqn:tdist}) $\shat_j$ is random whereas in (\ref{eqn:tlik}) it is conditioned on.
In principle it would be preferable, for a number of reasons, to model the randomness in $\shat_j$; we are
currently pursuing this improved approach (joint work with M.Lu) and results will be published elsewhere. 
%We note here that
%modelling the randomness in $\shat_j$ is particularly helpful when the number of observations used to estimate $\shat_j$ is very small (e.g. $<10$),
%where our simpler approach can produce unsatisfactory results.

\subsubsection*{Non-zero mode}

An addition to our software implementation, due to C.Dai, allows the mode to be estimated from the data by maximum likelihood, rather than fixed to 0.
This involves a simple grid search.

\subsection*{Implementation Details}

\subsubsection*{Likelihood for $\pi$}

We define the likelihood for $\pi$ to be the probability of the observed data $\bhat$ conditional on $\shat$: $L(\pi) := p(\bhat | \shat,\pi),$ which
by our conditional independence assumptions is equal to the product $\prod_j p(\bhat_j | \shat, \pi)$. [One might prefer to define the likelihood as $p(\bhat, \shat | \pi) = p(\bhat | \shat, \pi) p(\shat | \pi)$, in which case our definition comes down to assuming that the term $p(\shat | \pi)$ does not depend on $\pi$.]

Using the prior $\beta_j \sim \sum_{k=0}^K \pi_k f_k(\beta_j)$ given by (\ref{eqn:g-general}), and the normal likelihood (\ref{eqn:normlik}), integrating over $\beta_j$ yields
\begin{equation}
p(\bhat_j | \shat,\pi)  = \sum_{k=0}^K \pi_k \tilde{f}_k(\bhat_j)
\end{equation}
where 
\begin{equation}
\tilde{f}_k(\bhat_j) := \int f_k(\beta_j) N(\bhat_j; \beta_j,\shat^2_j) \, d\beta_j
\end{equation}
denotes the convolution of $f_k$ with a normal density.
These convolutions are straightforward to evaluate whether $f_k$ is a normal or uniform density.
Specifically, 
\begin{equation} \label{eqn:uconv}
\tilde{f}_k(\bhat_j)  = 
\begin{cases}
N(\bhat_j;0, \shat_j^2 +  \sigma_k^2) & \text{if $f_k(\cdot) = N(\cdot; 0, \sigma_k^2)$}, \\
\frac{\Psi((\bhat_j-a_k)/\shat_j) - \Psi((\bhat_j-b_k)/\shat_j)}{b_k-a_k} & \text{if $f_k(\cdot) = U(\cdot; a_k,b_k)$},
\end{cases}
\end{equation}
where $\Psi$ denotes the cumulative distribution function (c.d.f.) of the standard normal distribution.
If we replace the normal likelihood with the $t_\nu$ likelihood (\ref{eqn:tlik}) then the convolution 
for $f_k$ uniform the convolution is still given by (\ref{eqn:uconv}) but with $\Psi$ the c.d.f. 
of the $t_\nu$ distribution function. (The convolution for $f_k$ normal is tricky and we have not implemented it.)

\subsubsection*{Penalty term on $\pi$}

To make $\lfdr$ and $\lfsr$ estimates from our method ``conservative" we add a penalty 
term $log(h(\pi;\lambda))$ to the log-likelihood $\log L(\pi)$ to encourage over-estimation of $\pi_0$:
\begin{equation} \label{eqn:penalty}
h(\pi;\lambda) = \prod_{k=0}^K \pi_k^{\lambda_k-1}
\end{equation}
where $\lambda_k \geq 1 \, \forall k$. The default is $\lambda_0=10$ and $\lambda_k=1$, which yielded
consistently conservative estimation of $\pi_0$ in our simulations (Figure \ref{fig:pi0sims}).

Although this penalty is based on a Dirichlet density, we do not interpret this as a ``prior distribution" for $\pi$:
we chose it to provide conservative estimates of $\pi_0$ rather than to represent prior belief.

\subsubsection*{Problems with removing the penalty term in the half-uniform case}

It is straightforward to remove the penalty term by setting $\lambda_k=1$ in (\ref{eqn:penalty}).
We note here an unanticipated problem we came across when using no penalty term in the half-uniform case 
(i.e.~ $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$ in (\ref{eqn:g-general})): when the data
are nearly null, the estimated $g$ converges, as expected and desired, to a distribution where almost all the mass is near 0, but
sometimes all this mass is concentrated almost entirely just to one side (left or right) or 0. This can have a very profound effect on the local false sign rate:
for example, if all the mass is just to the right of 0 then all observations will be assigned a very high probability of being positive (but very small),
and a (misleading) low local false sign rate. For this reason we do not recommend use of the half-uniform with no penalty.

\subsubsection*{Optimization}

With this in place, the penalized log-likelihood for $\pi$ is given by:
\begin{equation}
\log L(\pi) + \log h(\pi) = \sum_{j=1}^n \log(\sum_{k=0}^K \pi_k l_{kj}) + \sum_{k=0}^K (\lambda_k-1) \log \pi_k
\end{equation}
where the $l_{kj}:= \tilde{f}_k(\bhat_j)$ are known. This is a convex optimization problem, which
can be solved very quickly and reliably using interior point (IP) methods. We used the {\tt KWdual} function from the {\tt R} package {\tt REBayes} \cite{REBayes}, which uses {\tt Rmosek} \cite{Rmosek}. We also found a simple EM algorithm \cite{dempster77}, accelerated using the elegant {\tt R} package {\tt SQUAREM} \cite{varadhan2008simple},
to provide adequate performance. In our EM implementation we initialized $\pi_k=1/n$ for $k=1,\dots,K$, with $\pi_0=1-\pi_1-\dots-\pi_K$,
and the one-step updates are:
\begin{align} \label{eqn:w}
w_{kj} & = \pi_k l_{kj} / \sum_{k'} {\pi_{k'} l_{k'j}} \\
n_k & = \sum_j w_{kj} + \lambda_k - 1 \quad \text{[E Step]} \\
\pi_k &= n_k/\sum_{k'} n_{k'} \quad \text{[M step]}.
\end{align}
One benefit to the EM algorithm is fewer software dependencies.
Both EM and IP methods are implemented in the {\tt ashr} package; results shown here are from the IP method, but graphs from EM are essentially the same.
See \url{http://stephenslab.github.io/ash/analysis/checkIP.html}  and \url{http://stephenslab.github.io/ash/analysis/IPvsEM.html}  for comparisons.

\subsubsection*{Conditional distributions}

Given $\hat\pi$, we compute the conditional distributions 
\begin{equation}
p(\beta_j | \hat\pi, \bhat, s) \propto g(\beta_j; \pi) L(\beta_j; \bhat_j, \shat_j).
\end{equation} 
Each posterior is a mixture on $K+1$ components:
\begin{equation}
p(\beta_j | \hat\pi, \bhat, s) = \sum_{k=0}^K w_{kj} p_k(\beta_j | \bhat_j, \shat_j)
\end{equation}
where the posterior weights $w_{kj}$ are computed as in (\ref{eqn:w}) with $\pi=\hat\pi$,
and the posterior mixture component $p_k$ is the posterior on $\beta_j$ that would be obtained using prior 
$f_k(\beta_j)$ and likelihood $L(\beta_j; \bhat_j,\shat_j)$.
All these posterior distributions are easily available.
For example, if $f_k$ is uniform and $L$ is $t_\nu$ then this is a truncated $t$ distribution.
If $f_k$ is normal and $L$ is normal, then this is a normal distribution.

\subsubsection*{Choice of grid for $\sigma_k, a_k$} \label{sec:grid}

\def\sigmamax{\sigma_{\text{max}}}
\def\sigmamin{\sigma_{\text{min}}}
When $f_k$ is $N(0,\sigma_k)$ we specify our grid by specifying: i) a maximum and minimum value $(\sigmamin,\sigmamax)$; ii) a multiplicative factor $m$ to be used
in going from one grid-point to the other, so that $\sigma_{k} = m \sigma_{k-1}$. The multiplicative factor affects the density of the grid; we used $m=\sqrt{2}$
as a default. We chose $\sigmamin$ to be small compared with the measurement precision ($\sigmamin=\min(\shat_j)/10$) and $\sigmamax= 2\sqrt{\max(\bhat_j^2-\shat_j^2)}$
based on the idea that $\sigmamax$ should be big enough so that  $\sigmamax^2 + \shat_j^2$ should exceed $\bhat_j^2$.   (In rare cases where $\max(\bhat_j^2-\shat_j^2)$
is negative we set $\sigmamax = 8\sigmamin$.)

When the mixture components $f_k$ are uniform, we use the same grid for the parameters $a_k$ as for $\sigma_k$ described above.

Our goal in specifying a grid was to make the limits sufficiently large and small, and the grid sufficiently dense, that results would not change appreciably with 
a larger or denser grid. For a specific data set one can of course check this by experimenting with the grid, but these defaults usually work well in our experience.

% Do NOT remove this, even if you are not including acknowledgments
\section*{Acknowledgements}

I thank J.~Lafferty for pointing out the convexity of the likelihood function, which lead to an improved implementation with interior point methods. 
Statistical analyses were conducted in the {\sf R} programming language \cite{Rcore:2012}, Figures produced using the {\tt ggplot2} package \cite{ggplot2}, and text
prepared using \LaTeX. Development of the methods in this paper was greatly enhanced by the use of the {\tt knitr} package \cite{xie2013dynamic}  within the RStudio GUI, and 
git and github. The {\tt ashr} R package is available from \url{http://github.com/stephens999/ashr} and includes contributions 
from Chaoxing (Rick) Dai, Mengyin Lu, Nan Xiao, and Tian Sen.

This work was supported by NIH grant HG02585 and a grant from the Gordon and Betty Moore Foundation.

%\section*{References}
% The bibtex filename
\bibliography{/Users/stephens/Dropbox/Documents/mainbib}

\clearpage
\appendix

\section{Appendix: Additional Figures and Tables}

\begin{figure}[h!] 
\begin{subfigure}{\textwidth}
\includegraphics[height=2in]{../analysis/figure/plot_cdf_eg.Rmd/mean_cdf-1.pdf}
\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here use penalty (\ref{eqn:penalty}) which leads to systematic overestimation of $\pi_0$ in some scenarios.}
\end{subfigure}
\begin{subfigure}{\textwidth}
\includegraphics[height=2in]{../analysis/figure/plot_cdf_eg.Rmd/mean_cdf_nopen-1.pdf}
\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here do not use penalty (\ref{eqn:penalty}) so $\pi_0$ is not systematically overestimated. Systematic differences from the truth in ``skew" and ``bimodal" scenarios highlight the effects of model mis-specification.}
\end{subfigure}
\caption{Comparisons of average estimated cdfs of $g$ with and without penalty term. See Figure \ref{fig:pi0sims} for simulation scenarios. In most cases the three different ash methods are very similar and so the lines lie on top of one another.} \label{fig:pen-nopen}
\end{figure}

\begin{figure}
\begin{center}
\begin{subfigure}{\textwidth}
\includegraphics[width=\textwidth]{../analysis/figure/plot_lfsr.Rmd/plot_lfsr_s_nn-1.png} 
\caption{Comparison of true and estimated $\lfsr$ when data are simulated with no point mass at zero ($\pi_0=0$), and also analyzed by \ash with no point mass on 0 (and mixture of normal components for $g$). Black line is $y=x$ and red line is $y=2x$. The results illustrate how estimates of $\lfsr$ can be more accurate in this case. That is, assuming there is no point mass can be beneficial if that is indeed true.}  \label{fig:lfsr-nn}
\end{subfigure}
\begin{subfigure}{\textwidth}
\includegraphics[width=\textwidth]{../analysis/figure/plot_lfsr.Rmd/plot_lfsr_s-1.png} 
\caption{Comparison of true and estimated $\lfsr$ when data are simulated with point mass at zero (drawn uniformly from [0,1] in each simulation), but analyzed by \ash with no point mass on 0 (and mixture of normal components for $g$). Black line is $y=x$ and red line is $y=2x$. The results illustrate how estimates of $\lfsr$ can be anti-conservative if we assume there is no point mass when the truth is that there is a point mass.} \label{fig:lfsr-s}
\end{subfigure}
\end{center}
\caption{Illustration of effects of excluding a point mass from the analysis.} \label{fig:lfsr-nopointmass}
\end{figure}

\begin{table}[!ht]
\centering\begin{tabular}{c c } \toprule
Scenario & Alternative distribution, $g_1$  \\ \midrule
spiky & $0.4 N(0,0.25^2) + 0.2 N(0,0.5^2) + 0.2 N(0,1^2), 0.2 N(0,2^2) $\\
near normal & $2/3 N(0,1^2) + 1/3 N(0,2^2)$ \\
flattop& $(1/7) [N(-1.5,.5^2) + N(-1,.5^2) + N(-.5,.5^2) +$ \\
 &  $N(0,.5^2) +N(0.5,.5^2)  +N(1.0,.5^2) + N(1.5,.5^2)]$  \\
skew & $(1/4) N(-2,2^2) + (1/4) N(-1,1.5^2) +  (1/3) N(0,1^2) + (1/6) N(1,1^2) $\\
big-normal & $N(0,4^2)$ \\ 
bimodal & $0.5 N(-2,1^2) + 0.5 N(2,1^2)$ \\ \bottomrule
\end{tabular}
\caption{Summary of simulation scenarios considered} \label{table:scenarios}
\end{table}

\begin{table}[!ht]
\begin{subtable}{\textwidth}
	\centering\input{../analysis/table/scoverage_all_nopen.tex}
	\caption{All observations}
\end{subtable}

\begin{subtable}{\textwidth}
\centering\input{../analysis/table/scoverage_neg_nopen.tex}
\caption{``Significant" negative discoveries.}
\end{subtable}

\begin{subtable}{\textwidth}
\centering\input{../analysis/table/scoverage_pos_nopen.tex}
\caption{``Significant" positive discoveries.}
\end{subtable}

\caption{Table of empirical coverage for nominal 95\% lower credible bounds for methods {\it without} the penalty term). } \label{tab:nopen}
\end{table}
%In simple settings (e.g. symmetric effects, with the same standard error) $\fdr_j$ is a monotonic function of the $p$ value $p_j$, and so
%\begin{equation}
%q_j = \widehat\FDR(\{k:p_k \leq p_j\})
%\end{equation}
%which is close to the definition in \cite{storey.03}.
\section*{Figure Legends}

\clearpage

\section*{Tables}

\section*{Supporting Information Legends}

Supplementary material can be found in {\bf Supplementary Information S1.}
%\begin{table}[!ht]
%\caption{
%\bf{Table title}}
%\begin{tabular}{|c|c|c|}
%table information
%\end{tabular}
%\begin{flushleft}Table caption
%\end{flushleft}
%\label{tab:label}
% \end{table}

\end{document}