### 2016-09-28 analysis/estimate_mode.Rmd

---
title: "Estimate Non-zero mode"
author: "Matthew Stephens"
date: 2016-02-01
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

The purpose of this analysis is to check the performance of the non-zero mode option.

## Simple simulation

I conjecture that the mean of the posterior means should be close to the optimum for the mode...
maybe even equal to it. (That is, this would constitute a fixed point for the update. We aren't
explicitly using that directly in the current implementation; the uniform mixture uses optim to
do it numerically; the normal mixture uses a true EM I think...)

```{r}
library(ashr)
check_mode = function(betahat, sebetahat,mixcompdist){
    z.ash=ash(betahat,sebetahat,mixcompdist=mixcompdist,mode="estimate")
    average.posteriormean = mean(get_pm(z.ash))
    fitted.mode = comp_mean(get_fitted_g(z.ash))[1]

    z.ash1 = ash(betahat-fitted.mode,sebetahat, mixcompdist=mixcompdist) #refit to get g
    g = get_fitted_g(z.ash1)
    loglik = get_loglik(z.ash1)

    loglik.down = ash(z-fitted.mode-0.01,1,g=g)$loglik
    loglik.up  = ash(z-fitted.mode+0.01,1,g=g)$loglik
    loglik.posteriormean = ash(z-average.posteriormean,1,g=g)$loglik

    return(list(fitted.mode=fitted.mode, average.posteriormean =average.posteriormean,
         loglik = c(loglik,loglik.down,loglik.up,loglik.posteriormean)))
  }
  set.seed(100)
  z = rnorm(1000)+3
  check_mode(z,1,mixcompdist="uniform")
  check_mode(z,1,mixcompdist="normal")
  check_mode(z,1,mixcompdist="halfuniform")
```

```{r}
  set.seed(100)
  beta = rexp(1000)
  betahat = beta + rnorm(1000,0,0.1)
  z.ash.hu = ash(betahat,0.1,mixcompdist="halfuniform",outputlevel=4,method="shrink")
  z.ash.pu = ash(betahat,0.1,mixcompdist="+uniform",outputlevel=4,method="shrink")

  z.ash.hu2 = ash(betahat-0.2,0.1,mixcompdist="halfuniform",outputlevel=4,method="shrink")
  z.ash.pu2 = ash(betahat-0.1,0.1,mixcompdist="+uniform",outputlevel=4,method="shrink")

  z.ash.hu.nzm = ash(betahat,0.1,mixcompdist="halfuniform",mode="estimate",method="shrink")
  z.ash.hu$loglik
  z.ash.hu.nzm$loglik

#  check_mode(betahat,.1,mixcompdist="halfuniform")
```

Note to self: check the normal version works too; check nonzeromodeEMobj  as it doesn't seem needed.

## Session information

```{r info}
sessionInfo()
```