### 2016-09-28 analysis/makefig_FDReg.Rmd

---
title: "Some figures to illustrate basic ideas in FDR"
author: "Matthew Stephens"
date: 2015-10-26
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

This file makes two figures:

1. illustrates basic idea of way q value does FDR analysis in p value space
2. compares methods in the way they decompose p values and z scores into two groups

```{r}
require(ashr)
require(qvalue)
require(locfdr)
require(mixfdr)
source("../R/plot_FDReg_hist.R") #some plotting functions
source("../R/nullalthist.R")

############################ FDReg_hist.pdf ##################################################################

#simple simulated example
ncz = 100 # number of bins in z score histograms
ntest = 10000
set.seed(111)

#simulate (with all tests alternative, true effects $\beta \sim N(0,1)$)
beta = rnorm(ntest)
sebetahat = 1
betahat = beta + rnorm(ntest, 0, sebetahat)
zscore = betahat/sebetahat
pval = pchisq(zscore^2,df=1,lower.tail=F)

#apply the different methods
res.qvalue = qvalue(pval)
res.locfdr = locfdr(zscore,nulltype=0,plot=0)
res.mixfdr = mixFdr(zscore,noiseSD=1,theonull=TRUE,plot=FALSE)
res.ash = ash(betahat,1,method="fdr")

#roughly compute a local fdr for qvalue (to aid the plotting of the decomposition in zscore space)
temp=hist(pval,nclass=50)
bin_fdr = res.qvalue$pi0/temp$density #compute local fdr in each bin of histogram
qval_fdr = bin_fdr[as.numeric(cut(pval,temp$breaks))]
qval_fdr = pmin(1,qval_fdr) #threshold at 1, so fdr never bigger than 1
qval_fdr = qvalue::lfdr(pval)
```

Here's an example of how qvalue decomposes $p$ values into null and alternative components. First the $p$ values:
```{r}
#pdf("../figures/FDReg_hist.pdf",width=6,height=3)
#par(mai=c(0.5,0.5,0.2,0.2),mgp = c(3, 0.5, 0))
#layout(matrix(1:2,ncol=2,byrow=TRUE))
plot_FDReg_hist(pval,res.qvalue$pi0,type=1,title="p values",cex.axis=0.8,cex.main=0.8)

```

Now the decomposition:
```{r}
plot_FDReg_hist(pval,res.qvalue$pi0,type=4,cex.axis=0.8,yaxt='n',textsize=0.9,cex.main=0.8,title="Decomposition into null/alternative")
axis(side=2, labels=FALSE,tck=-0.01)
```

############################## decomp_ZA.pdf #############################################
```{r decomp_ZA, dev='pdf', fig.width=6, fig.height=6, crop=TRUE}
#pdf("../figures/decomp_ZA.pdf",width=6,height=6)
layout(matrix(1:12,ncol=3,byrow=FALSE))
plotlabel=function(label,cex=1.5){
  plot(0,0,type="n",axes=FALSE,xlab="",ylab="")
  text(0,0,label,cex=cex)
}
plotlabel("qvalue")
plotlabel("locfdr")
plotlabel("mixfdr")
plotlabel("ash")

par(mar = c(1.3,2,1.5,0.2),mgp = c(3, 0.5, 0))

#p value histograms
altnullhist(pval,qval_fdr,main="p values",ncz=50,xaxt='n',cex.axis=0.8)
#plot_FDReg_hist(pval,res.qvalue$pi0,type=2,title="p values",xaxt='n',cex.axis=0.8)
axis(side=1, labels=FALSE,tck=-0.01)
#mtext(side=3,"p values",line=1)
altnullhist(pval,res.locfdr$fdr,main="",ncz=50,xaxt='n',cex.axis=0.8)
axis(side=1, labels=FALSE,tck=-0.01)
altnullhist(pval,res.mixfdr$fdr,main="",ncz=50,xaxt='n',cex.axis=0.8)
axis(side=1, labels=FALSE,tck=-0.01)
altnullhist(pval,get_lfdr(res.ash),main="",ncz=50,xaxt='n',cex.axis=0.8)
axis(side=1, labels=TRUE,tck=-0.01,cex.axis=0.8)
#mtext(side=1,"p values",line=2)

#z score histograms
nullalthist(zscore,qval_fdr,main="z scores",ncz=50,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=FALSE,tck=-0.01)
#mtext(side=3,"z scores",line=1)
nullalthist(zscore,res.locfdr$fdr,main="",ncz=50,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=FALSE,tck=-0.01)
nullalthist(zscore,res.mixfdr$fdr,main="",ncz=50,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=FALSE,tck=-0.01)
nullalthist(zscore,get_lfdr(res.ash),main="",ncz=50,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=TRUE,tck=-0.01,cex.axis=0.8)
#mtext(side=1,"z scores",line=2)

#dev.off()

```

This one is a different layout (4 columns, 2 rows) for my Tukey poster

```{r decomp_ZA_poster, dev='pdf', fig.width=6, fig.height=2, crop=TRUE}
#pdf("../figures/decomp_ZA_poster.pdf",width=6,height=2)
layout(matrix(1:8,ncol=4,byrow=TRUE))
par(mar = c(1.3,2,1.5,0.2),mgp = c(3, 0.5, 0))
ncz=25

#p value histograms
altnullhist(pval,qval_fdr,main="p values: qvalue",ncz=ncz,xaxt='n',cex.axis=0.8)
#plot_FDReg_hist(pval,res.qvalue$pi0,type=2,title="p values",xaxt='n',cex.axis=0.8)
axis(side=1, labels=FALSE,tck=-0.01)
#mtext(side=3,"p values",line=1)
altnullhist(pval,res.locfdr$fdr,main="locfdr",ncz=ncz,xaxt='n',cex.axis=0.8)
axis(side=1, labels=FALSE,tck=-0.01)
altnullhist(pval,res.mixfdr$fdr,main="mixfdr",ncz=ncz,xaxt='n',cex.axis=0.8)
axis(side=1, labels=FALSE,tck=-0.01)
altnullhist(pval,get_lfdr(res.ash),main="ash",ncz=ncz,xaxt='n',cex.axis=0.8)
axis(side=1, labels=TRUE,tck=-0.01,cex.axis=0.8)
#mtext(side=1,"p values",line=2)

#z score histograms
nullalthist(zscore,qval_fdr,main="z scores: qvalue",ncz=ncz,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=FALSE,tck=-0.01)
#mtext(side=3,"z scores",line=1)
nullalthist(zscore,res.locfdr$fdr,main="locfdr",ncz=ncz,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=FALSE,tck=-0.01)
nullalthist(zscore,res.mixfdr$fdr,main="mixfdr",ncz=ncz,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=FALSE,tck=-0.01)
nullalthist(zscore,get_lfdr(res.ash),main="ash",ncz=ncz,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=TRUE,tck=-0.01,cex.axis=0.8)
#mtext(side=1,"z scores",line=2)

dev.off()
```

## Session information

```{r info}
sessionInfo()
```