### 2016-09-29 analysis/adaptive_shrinkage.Rmd
Working through adaptive_shrinkage.Rmd.

---
title: "Illustration of Adaptive Shrinkage"
author: "Matthew Stephens"
date: 2015-10-26
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

The goal here is to illustrate the "adaptive" nature of the adaptive
shrinkage. The shrinkage is adaptive in two senses. First, the amount
of shrinkage depends on the distribution $g$ of the true effects,
which is learned from the data: when $g$ is very peaked about zero
then ash learns this and deduces that signals should be more strongly
shrunk towards zero than when g is less peaked about zero.  Second,
the amount of shrinkage of each observation depends on its standard
error: the smaller the standard error, the more informative the data,
and so the less shrinkage that occurs. From an Empirical Bayesian
perspective both of these points are entirely natural: the posterior
depends on both the prior and the likelihood; the prior, $g$, is
learned from the data, and the likelihood incorporates the standard
error of each observation.

TO DO: explain what this chunk does.

```{r chunk-options,include = FALSE}
source("chunk-options.R")
```

We simulate from two scenarios, the first with effects more peaked
about zero, the second with effects less peaked.

```{r initialize}
set.seed(100)
library(ashr)
library(ggplot2)
library(scales)
source("../code/dsc-shrink/datamakers/datamaker.R")

NSAMP=1000
s = 1/rgamma(NSAMP,5,5)
sim.bignormal = rnormmix_datamaker(args=list(g=normalmix(c(1),c(0),c(4)),
                        min_pi0=0,
                        max_pi0=0,
                        nsamp=NSAMP,
                        betahatsd=s
                      ))

sim.spiky = rnormmix_datamaker(args=list(g=normalmix(c(.4,.2,.2,.2),c(0,0,0,0),c(.25,.5,1,2)),
                        min_pi0=0,
                        max_pi0=0,
                        nsamp=NSAMP,
                        betahatsd=s
                      ))

```

Now run ash on both:
```{r run-ash}
beta.bignormal.ash = ash(sim.bignormal$input$betahat, s)
beta.spiky.ash = ash(sim.spiky$input$betahat, s)
```

Here we plot the shrunken estimates against the observed values, colored
according to the (square root of) precision: precise esitmates being colored red, and less precise estimates being blue. Two key features of the plots
illustrate the ideas of adaptive shrinkage: i) the estimates under the spiky scenario are shrunk more strongly, illustrating that shrinkage adapts to the underlying distribution of beta; ii) in both cases, estimates with large standard error (blue) are shrunk more than estimates with small standard error (red) illustrating that shrinkage adapts to measurement precision.

```{r plotting}
make_df_for_ashplot = function(sim1,sim2,ash1,ash2,name1="spiky",name2="big-normal"){
  df = data.frame(betahat = c(sim1$input$betahat,sim2$input$betahat),
                beta_est = c(get_pm(ash1), get_pm(ash2)),
                lfsr = c(get_lfsr(ash1), get_lfsr(ash2)),
                s=c(sim1$input$sebetahat,sim2$input$sebetahat),
                scenario=c(rep(name1,NSAMP),rep(name2,NSAMP)))
  df$signif = df$lfsr<0.05
  return(df)
}

ashplot=function(df,xlab="Observed beta-hat",ylab="Shrunken beta estimate"){
  ggplot(df,aes(x=betahat,y=beta_est,color=1/s)) + 
    xlab(xlab) +
    ylab(ylab) + geom_point() +  facet_grid(.~scenario) +   geom_abline(intercept=0,slope=1) +
  scale_colour_gradient2(midpoint = median(1/s),low="blue", mid="white",high="red",space="Lab")  +coord_fixed(ratio=1)
}
df=make_df_for_ashplot(sim.spiky,sim.bignormal,beta.spiky.ash,beta.bignormal.ash)
ashplot(df)
```

Now plot lfsr against z scores, colored
according to the (square root of) precision

```{r plotting lfsr}

z_lfsr_plot=function(df,ylab="Observed Z score",xlab="lfsr"){
  ggplot(df,aes(x=lfsr,y=betahat/s,color=1/s)) + 
    xlab(xlab) +
    ylab(ylab) + geom_point() +  facet_grid(.~scenario) + scale_colour_gradient2(midpoint = median(1/s),low="blue", mid="white",high="red",space="Lab") 
}

z_lfsr_plot(df)
```

A related consequence is that significance of each observation is no longer monotonic with $p$ value.

```{r}
pval_plot=function(df){
  ggplot(df,aes(x=pnorm(-abs(betahat/s)),y=lfsr,color=log(s))) + geom_point() +  facet_grid(.~scenario) +xlim(c(0, 0.025)) + xlab("p value") + ylab("lfsr") +
  scale_colour_gradient2(midpoint = 0,low="blue", mid="white",high="red")  
}
pval_plot(df)
```

Let's see how these are affected by changing the modelling assumptions so that
the standardized beta are exchangeable (rather than the beta being exchangeable).
```{r run-ash-ET}
beta.bignormal.ash.ET = ash(sim.bignormal$input$betahat, s,alpha=1,mixcompdist="normal")
beta.spiky.ash.ET = ash(sim.spiky$input$betahat, s,alpha=1,mixcompdist="normal")
df.ET = make_df_for_ashplot(sim.spiky,sim.bignormal,beta.spiky.ash.ET,beta.bignormal.ash.ET)
ashplot(df.ET,ylab="Shrunken beta estimate (ET model)")
pval_plot(df.ET)
```

This is a "volcano plot" showing effect size against p value. The blue points are "significant" in that they have lfsr<0.05. In this case the significance by lfsr is not quite the same as cutting off at a given $p$ value (you can see that the decision boundary is not quite the same as drawing a horizontal line), but also not that different, presumably because the standard errors, although varying across observations, do not vary greatly.
```{r volcano}
  ggplot(df,aes(x=betahat,y=-log10(2*pnorm(-abs(betahat/s))),col=signif)) + geom_point(alpha=1,size=1.75) +  facet_grid(.~scenario) + theme(legend.position = "none") +
  xlim(c(-10, 10)) + ylim(c(0, 15)) +
  xlab("Effect (beta)") + ylab("-log10 p-value")
```

```{r sehist}
hist(s,main="histogram of standard errors")
summary(s)
```

## Session information

```{r info}
sessionInfo()
```