### 2016-10-14 analysis/estimate_mode.Rmd

---
title: "Estimate Non-zero mode"
author: "Matthew Stephens"
date: 2016-02-01
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

First, we load the necessary libraries, and specify settings for
displaying the plots in the rendered document.

```{r chunk-options, include=FALSE}
library(ashr)
source("chunk-options.R")
```

The purpose of this analysis is to check the performance of the non-zero mode
option.

## Simple simulation

I conjecture that the mean of the posterior means should be close to the
optimum for the mode... maybe even equal to it. (That is, this would constitute
a fixed point for the update. We aren't explicitly using that directly in the
current implementation; the uniform mixture uses `optim` to do it numerically;
the normal mixture uses a true EM I think...)

```{r}
check_mode = function(betahat, sebetahat, mixcompdist) {
  z.ash = ash(betahat,sebetahat,mixcompdist = mixcompdist,mode = "estimate")
  average.posteriormean = mean(get_pm(z.ash))
  fitted.mode = comp_mean(get_fitted_g(z.ash))[1]

  # Refit to get g.
  z.ash1 = ash(betahat - fitted.mode,sebetahat,mixcompdist = mixcompdist)
  g      = get_fitted_g(z.ash1)
  loglik = get_loglik(z.ash1)

  loglik.down          = ash(z - fitted.mode - 0.01,1,g = g)$loglik
  loglik.up            = ash(z - fitted.mode + 0.01,1,g = g)$loglik
  loglik.posteriormean = ash(z - average.posteriormean,1,g = g)$loglik

  return(list(fitted.mode = fitted.mode,
              average.posteriormean = average.posteriormean,
              loglik = c(loglik,loglik.down,loglik.up,loglik.posteriormean)))
}

set.seed(100)
z = rnorm(1000) + 3
print(check_mode(z,1,mixcompdist="uniform"))
print(check_mode(z,1,mixcompdist="normal"))
print(check_mode(z,1,mixcompdist="halfuniform"))
```

```{r}
  set.seed(100)
  beta = rexp(1000)
  betahat = beta + rnorm(1000,0,0.1)
  z.ash.hu = ash(betahat,0.1,mixcompdist="halfuniform",outputlevel=4,method="shrink")
  z.ash.pu = ash(betahat,0.1,mixcompdist="+uniform",outputlevel=4,method="shrink")

  z.ash.hu2 = ash(betahat-0.2,0.1,mixcompdist="halfuniform",outputlevel=4,method="shrink")
  z.ash.pu2 = ash(betahat-0.1,0.1,mixcompdist="+uniform",outputlevel=4,method="shrink")

  z.ash.hu.nzm = ash(betahat,0.1,mixcompdist="halfuniform",mode="estimate",method="shrink")
  z.ash.hu$loglik
  z.ash.hu.nzm$loglik

#  check_mode(betahat,.1,mixcompdist="halfuniform")
```

Note to self: check the normal version works too; check nonzeromodeEMobj  as it doesn't seem needed.

## Session information.

```{r info}
sessionInfo()
```