### 2016-10-20 analysis/ash_npmle.Rmd
Revised ash_nplme.Rmd.

---
title: "Using ash for NPMLE"
author: "Matthew Stephens"
date: 2016-05-31
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

First, we load the necessary libraries, and specify settings for
displaying the plots in the rendered document.

```{r chunk_options, include=FALSE}
library(REBayes)
library(ashr)
source("chunk-options.R")
```

## The NPMLE

Here we show how we can use ash to (approximately) compute the unconstrained
NPMLE - that is, estimate the underlying distribution $g$ by maximizing the
likelihood without the unimodal constraint. See Koenker and Mizera, JASA 2014,
for background.

Within ash one can approximate the npmle as a mixture of uniforms on a dense
grid of non-overlapping values - this results in a piecewise constant density
with changes in the density only at the grid points.

The following example comes from the REBayes vignette by Koenker and Gu. The
underlying $g$ is a mixture of a point mass (weight 0.8) at 0 and a point
mass (weight 0.2) at 2.
```{r fit_glmix_model}
set.seed(102)
y <- c(rep(0,800),rnorm(200,2)) + rnorm(1000)
z <- GLmix(y)
```

Now we fit the NPMLE using 'ashr' and compare with the REBayes solution.
```{r fit_ash_model}
grid = seq(from = min(z$x),to = max(z$x),length = 1000)
k    = length(grid)
y.ash.npmle =
  ash(y,1,g = unimix(pi=rep(1/(k-1),(k-1)),a = grid[-k],b = grid[-1]),
      method = "shrink")

# Plot the model fitted using GLmix against the fitted ash model.
plot(z$x,cumsum(z$y)/sum(z$y),col = 2,main = "Estimated cdf",type = "l")
lines(cdf.ash(y.ash.npmle,x = z$x),type = "l",lty = "dashed",lwd = 2)

# Compare the likelihood given the GLmix model and given ash model.
print(z$logLik,digits = 12)
print(y.ash.npmle$loglik,digits = 12)
```

## Session information

```{r info}
sessionInfo()
```