### 2016-10-21 analysis/efron.fcr.Rmd

---
title: "Post-selection CI example, assymetric, not unimodal at 0"
author: "Matthew Stephens"
date: 2016-05-09
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

First, we load the necessary libraries, and specify settings for
displaying the plots in the rendered document.

```{r chunk_options, include=FALSE}
library(ashr)
library(qvalue)
library(locfdr)
source("chunk-options.R")
```

## Introduction

This example comes from Efron (2008) p16 when examining the false coverage rate
(FCR). I selected this example because the distribution of the non-zero effects
is highly assymetric and not at all unimodal at zero, both issues a referee
asked me to elaborate on. Specifically, the distribution of the non-zero
effects is N(-3,1). Here I simulate data, and apply ash (with the "halfuniform"
option to allow for asymmetric g).

```{r fit_halfuniform_model}
set.seed(10)
nsamp   = 1e4
altmean = -3
mu0     = rep(0,nsamp)
mu1     = rnorm(nsamp,altmean,1)
comp    = rbinom(nsamp,1,0.1)
mu      = ifelse(comp == 0,mu0,mu1)
z       = rnorm(nsamp,mu,1)

# Fit the model.
res.ash = ash(z,1,mixcompdist = "halfuniform")
```

We can also run ash with the "true" g, to allow us to compare the lfsr, lfdr,
etc.

```{r fit_normalmix_model}
true.g       = normalmix(c(0.9,0.1),c(0,-3),c(0,1))
res.ash.true = ash(z,1,g = true.g,fixg = TRUE)
```

Here we can see how the partition of $z$ scores compares with the truth. Note
the effect of the unimodal assumption is to extend the inferred alternative
distribution toward 0. Here, `nullalthist` plots a histogram of z scores,
highlighting the alternative distribution of z scores that is implied by
the localfdr values lfdr.

```{r plot_zscore_hist}
source("../R/nullalthist.R")
par(mfcol = c(2,1),cex.main = 1,font.main = 1)
nullalthist(z,lfdr = get_lfdr(res.ash.true),main = "true partition (res.ash.true)")
nullalthist(z,lfdr = get_lfdr(res.ash),main = "inferred partition (res.ash)")
```

Comparing the inferred posterior means, lfdr, and lfsr with the true values of
these quantities, we find reassuringly good correspondence.

```{r scatterplots}
par(mfrow = c(1,3))
plot(get_pm(res.ash.true),get_pm(res.ash),xlab = "Truth",ylab = "ash.hu",
     main = "Posterior Mean (inferred vs truth)",pch = 20,cex = 0.6,
     xlim = c(-6,1),ylim = c(-6,1))
abline(a = 0,b = 1,col = "orangered",lwd = 2,lty = "dotted")

plot(get_lfdr(res.ash.true),get_lfdr(res.ash),xlab = "Truth", ylab = "ash.hu",
     main = "lfdr (inferred vs truth)",pch = 20,cex = 0.6,
     xlim = c(0,1),ylim = c(0,1))
abline(a = 0,b = 1,col = "orangered",lwd = 2,lty = "dotted")

plot(get_lfsr(res.ash.true),get_lfsr(res.ash),xlab = "Truth", ylab = "ash.hu",
     main = "lfsr (inferred vs truth)",pch = 20,cex = 0.6,
     xlim = c(0,1),ylim = c(0,1))
abline(a = 0,b = 1,col = "orangered",lwd = 2,lty = "dotted")
```

## Comparison with qvalue and locfdr

We can also run qvalue and locfdr. We see that locfdr perhaps performs a bit
better than ash for the decomposition here, but the estimated local fdrs are
pretty similar. Here qvalue does less well because of the asymmetry which it
didn't take account of.

```{r comparison_to_other_methods}
res.locfdr = locfdr(z,nulltype = 0)
res.qvalue = qvalue(p = pchisq(z^2,df = 1,lower.tail = FALSE))

par(mfrow = c(1,3))
plot(get_lfdr(res.ash.true),get_lfdr(res.ash),pch = 20,cex = 0.6,
     xlab = "Truth (lfdr)",ylab = "ash.hu",main = "ash.hu",
     xlim = c(0,1),ylim = c(0,1))
abline(a = 0,b = 1,col = "orangered",lty = "dotted",lwd = 2)

plot(get_lfdr(res.ash.true),res.locfdr$fdr,pch = 20,cex = 0.6,
     xlab = "Truth (lfdr)",ylab = "Estimate",main = "locfdr",
     xlim = c(0,1),ylim = c(0,1))
abline(a = 0,b = 1,lwd = 2,col = "orangered",lty = "dotted")

plot(get_lfdr(res.ash.true),res.qvalue$lfdr,pch = 20,cex = 0.6,
     xlab = "Truth (lfdr)",ylab = "Estimate",main = "qvalue",
     xlim = c(0,1),ylim = c(0,1))
abline(a = 0,b = 1,lwd = 2,col = "orangered",lty = "dotted")
```

## Uniform tail curtails Credible Intervals

The following plot compares the (symmetric-tail) 95% CIs from ash (red) for
the "significant" observations with Bayes rule (green), similar to Figure 8
from Efron. Note that the lower 97.5% point is pretty accurate, but the
upper 97.5% point is curtailed - presumably due, at least in part, to the
short tails of the uniform mixture.

```{r plot_credints}
CImatrix = ashci(res.ash,level = 0.95)
BayesComparePlot = function(CImatrix, altmean = -3,...) {
  plot(z,mu,xlim = c(-8,0),pch = 20,cex = 0.6,...)
  points(z,CImatrix[,1],col = "orangered",cex = 0.6,pch = 20)
  points(z,CImatrix[,2],col = "orangered",cex = 0.6,pch = 20)

  fdr   = 0.9*dnorm(z)/(0.9*dnorm(z) + 0.1*dnorm(z,altmean,sqrt(2)))
  o     = order(z)
  upper = ifelse(fdr[o] < 0.025,
                 (z[o] + altmean)/2 + qnorm(0.975 + fdr[o])/sqrt(2),0)
  lines(z[o],upper,col = "limegreen",lwd = 2)
  lines(z[o],(z[o] + altmean)/2 - qnorm(0.975)/sqrt(2),col = "limegreen",lwd = 2)
  abline(v = max(z[fdr < 0.05]))
}
par(mfrow = c(1,1))
BayesComparePlot(CImatrix,
  main = "CIs for highly asymmetric and non-unimodal-at-zero data")
```

## Variational version

Although not a focus of the paper, ash does have an option to do variational
inference for the mixture components (with a Dirichlet prior). In practice
this approach usually ends up spreading the posterior mass up more among the
mixture components. It seemed plausible that this might lead to slightly
less extreme tail behaviour than above (because the model will put a little
more weight on the uniforms with larger variance, which are essentially set
to zero in the above).

```{r vb_model_fitting}
res.ash.VB  = ash(z,1,mixcompdist = "halfuniform",optmethod = "mixVBEM")
CImatrix.VB = ashci(res.ash.VB,level = 0.95)
```

Again, we can compare results with Bayes rule:

```{r plot_vb_credints}
BayesComparePlot(CImatrix.VB,
  main = paste("CIs for highly asymmetric and non-unimodal-at-zero data\n",
              "Variational Version"))
```

## Session information

```{r info}
sessionInfo()
```