### 2016-10-21 analysis/estimate_mode.Rmd
Revised estimate_mode.Rmd.

---
title: "Estimate Non-zero mode"
author: "Matthew Stephens"
date: 2016-02-01
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

The purpose of this analysis is to check the performance of the non-zero mode
option.

First, we load the necessary libraries, and specify settings for
displaying the plots in the rendered document.

```{r chunk-options, include=FALSE}
library(ashr)
source("chunk-options.R")
```

## Simple simulation

I conjecture that the mean of the posterior means should be close to the
optimum for the mode... maybe even equal to it. (That is, this would constitute
a fixed point for the update. We aren't explicitly using that directly in the
current implementation; the uniform mixture uses `optim` to do it numerically;
the normal mixture uses a true EM I think...)

```{r evaluate_mode_estimates}
check_mode = function(betahat, sebetahat, mixcompdist) {
  z.ash = ash(betahat,sebetahat,mixcompdist = mixcompdist,mode = "estimate")
  average.posteriormean = mean(get_pm(z.ash))
  fitted.mode           = comp_mean(get_fitted_g(z.ash))[1]

  # Refit to get g.
  z.ash1 = ash(betahat - fitted.mode,sebetahat,mixcompdist = mixcompdist)
  g      = get_fitted_g(z.ash1)
  loglik = get_loglik(z.ash1)

  loglik.down          = ash(z - fitted.mode - 0.01,1,g = g)$loglik
  loglik.up            = ash(z - fitted.mode + 0.01,1,g = g)$loglik
  loglik.posteriormean = ash(z - average.posteriormean,1,g = g)$loglik

  return(list(fitted.mode = fitted.mode,
              average.posteriormean = average.posteriormean,
              loglik = c(loglik,loglik.down,loglik.up,loglik.posteriormean)))
}

set.seed(100)
z = rnorm(1000) + 3
print(check_mode(z,1,mixcompdist = "uniform"))
print(check_mode(z,1,mixcompdist = "normal"))
print(check_mode(z,1,mixcompdist = "halfuniform"))
```

An additional experiment:

```{r}
set.seed(100)
beta     = rexp(1000)
betahat  = beta + rnorm(1000,0,0.1)
z.ash.hu = ash(betahat,0.1,mixcompdist="halfuniform",outputlevel=4,method="shrink")
z.ash.pu = ash(betahat,0.1,mixcompdist="+uniform",outputlevel=4,method="shrink")

z.ash.hu2    = ash(betahat - 0.2,0.1,mixcompdist = "halfuniform",outputlevel = 4,
                   method = "shrink")
z.ash.pu2    = ash(betahat - 0.1,0.1,mixcompdist = "+uniform",outputlevel = 4,
                   method = "shrink")
z.ash.hu.nzm = ash(betahat,0.1,mixcompdist = "halfuniform",mode = "estimate",
                   method = "shrink")
print(z.ash.hu$loglik,digits = 8)
print(z.ash.hu.nzm$loglik,digits = 8)
```

*Note to self:* check that the normal version works too; check
nonzeromodeEMobj as it doesn't seem needed.

## Session information

```{r info}
sessionInfo()
```