### 2016-10-22 analysis/check_mixfdr_lownoise.Rmd
Revised check_mixfdr_lownoise.Rmd---not much to do there.

---
title: "Check mixfdr in low noise simulation"
author: "Matthew Stephens"
date: 2015-10-26
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

Here we check that mixfdr actually does OK at fitting the bimodal distribution
when noise is small compared with bimodality.

First, we load the necessary libraries, and specify settings for
displaying the plots in the rendered document.

```{r chunk_options, include=FALSE}
source("chunk-options.R")
```

Simulate bimodal data with well-separated peaks. The histogram shows the
multi-modal nature of the underlying g is clear.

```{r sim_data}
source("../code/dsc-shrink/datamakers/datamaker.R")
sim.bimodal =
  rnormmix_datamaker(args = list(g = normalmix(c(0.5,0.5),c(-5,5),c(1,1)),
                                 min_pi0   = 0.4,
                                 max_pi0   = 0.4,
                                 nsamp     = 1000,
                                 betahatsd = 1))
hist(sim.bimodal$input$betahat,nclass = 20,xlim = c(-10,10),prob = TRUE,
     main="histogram of simulated betahat values")
```

Now run mixfdr on the simulated data. Note that the fit captures the
multi-modal nature of g, although mixfdr continues to overestimate pi0.

```{r fit_models}
source("../code/dsc-shrink/methods/mixfdr.wrapper.R")
sim.bimodal.mixfdroutput =
  mixfdr.wrapper(sim.bimodal$input,args = list(theonull = TRUE))
g = mixfdr2fitted.g(sim.bimodal.mixfdroutput)$fitted.g
x = seq(-10,10,length = 100)
par(cex.main = 1,font.main = 1)
plot(ecdf(sim.bimodal$meta$beta),cex = 0.6,
     main = "CDF of true beta: empirical (black) and fitted (red)")
lines(x,mixcdf(g,x),lwd = 2,lty = "dotted",col = "orangered")
```

## Session information

```{r info}
sessionInfo()
```