### 2016-10-25 analysis/metaplot_examples.Rmd
add some plot functions for talks

---
title: "DDD plots"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Simulate Z scores

```{r}
library(rmeta)
library(ashr)
n=50
set.seed(1)
beta = c(rep(0,n/2),rnorm(n/2,0,2))
betahat = beta+rnorm(n)
beta.ash=ashr::ash(betahat,1)
o = order(betahat)
```

# Meta Plots

```{r unshrunk}
rmeta::metaplot(betahat[o],rep(1,n),ylab="",xlab="",xlim=c(-6,6),colors = meta.colors(box=c(rep("red",n/2),rep("black",n/2))[o]))
```

```{r shrunk}
rmeta::metaplot(get_pm(beta.ash)[o],get_psd(beta.ash)[o],ylab="",xlab="",xlim=c(-6,6),colors = meta.colors(box=c(rep("red",n/2),rep("black",n/2))[o]))
```