### 2016-12-06 analysis/makefig_FDReg.Rmd
Working on makefig_FDRreg.Rmd.

---
title: "Some figures to illustrate basic ideas in FDR"
author: "Matthew Stephens"
date: 2015-10-26
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

This file makes two figures:

1. Illustrates basic idea of way q value does FDR analysis in p value space.

2. Compares methods in the way they decompose p values and z scores into
two groups.

First, we load the necessary libraries, and specify settings for
displaying the plots in the rendered document.

```{r chunk-options, include=FALSE}
require(ashr)
require(qvalue)
require(locfdr)
require(mixfdr)
source("chunk-options.R")
source("../R/plot_FDReg_hist.R")
source("../R/nullalthist.R")
```

## Simple simulated example

Generates DReg_hist.pdf?

```{r qvalue_analysis}

# ncz is the number of bins in the z score histograms.
ncz   = 100
ntest = 10000
set.seed(111)

# Simulate (with all tests alternative, true effects $\beta \sim N(0,1)$).
beta      = rnorm(ntest)
sebetahat = 1
betahat   = beta + rnorm(ntest,0,sebetahat)
zscore    = betahat/sebetahat
pval      = pchisq(zscore^2,df = 1,lower.tail = FALSE)

# Apply the different methods.
res.qvalue = qvalue(pval)
res.locfdr = locfdr(zscore,nulltype = 0,plot = 0)
res.mixfdr = mixFdr(zscore,noiseSD = 1,theonull = TRUE,plot = FALSE)
res.ash    = ash(betahat,1,method = "fdr")

# Roughly compute a local fdr for qvalue (to aid the plotting of the
# decomposition in zscore space).
temp     = hist(pval,nclass = 50)
bin_fdr  = res.qvalue$pi0/temp$density # compute local fdr in each bin of histogram
qval_fdr = bin_fdr[as.numeric(cut(pval,temp$breaks))]
qval_fdr = pmin(1,qval_fdr) # threshold at 1, so fdr never bigger than 1
qval_fdr = qvalue::lfdr(pval)
```
Here's an example of how qvalue decomposes $p$ values into null and alternative components. First the $p$ values:

```{r plot_pvalues}
plot_FDReg_hist(pval,res.qvalue$pi0,type = 1,title = "p values",
                cex.axis = 0.8,cex.main = 0.8)

```

Now the decomposition:

```{r decompose_pvalues}
plot_FDReg_hist(pval,res.qvalue$pi0,type = 4,cex.axis = 0.8,yaxt = 'n',
                textsize = 0.9,cex.main = 0.8,
                title = "Decomposition into null/alternative")
axis(side = 2,labels = FALSE,tck = -0.01)
```

```{r decomp_ZA, dev='pdf', fig.width=6, fig.height=6, crop=TRUE}
layout(matrix(1:12,ncol = 3,byrow = FALSE))
plotlabel = function(label,cex = 1.5){
  plot(0,0,type = "n",axes = FALSE,xlab = "",ylab = "")
  text(0,0,label,cex = cex)
}
plotlabel("qvalue")
plotlabel("locfdr")
plotlabel("mixfdr")
plotlabel("ash")

par(mar = c(1.3,2,1.5,0.2),mgp = c(3, 0.5, 0))

# p value histograms
altnullhist(pval,qval_fdr,main="p values",ncz=50,xaxt='n',cex.axis=0.8)
axis(side=1, labels=FALSE,tck=-0.01)
altnullhist(pval,res.locfdr$fdr,main="",ncz=50,xaxt='n',cex.axis=0.8)
axis(side=1, labels=FALSE,tck=-0.01)
altnullhist(pval,res.mixfdr$fdr,main="",ncz=50,xaxt='n',cex.axis=0.8)
axis(side=1, labels=FALSE,tck=-0.01)
altnullhist(pval,get_lfdr(res.ash),main="",ncz=50,xaxt='n',cex.axis=0.8)
axis(side=1, labels=TRUE,tck=-0.01,cex.axis=0.8)

# z score histograms
nullalthist(zscore,qval_fdr,main="z scores",ncz=50,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=FALSE,tck=-0.01)
nullalthist(zscore,res.locfdr$fdr,main="",ncz=50,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=FALSE,tck=-0.01)
nullalthist(zscore,res.mixfdr$fdr,main="",ncz=50,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=FALSE,tck=-0.01)
nullalthist(zscore,get_lfdr(res.ash),main="",ncz=50,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=TRUE,tck=-0.01,cex.axis=0.8)
```

This one is a different layout (4 columns, 2 rows) for my Tukey poster

```{r decomp_ZA_poster, dev='pdf', fig.width=6, fig.height=2, crop=TRUE}
#pdf("../figures/decomp_ZA_poster.pdf",width=6,height=2)
layout(matrix(1:8,ncol=4,byrow=TRUE))
par(mar = c(1.3,2,1.5,0.2),mgp = c(3, 0.5, 0))
ncz=25

#p value histograms
altnullhist(pval,qval_fdr,main="p values: qvalue",ncz=ncz,xaxt='n',cex.axis=0.8)
#plot_FDReg_hist(pval,res.qvalue$pi0,type=2,title="p values",xaxt='n',cex.axis=0.8)
axis(side=1, labels=FALSE,tck=-0.01)
#mtext(side=3,"p values",line=1)
altnullhist(pval,res.locfdr$fdr,main="locfdr",ncz=ncz,xaxt='n',cex.axis=0.8)
axis(side=1, labels=FALSE,tck=-0.01)
altnullhist(pval,res.mixfdr$fdr,main="mixfdr",ncz=ncz,xaxt='n',cex.axis=0.8)
axis(side=1, labels=FALSE,tck=-0.01)
altnullhist(pval,get_lfdr(res.ash),main="ash",ncz=ncz,xaxt='n',cex.axis=0.8)
axis(side=1, labels=TRUE,tck=-0.01,cex.axis=0.8)
#mtext(side=1,"p values",line=2)

#z score histograms
nullalthist(zscore,qval_fdr,main="z scores: qvalue",ncz=ncz,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=FALSE,tck=-0.01)
#mtext(side=3,"z scores",line=1)
nullalthist(zscore,res.locfdr$fdr,main="locfdr",ncz=ncz,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=FALSE,tck=-0.01)
nullalthist(zscore,res.mixfdr$fdr,main="mixfdr",ncz=ncz,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=FALSE,tck=-0.01)
nullalthist(zscore,get_lfdr(res.ash),main="ash",ncz=ncz,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=TRUE,tck=-0.01,cex.axis=0.8)
#mtext(side=1,"z scores",line=2)

dev.off()
```

## Session information

```{r info}
sessionInfo()
```