### 2016-12-09 analysis/makefig_FDReg.Rmd
knit makefig_FDReg.Rmd (note that this is in RStudio 1.0.44).

---
title: "Some figures to illustrate basic ideas in FDR"
author: "Matthew Stephens"
date: 2015-10-26
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

This file makes two figures:

1. Illustrates basic idea of way q value does FDR analysis in p value space.

2. Compares methods in the way they decompose p values and z scores into
two groups.

First, we load the necessary libraries, and specify settings for
displaying the plots in the rendered document.

```{r chunk-options, include=FALSE}
library(ashr)
library(qvalue)
library(locfdr)
library(mixfdr)
source("chunk-options.R")
source("../R/plot_FDReg_hist.R")
source("../R/nullalthist.R")
```

## Simple simulated example

Generate a small data set, and run the different methods on this data set.

```{r qvalue_analysis}

# ncz is the number of bins in the z score histograms.
ncz   = 100
ntest = 10000
set.seed(111)

# Simulate (with all tests alternative, true effects $\beta \sim N(0,1)$).
beta      = rnorm(ntest)
sebetahat = 1
betahat   = beta + rnorm(ntest,0,sebetahat)
zscore    = betahat/sebetahat
pval      = pchisq(zscore^2,df = 1,lower.tail = FALSE)

# Apply the different methods.
res.qvalue = qvalue(pval)
res.locfdr = locfdr(zscore,nulltype = 0,plot = 0)
res.mixfdr = mixFdr(zscore,noiseSD = 1,theonull = TRUE,plot = FALSE)
res.ash    = ash(betahat,1,method = "fdr")

# Roughly compute a local fdr for qvalue (to aid the plotting of the
# decomposition in zscore space) in each bin of histogram. I set the 
# threshold at 1 so that fdr is never larger than 1.
temp     = hist(pval,nclass = 50)
bin_fdr  = res.qvalue$pi0/temp$density 
qval_fdr = bin_fdr[as.numeric(cut(pval,temp$breaks))]
qval_fdr = pmin(1,qval_fdr)
qval_fdr = qvalue::lfdr(pval)
```

Here's an example of how qvalue decomposes $p$ values into null and alternative components. First the $p$ values:

```{r plot_pvalues}
plot_FDReg_hist(pval,res.qvalue$pi0,type = 1,title = "p values",
                cex.axis = 0.8,cex.main = 0.8)
```

Now plot the decomposition:

```{r decompose_pvalues}
plot_FDReg_hist(pval,res.qvalue$pi0,type = 4,cex.axis = 0.8,yaxt = 'n',
                textsize = 0.9,cex.main = 0.8,
                title = "Decomposition into null/alternative")
axis(side = 2,labels = FALSE,tck = -0.01)
```

Show the distribution of the p-values and the z-scores, stratified by their
classification into the "null" and "alternative".

```{r decomp_ZA, dev='pdf', fig.width=6, fig.height=6, crop=TRUE}
layout(matrix(1:12,ncol = 3,byrow = FALSE))
plotlabel = function(label,cex = 1.5){
  plot(0,0,type = "n",axes = FALSE,xlab = "",ylab = "")
  text(0,0,label,cex = cex)
}
plotlabel("qvalue")
plotlabel("locfdr")
plotlabel("mixfdr")
plotlabel("ash")

par(mar = c(1.3,2,1.5,0.2),mgp = c(3, 0.5, 0))

# p value histograms
altnullhist(pval,qval_fdr,main = "p values",ncz = 50,xaxt = 'n',cex.axis = 0.8)
axis(side = 1, labels = FALSE,tck = -0.01)
altnullhist(pval,res.locfdr$fdr,main = "",ncz = 50,xaxt = 'n',cex.axis = 0.8)
axis(side = 1, labels = FALSE,tck = -0.01)
altnullhist(pval,res.mixfdr$fdr,main = "",ncz = 50,xaxt = 'n',cex.axis = 0.8)
axis(side = 1, labels = FALSE,tck = -0.01)
altnullhist(pval,get_lfdr(res.ash),main = "",ncz = 50,xaxt = 'n',cex.axis = 0.8)
axis(side = 1, labels = TRUE,tck = -0.01,cex.axis = 0.8)

# z score histograms
nullalthist(zscore,qval_fdr,main = "z scores",ncz = 50,xaxt = 'n',
            cex.axis = 0.8,ylim = c(0,0.3),xlim = c(-6,6))
axis(side = 1,labels = FALSE,tck = -0.01)
nullalthist(zscore,res.locfdr$fdr,main = "",ncz = 50,xaxt = 'n',
            cex.axis = 0.8,ylim = c(0,0.3),xlim = c(-6,6))
axis(side = 1,labels = FALSE,tck = -0.01)
nullalthist(zscore,res.mixfdr$fdr,main = "",ncz = 50,xaxt = 'n',
            cex.axis = 0.8,ylim = c(0,0.3),xlim = c(-6,6))
axis(side = 1,labels = FALSE,tck = -0.01)
nullalthist(zscore,get_lfdr(res.ash),main = "",ncz = 50,xaxt = 'n',
            cex.axis = 0.8,ylim = c(0,0.3),xlim = c(-6,6))
axis(side = 1,labels = TRUE,tck = -0.01,cex.axis = 0.8)
```

This one is a different layout (4 columns, 2 rows) for my Tukey poster

```{r decomp_ZA_poster, dev='pdf', fig.width=6, fig.height=2, crop=TRUE}
# pdf("../figures/decomp_ZA_poster.pdf",width=6,height=2)
layout(matrix(1:8,ncol = 4,byrow = TRUE))
par(mar = c(1.3,2,1.5,0.2),mgp = c(3, 0.5, 0))
ncz = 25

# p value histograms
altnullhist(pval,qval_fdr,main = "p values: qvalue",ncz = ncz,
            xaxt = 'n',cex.axis = 0.8)
# plot_FDReg_hist(pval,res.qvalue$pi0,type = 2,title = "p values", 
#                 xaxt = 'n',cex.axis = 0.8)
axis(side = 1,labels = FALSE,tck = -0.01)
# mtext(side = 3,"p values",line = 1)
altnullhist(pval,res.locfdr$fdr,main = "locfdr",ncz = ncz,xaxt = 'n',
            cex.axis = 0.8)
axis(side = 1,labels = FALSE,tck = -0.01)
altnullhist(pval,res.mixfdr$fdr,main = "mixfdr",ncz = ncz,xaxt = 'n',
            cex.axis = 0.8)
axis(side = 1,labels = FALSE,tck = -0.01)
altnullhist(pval,get_lfdr(res.ash),main = "ash",ncz = ncz,xaxt = 'n',
            cex.axis = 0.8)
axis(side = 1,labels = TRUE,tck = -0.01,cex.axis = 0.8)
# mtext(side = 1,"p values",line = 2)

#z score histograms
nullalthist(zscore,qval_fdr,main = "z scores: qvalue",ncz = ncz,xaxt = 'n',
            cex.axis = 0.8,ylim = c(0,0.3),xlim = c(-6,6))
axis(side = 1,labels = FALSE,tck = -0.01)
# mtext(side = 3,"z scores",line = 1)
nullalthist(zscore,res.locfdr$fdr,main = "locfdr",ncz = ncz,xaxt = 'n',
            cex.axis = 0.8,ylim = c(0,0.3),xlim = c(-6,6))
axis(side = 1,labels = FALSE,tck = -0.01)
nullalthist(zscore,res.mixfdr$fdr,main = "mixfdr",ncz = ncz,xaxt = 'n',
            cex.axis = 0.8,ylim = c(0,0.3),xlim = c(-6,6))
axis(side = 1,labels = FALSE,tck = -0.01)
nullalthist(zscore,get_lfdr(res.ash),main = "ash",ncz = ncz,xaxt = 'n',
            cex.axis = 0.8,ylim = c(0,0.3),xlim = c(-6,6))
axis(side = 1,labels = TRUE,tck = -0.01,cex.axis = 0.8)
# mtext(side = 1,"z scores",line = 2)
# dev.off()
```

## Session information

```{r info}
sessionInfo()
```