### 2016-12-20 analysis/checkIP.Rmd
Successfully knit summarize_dsc_znull.Rmd using simulation results from res.znull.RData.

---
title: "Check interior point method"
author: "Matthew Stephens"
date: 2015-11-18
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

Here we compare the likelihoods achieved by the newly implemented
interior point method against the EM algorithm.

First, we load the necessary libraries, and specify settings for
displaying the plots in the rendered document.

```{r chunk_options, include=FALSE}
library(ggplot2)
library(dplyr)
library(reshape2)
source("chunk-options.R")
```

## Log-likelihood comparison

The 'dsc-opt' script simply runs both methods (IP and EM) on the same data sets,
and compares the difference in the log-likelihoods. We can see that in >90% of 
cases the difference is negligible. Then there is a tail of cases where the IP 
log-likelihood is higher - presumably where the EM doesn't really converge so
well.

```{r compare_likelihoods}
load("../output/dsc-opt-files/dsc_opt.RData")
plot(ecdf(dsc_opt$res$diff1),
     main = "ECDF of log-likelihood difference (IP - EM)")
```

## lfsr comparison

However, what we really care about is inference quantities, such as the lfsr, 
rather than the log-likelihood. Here we compare the lfsr for the simulations 
where we ran both the IP method and EM method (nocxx) in dsc-shrink.

```{r compare_lfsr}
load("../output/dsc-shrink-files/res.RData")

# Select out the lfsr estimates for each method.
df1 = res$lfsr %>% filter(method %in% c("ash.hu.nocxx")) %>% 
        select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
        melt(id.vars = c("method","scenario","seed",".id"),
             value.name = "lfsr.nocxx")
df2 = res$lfsr %>% filter(method %in% c("ash.hu")) %>%
        select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
        melt(id.vars = c("method","scenario","seed",".id"),
             value.name = "lfsr.IP")
df = inner_join(df1,df2,by = c("scenario","seed","variable"))
ggplot(df,aes(lfsr.nocxx,lfsr.IP)) + geom_point(shape = 1) +
facet_wrap(~ scenario,nrow = 2) +
geom_abline(colour = "black") +
xlab("EM algorithm (no cxx)") +
ylab("IP method")
```

## Session information

```{r info}
sessionInfo()
```