### 2016-12-20 analysis/investigate_hu_badcoverage.Rmd

---
title: "Investigate poor coverage for ash.hu"
author: "Matthew Stephens"
date: 2015-10-26
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

First, we load the necessary libraries, and specify settings for
displaying the plots in the rendered document.

```{r chunk_options, include=FALSE}
library(ashr)
library(reshape2)
library(ggplot2)
library(dplyr)
source("chunk-options.R")
```

```{r}
load("../output/dsc-shrink-files/res.RData")

out <- res$negprob
neglong = 
  out %>% 
    select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
    melt(id.vars=c("method","scenario","seed",".id"),value.name = "negprob") %>%
    filter(negprob > 0.95)

poslong = 
  res$posprob %>% 
    select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
    melt(id.vars = c("method","scenario","seed",".id"),value.name="posprob") %>%
    filter(posprob > 0.95)

reslong = 
  res$cdf_score %>% 
    select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%    
    melt(id.vars = c("method","scenario","seed",".id"))

reslong.pos = inner_join(reslong,poslong)
reslong.neg = inner_join(reslong,neglong)

```

```{r}
ash.hu.ft = reslong.pos %>% dplyr::filter(method=="ash.hu.s" & scenario=="flat-top")
head(ash.hu.ft)
```

```{r}
  eg=load_example(dsc_shrink,42,"flat-top",method="ash.hu.s","../code/dsc-shrink")
  eg$output$fitted_g
```

Notice how almost all the inferred weight is on a small positive component. As a result false sign rate will be small, and there will be a strong tendency
to overestimate 0 effects. This leads to coverage problems observed.

Now let's look at an example with u and spiky which seems to be somewhat badly calibrated for the negative discoveries. 
```{r}
ash.u.s.spiky = reslong.neg %>% dplyr::filter(method=="ash.u.s" & scenario=="spiky")
ash.n.s.spiky = reslong.neg %>% dplyr::filter(method=="ash.n.s" & scenario=="spiky")
hist(ash.u.s.spiky$value,nclass=100,main="histogram of quantile where observation falls in its CI; significant negative discoveries")
hist(ash.n.s.spiky$value,nclass=100,main="histogram of quantile where observation falls in its CI; significant negative discoveries")
```

So what seems to be happening here is that the uniform tail is too short; when observation falls outside of this tail it gets a zero quantile of posterior interval.

Can we find an example?
```{r}
  table((ash.u.s.spiky %>% filter(value==0))$seed)
  eg =load_example(dsc_shrink,seed = 39,method="ash.u.s",scenario="spiky","../code/dsc-shrink")
```

For comparison, here are the positive discoveries; here there is not too bad.
```{r}
ash.u.s.spiky = reslong.pos %>% dplyr::filter(method=="ash.u.s" & scenario=="spiky")
ash.n.s.spiky = reslong.pos %>% dplyr::filter(method=="ash.n.s" & scenario=="spiky")
hist(ash.n.s.spiky$value)
hist(ash.u.s.spiky$value)
```

## Session information

```{r info}
sessionInfo()
```