### 2016-12-20 analysis/summarize_dsc_znull.Rmd

---
title: "Summarize distribution of LR statistic under null"
author: "Matthew Stephens"
date: '2015-10-26'
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

First, we load the necessary libraries, and specify settings for
displaying the plots in the rendered document.

```{r chunk_options, include=FALSE}
library(ggplot2)
library(dplyr)
source("chunk-options.R")
```

Load and summarize the results of the data simulations.

```{r load_sim_results}
load("../output/dsc-znull-files/res.znull.RData")
out <- ungroup(res %>% group_by(scenario,method) %>% 
                  summarise(gt0 = mean(logLR > 0)))
xtabs(gt0 ~ method + scenario,out)
```

Note that the ash.n.s conforms closely to the null expectation of 0.5 
chi2_0 + 0.5 chi2_1 from Stram and Lee, Biometrics.

*Conjecture:* the uniform has the same asymptotic behaviour and the half 
uniform is like the sum of two of these?

From this a 95% procedure would need to check for logLR > `r qchisq(0.9,df=1)/2`.

```{r scatterplots}
ggplot(res %>% filter(logLR > 0) %>% filter(grepl("ash.hu",method)),
       aes(sample = logLR)) + 
  facet_grid(scenario ~ method) +
  stat_qq(distribution = qchisq,dparams = list(df=1)) + 
  geom_abline() + 
  ggtitle("qqplot under null vs chisq-1; line slope=1")

ggplot(res %>% filter(logLR > 0) %>% filter(grepl("ash.u",method)),
       aes(sample = logLR)) + 
  facet_grid(scenario~method) +
  stat_qq(distribution = qchisq,dparams = list(df = 1)) + 
  geom_abline(intercept = 0,slope = 0.5) + 
  ggtitle("qqplot under null vs chisq-1; line slope=0.5")

ggplot(res %>% filter(logLR > 0) %>% filter(grepl("ash.n",method)),
       aes(sample = logLR)) + 
  facet_grid(scenario ~ method) +
  stat_qq(distribution = qchisq,dparams = list(df = 1)) + 
  geom_abline(intercept = 0,slope = 0.5) + 
  ggtitle("qqplot under null vs chisq-1; line slope=0.5")
```

## Session information

```{r info}
sessionInfo()
```