### 2016-12-20 analysis/plot_egdens.Rmd
Successfully re-knit plot_egdens.Rmd using ashr 2.0.4.

---
title: "Plot densities for simulation studies"
author: "Matthew Stephens"
date: 2015-10-26
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk_options, include=FALSE}
library(ashr)
library(ggplot2)
library(dplyr)
source("chunk-options.R")
```

Load the results of the simulation experiments, and generate the density
data for all the simulation scenarios in a single data frame ("df").

```{r chunk_label}
load("../output/dsc-shrink-files/res.RData")
PLOTSCENARIOS = c("spiky","near-normal","flat-top","skew","bimodal")
PLOTNAMES     = PLOTSCENARIOS

df = data.frame()
for(i in PLOTSCENARIOS) {
  s  = dsc_shrink$scenarios[[i]]
  g  = s$args$g
  x  = seq(-6,6,length = 100)
  y  = as.numeric(dens(g,x))
  df = rbind(df,data.frame(x = x,y = y,scenario = i))
}

df$scenario         = factor(df$scenario,levels = PLOTSCENARIOS)
levels(df$scenario) = PLOTNAMES
```

Generate density plots using ggplot.

```{r scenario_density, dev='pdf',fig.height=3, fig.width=9, crop=TRUE}
ggplot(df,aes(x = x,y = y)) + 
  geom_line(size = 1.2,linetype = 1) + 
  facet_grid(.~scenario) + 
  ylab("density")
```

## Session information

```{r info}
sessionInfo()
```