### 2017-01-02 analysis/checkIP.Rmd
Successfully rendered checkIP.Rmd using new dsc-opt and dsc-shrink output.

---
title: "Check interior point method"
author: "Matthew Stephens"
date: 2015-11-18
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

Here we compare the likelihoods achieved by the newly implemented
interior point method against the EM algorithm.

**Note:** the last code chunk ("compare_lfsr") may take several minutes to 
render in R or RStudio because it uses ggplot to plot more than a million data 
points. Also, it it may not be possible to run the last code chunk 
interactively in RStudio.

First, we load the necessary libraries, and specify settings for
displaying the plots in the rendered document.

```{r chunk_options, include=FALSE}
library(ggplot2)
library(dplyr)
library(reshape2)
source("chunk-options.R")
```

## Log-likelihood comparison

The 'dsc-opt' script simply runs both methods (IP and EM) on the same
data sets, and compares the difference in the log-likelihoods. We can
see that in >90% of cases the difference is negligible. Then there is
a tail of cases where the IP log-likelihood is higher - presumably
where the EM doesn't really converge so well.

```{r compare_likelihoods}
load("../output/dsc-opt-files/dsc_opt.RData")
plot(ecdf(dsc_opt$res$diff1),
     main = "ECDF of log-likelihood difference (IP - EM)")
```

## lfsr comparison

However, what we really care about is inference quantities, such as
the lfsr, rather than the log-likelihood. Here we compare the lfsr for
the simulations where we ran both the IP method and EM method (nocxx)
in dsc-shrink.

```{r compare_lfsr}
load("../output/dsc-shrink-files/res.RData")

# Select out the lfsr estimates for each method.
df1 = res$lfsr %>% filter(method %in% c("ash.hu.nocxx")) %>% 
        select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
        melt(id.vars = c("method","scenario","seed",".id"),
             value.name = "lfsr.nocxx")
df2 = res$lfsr %>% filter(method %in% c("ash.hu")) %>%
        select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
        melt(id.vars = c("method","scenario","seed",".id"),
             value.name = "lfsr.IP")
df = inner_join(df1,df2,by = c("scenario","seed","variable"))
df = transform(df,scenario = factor(scenario))
print(ggplot(df,aes(lfsr.nocxx,lfsr.IP)) + 
      geom_point(shape = 1) +
      facet_wrap(~ scenario,nrow = 2) +
      geom_abline(colour = "black") +
      xlab("EM algorithm (no cxx)") +
      ylab("IP method"))
```

## Session information

```{r info}
sessionInfo()
```