### 2017-01-03 analysis/plot_lfsr.Rmd

---
title: "Plot local false sign rate results"
author: "Matthew Stephens"
date: 2015-10-26
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

First, we load the necessary libraries, and specify settings for
displaying the plots in the rendered document.

```{r chunk_options, include=FALSE}
library(dplyr)
library(reshape2)
library(ggplot2)
source("chunk-options.R")
source("../R/set_plot_colors.R")
```

Load the results of the simulation experiments generated by 
code/dsc-shrink/run_dsc.R, and prepare the posterior statistics for 
plotting.

```{r chunk_label}
load("../output/dsc-shrink-files/res.RData")

#' @param df dataframe of scores for many methods/scenrios etc
#' @return tall dataframe with columns of scores for each method and the "goldmethod" against which plot is to be made
process_score_for_plotting_against_gold = 
  function(df, 
           PLOTMETHODS   = c("ash.n","ash.u","ash.hu"),
           GOLDMETHOD    = "bayes",PLOTSEEDS = 1:100,
           PLOTSCENARIOS = c("spiky","near-normal","flat-top","skew",
                             "big-normal","bimodal"),
           PLOTNAMES     = PLOTSCENARIOS) {
  df %<>% filter(seed %in% PLOTSEEDS) %>% 
    filter(scenario %in% PLOTSCENARIOS) %>% 
    filter(method %in% c(PLOTMETHODS,GOLDMETHOD))
  df$scenario = factor(df$scenario,levels = PLOTSCENARIOS)
  levels(df$scenario) = PLOTNAMES

  # Create "tall"" version of dataframe.
  df %<>% select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
    melt(id.vars = c("method","scenario","seed",".id"),value.name = "val")

  #separate bayes and remainder
  df.bayes = df %>% filter(method == GOLDMETHOD)
  df.rest  = df %>% filter(method != GOLDMETHOD)

  # Join bayes with others, so each line has both the bayes and the 
  # non-bayes version.
  return(inner_join(df.bayes,df.rest,by = c("scenario","seed","variable")))
}

plot_lfsr = function(lfsr,xlab = "True lfsr",ylab = "Estimated lfsr",
                     xlim = c(0,0.2),ylim = c(0,0.2),
                     legend.position = "bottom")
  ggplot(lfsr,aes(val.x,val.y,colour = method.y)) +
    facet_grid(. ~ scenario) +
    guides(alpha = FALSE) +
    geom_abline(colour = "black") +
    geom_abline(colour = "red",slope = 2) +
    xlab(xlab) + ylab(ylab) +
    geom_point(shape = 1,size = 0.1,alpha = 0.2) +
    scale_y_continuous(limits = ylim) +
    scale_x_continuous(limits = xlim)

lfsr = process_score_for_plotting_against_gold(res$lfsr,PLOTSEEDS = 1:100,
                                               PLOTMETHODS = "ash.n")
lfdr = process_score_for_plotting_against_gold(res$lfdr,PLOTSEEDS = 1:100,
                                               PLOTMETHODS = "ash.n")

p1 = plot_lfsr(lfsr,ylim = c(0,1),xlim = c(0,0.2))
p2 = plot_lfsr(lfdr,ylim = c(0,1),xlim = c(0,0.2),
               xlab = "True lfdr",ylab = "Estimated lfdr")
```

Separately for each model, create a scatterplot comparing the estimated LFSR 
against the "gold-standard" LFSR.

```{r plot_lfsr, dev=c('png','pdf'), fig.width=9, fig.height=3, crop=TRUE}
print(p1 + theme(legend.position = "none",
                 axis.text.x = element_text(size = 8,angle = 45)) +
      coord_equal(ratio = 1/5) + colScale)
```

Separately for each model, create a scatterplot comparing the estimated LFDR 
against the "gold-standard" LFDR.

```{r plot_lfdr, dev=c('png','pdf'), fig.width=9, fig.height=3, crop=TRUE}
print(p2 + theme(legend.position = "none",
                 axis.text.x = element_text(size = 8,angle = 45)) +
      coord_equal(ratio = 1/5) + colScale)
```

Separately for each of the ash.n.s methods, create a scatterplot comparing the 
estimated LFSR against the "gold-standard" LFSR.

```{r plot_lfsr_s, dev=c('png','pdf'), fig.width=9, fig.height=3, crop=TRUE}
lfsr.s = process_score_for_plotting_against_gold(res$lfsr,PLOTSEEDS = 1:100,
                                                 PLOTMETHODS = "ash.n.s")
p1.s   = plot_lfsr(lfsr.s,ylim = c(0,1),xlim = c(0,0.2))
print(p1.s + theme(legend.position = "none",
                   axis.text.x = element_text(size = 8,angle = 45)) +
      coord_equal(ratio = 1/5))
```

Separately for each of the ash.n.s methods with the -nn data sets, 
create a scatterplot comparing the estimated LFSR against the 
"gold-standard" LFSR.

```{r plot_lfsr_s_nn, dev=c('png','pdf'), fig.width=9, fig.height=3, crop=TRUE}
lfsr.s.nn = 
  process_score_for_plotting_against_gold(res$lfsr,PLOTSEEDS = 1:100,
    PLOTMETHODS = "ash.n.s",
    PLOTSCENARIOS = paste0(c("spiky","near-normal","flat-top","skew",
                             "big-normal","bimodal"),"-nn"))
p1.s.nn = plot_lfsr(lfsr.s.nn,ylim = c(0,1),xlim = c(0,0.2))
print(p1.s.nn + theme(legend.position = "none",
                      axis.text.x = element_text(size = 8,angle = 45)) +
      coord_equal(ratio = 1/5))
```

## Session information.

```{r info}
print(sessionInfo())
```