### 2017-01-03 analysis/plot_cdf_eg.Rmd
Successfully rendered plot_cdf_eg.Rmd, but only after dropping the 'big-normal' scenarios from the plots; it seems that the big-normal results are no longer included.

---
title: "Various examples of cdf plots"
author: "Matthew Stephens"
date: 2015-10-26
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

First, we load the necessary libraries, and specify settings for
displaying the plots in the rendered document.

```{r chunk_options, include=FALSE}
library(plyr)
library(dplyr)
library(magrittr)
library(ashr)
library(ggplot2)
library(dscr)
library(REBayes)
library(ashr)
source("chunk-options.R")
source("../code/dsc-shrink/methods/mixfdr.wrapper.R")
source("../R/set_plot_colors.R")
```

Define a couple functions that compile CDF data in such a way that it is 
easy to plot the CDFs using ggplot2.

```{r chunk_label}
load("../output/dsc-shrink-files/res.RData")

# df is a list with components method,seed, scenario
# cdf is evaluated at x
# returns list of x,y,pi0, where cdf values are in y
get_cdf = function (df,dsc = dsc_shrink,x = seq(-6,6,length = 100),
                    homedir = "../code/dsc-shrink") {
  m = df$method
  if (m == "truth")
    m = "ash.n"
  temp = load_example(dsc,df$seed,df$scenario,m,homedir)
  pi0  = temp$meta$pi0

  if (df$method == "truth") {
    s    = dsc$scenarios[[df$scenario]]
    galt = s$args$g
    g    = normalmix(c(pi0,(1 - pi0) * mixprop(galt)),
                     c(0,comp_mean(galt)),
                     c(0,comp_sd(galt)))
  } else {
    if (grepl("mixfdr",df$method)) {
      temp$output$fitted_g = mixfdr2fitted.g(temp$output)$fitted.g
    }
    g = temp$output$fitted_g
  }
  return(data.frame(x = x,y = as.numeric(mixcdf(g,x)),pi0 = pi0))
}

plot_mean_cdf = 
  function(SEEDS,
           PLOTMETHODS   = c("ash.n","ash.u","ash.hu","truth","mixfdr.tnull"),
           PLOTSCENARIOS = c("spiky","near-normal","flat-top","skew","bimodal"),
           pi0filter     = FALSE,...) {
  PLOTNAMES = PLOTSCENARIOS

  #set up dataframe with cdf for all methods and all datasets
  df     = expand.grid(seed = SEEDS,scenario = PLOTSCENARIOS,
                       method = PLOTMETHODS,stringsAsFactors = FALSE)
  df.cdf = ddply(df,.(seed,scenario,method),get_cdf)

  if (pi0filter)
    df.cdf %<>% filter(pi0 < 0.55 & pi0 > 0.45)
  if (length(SEEDS) > 1)
    df.cdf %<>% group_by(x,method,scenario) %>% summarise(y = mean(y))

  df.cdf$scenario         = factor(df.cdf$scenario,levels = PLOTSCENARIOS)
  levels(df.cdf$scenario) = PLOTNAMES

  return(ggplot(df.cdf,aes(x = x,y = y,color = method),...) + colScale + 
    geom_line(lwd = 1.5,alpha = 0.7) + facet_grid(.~scenario) + 
      theme(legend.position = "bottom"))
}
```

Show the CDFs for all methods and all scenarios, based on a single data set.

```{r egcdf, dev='pdf',fig.height=3, fig.width=9, crop=TRUE}
plot_mean_cdf(1)
```

Show the CDFs for all scenarios, this time averaged over 100 simulated data 
sets.

```{r mean_cdf, dev='pdf',fig.height=3, fig.width=9, crop=TRUE}
plot_mean_cdf(1:100,PLOTMETHODS = c("ash.n","ash.u","ash.hu","truth"),
              pi0filter = TRUE)
```

Same as above, but with custom colours.

```{r mean_cdf_nopen, dev='pdf',fig.height=3, fig.width=9, crop=TRUE}
names(myColors) <- c("truth","ash.hu.s","ash.n.s","ash.u.s","qvalue",
                     "locfdr","mixfdr.tnull")
colScale <- scale_colour_manual(name = "method",values = myColors)
plot_mean_cdf(1:100,PLOTMETHODS = c("ash.n.s","ash.u.s","ash.hu.s","truth"),
              pi0filter = TRUE)

# Reset color scale for other plots.
source("../R/set_plot_colors.R") 
```

The following plots have fewer methods and scenarios for more clarity

```{r egcdf-reduce, fig.height=3, fig.width=6, crop=TRUE, dev='pdf'}
plot_mean_cdf(1,PLOTMETHODS = c("ash.n","truth"),
              PLOTSCENARIOS = c("spiky","near-normal","bimodal"))
```

Same as above, but this time the CDFs are averaged over 100 simulated data sets.

```{r mean_cdf-reduce, dev='pdf',fig.height=3, fig.width=6, crop=TRUE}
plot_mean_cdf(1:100,PLOTMETHODS = c("ash.n","truth"),
              PLOTSCENARIOS = c("spiky","near-normal","bimodal"),
              pi0filter = TRUE)
```

Now with custom colours.

```{r mean_cdf_nopen-reduce, fig.height=3, fig.width=6, crop=TRUE, dev='pdf'}
names(myColors) <- c("truth","ash.hu.s","ash.n.s","ash.u.s","qvalue","locfdr",
                     "mixfdr.tnull")
colScale        <- scale_colour_manual(name = "method",values = myColors)
plot_mean_cdf(1:100,PLOTMETHODS = c("ash.n.s","truth"),
              PLOTSCENARIOS = c("spiky","near-normal","bimodal"),
              pi0filter = TRUE)
source("../R/set_plot_colors.R")
```

I wanted to add the npmle method (fit using the GLmix function) to the plot...

```{r egcdf-reduce-with-npmle, dev='pdf',fig.height=4.5, fig.width=6, crop=TRUE}
run_nplme = function (SEED = 1,PLOTSCENARIOS = c("spiky","near-normal",
                                                 "flat-top","skew","bimodal")) {
  df = data.frame(seed = NULL,scenario = NULL,method = NULL,x = NULL,
                  y = NULL,pi0 = NULL)
  for (SCENARIO in PLOTSCENARIOS) {
    temp = load_example(dsc_shrink,SEED,SCENARIO,"ash.n","../output/dsc-shrink-files/")
    z    = GLmix(temp$input$betahat)
    df   = rbind(df,data.frame(seed = SEED,scenario = SCENARIO,method = "NPMLE",
                               x = z$x,y = cumsum(z$y)/sum(z$y),pi0 = NA))
  }
  return(df)
}

plot_mean_cdf_with_npmle = 
  function(SEED          = 1,
           PLOTMETHODS   = c("ash.n","ash.u","ash.hu","truth","mixfdr.tnull"),
           PLOTSCENARIOS = c("spiky","near-normal","flat-top","skew","bimodal"),
           pi0filter     = FALSE,...) {
  if (length(SEED) > 1)
    stop("plot with npmle only implemented for a single seed")
  PLOTNAMES = PLOTSCENARIOS

  # Set up dataframe with cdf for all methods and all datasets.
  df = expand.grid(seed = SEED,scenario = PLOTSCENARIOS,method = PLOTMETHODS,
                   stringsAsFactors = FALSE)
  df.cdf = ddply(df,.(seed,scenario,method),get_cdf)

  df.cdf$scenario         = factor(df.cdf$scenario,levels = PLOTSCENARIOS)
  levels(df.cdf$scenario) = PLOTNAMES

  df.npmle = run_nplme(SEED,PLOTSCENARIOS)
  df.cdf   = rbind(df.cdf,df.npmle)

  p = ggplot(df.cdf,aes(x = x,y = y,color = method),...) + 
    xlim(c(-4,4)) + colScale + geom_line(lwd = 1.2,alpha = 0.7) + 
    theme(legend.position = "bottom")
  return(p)
}

plot_mean_cdf_with_npmle(1,PLOTMETHODS = c("ash.hu","truth"),
                         PLOTSCENARIOS = c("spiky","near-normal","flat-top",
                                           "skew","bimodal")) +
  facet_wrap(~ scenario, nrow = 2)
```

## Session information

```{r info}
sessionInfo()
```